# Material and Methods

```{r load-files, warning=F, echo=F, include=F}
source("src/setup.R", local = knitr::knit_global())
source("src/mm-cleanup.R", local = knitr::knit_global())
```

```{r, include=FALSE}
if (file.exists("output/r-model.rds")) {
  r_model <- readRDS("output/r-model.rds")
} else {
  source("src/r-model.R", local = knitr::knit_global())
}
```


## Survey

Annually, since the spring in 2008, a citizen science survey about winter honey bee colony losses in Austria is conducted by the Institute of Biology at the University of Graz [@brodschneider2011]. It is available online, published in a beekeeping journal and distributed as paper submissions. The main focus of the survey are winter honey bee colony losses in Austria. The survey is part of the COLOSS questionnaire and most questions are in accordance with other participating countries [@vanderzee2013;@brodschneider2022]. Most of the questions are relate to possible risk factors which may influence overwintering success as seen in previous studies [e.g. @brodschneider2010a;@brodschneider2018;@oberreiter2020;@gray2020]. The survey is only obligatory for beekeepers who want to participate in the Austrian bee health program. The confirmation of participation is one of the options which allows a beekeeper to apply and/or get a higher rate of government aid for investments in technical equipment to improve production quality and hygiene in their beekeeping operation.^[AgrarMarkt Austria, 06/2021 - Version 04] The survey is in German and questions are translated for this study.

The survey is anonymous and private data like IP addresses were only used in the first step to help identify possible duplicate entries. In order to protect the personal data of the participants, a privacy statement was drawn between the primary persons involved in the data collection. If voluntary contact details were given, they were only used for enquiries to correct missing or incorrect data. All personal data was anonymised prior to the analysis.

As most questions are related to hive management and winter colony losses, I am only interested in a subset of those for my research question: number of colonies wintered (size of operation), number of colonies lost over the winter period (due to natural or queen loss, without losses due to natural disaster), state and district of the main wintering apiary (as a possible covariate), month(s) and type of treatment method used. In addition, two operational questions are also explored, if the beekeeper migrates with her/his colonies and if she/he is an organic certified beekeeper, with the possible answers 'Yes', 'No', and 'Uncertain'.

The answers to these questions of interest are explored in relation to the estimated treatment expenses per colony without labour costs. Participants could estimate the expenses with any positive number from zero without any upper limit. The question on treatment expense and motivation was only available in the online survey as the questionnaire, which are printed in a beekeeping journal and the handed out paper submissions, are shorter and more compact versions.

To further comprehend why participants choose a particular *Varroa* control method, the survey 2020/21 did additionally include the question: 'What are your main motivations in choosing your *Varroa* control method (max. 5)?'. The possible answers were displayed in a randomized order and included: 'Habit', 'Economic costs', 'Marketing', 'Subsidies', 'Prescription', 'Philosophy', 'Efficacy', 'Method applicable with, and without brood', 'Time of season where they can be applied', 'Colony strength', 'Colony infestation level (*Varroa*)', 'Environmental conditions', 'Recommendations from beekeeping associations, extension workers, veterinarians, magazines, etc.', 'Legal status (regulatory framework)', 'Availability of products', 'Resistances', 'Residues', 'Side effects on bees', 'Level of difficulty and needed equipment', 'Time and number of applications' and 'Other'. A total of `r length(mmList$motivation$no_answer)+12` participants did choose not to answer this question or did by accident choose more than five answers and were not included. 
<!-- the +12 are participants who could not finish the survey as they did answer more than 5 categories and had to be manually imported -->

As I analysed three continuous survey years and the survey is anonymous I have no assumption if the same beekeepers participated in more than one survey. Therefore, I can not tell which survey replies are independent over the years and which are not, and analysed them separately. This would only be possible for a handful of beekeepers who did consistently give the same contact details. The advantages of comparing multiple years is that I can get a more thorough overview of whether the survey is producing robust results, since colony expenditures and distribution, as well as apiary size, are not likely to change significantly over the survey years. Furthermore, overall winter loss rates fluctuate between years of the surveys. This means I can check if a given treatment method did result in low or high winter mortality or not in all survey years.

## Expenses Estimation

Expenses estimation was carried out to check if detailed desk research of costs would be comparable to participants' survey answers. To calculate the estimates, I researched online store prices in 2020 in Austria for officially registered medication[^mm-1] and used the treatment agent quantity per colony as given by manufacturer instructions (Table&nbsp;\@ref(tab:standard-cost)). The per colony estimation is based on equation \@ref(eq:estimate-calculation). With this equation an estimate was calculated for each survey participant with the given answer for treatment method (or combination), number of colonies and number of months used. No estimation was done for 'Drone brood removal', which has no obvious material expenses, and 'Another method', which is undefined.

[^mm-1]: Austrian medicinal product index: <https://aspregister.basg.gv.at/>, accessed November 2020

```{r standard-cost, include=TRUE}
c <- "Calculated estimates for treatment expenses per method in Euro. Investments (e.g. hyperthermia equipment) and material (e.g. queen cages) can be used long term and therefore a depreciation was calculated with 10 or 7 years. Consumables amount was taken by manufacturer instruction for a deep brood box (e.g. formic acid long term was calculated with 200ml use of formic acid)."
cn <- c("Methode", "Investment", "Material", "Consumables")
tab <- kbl(
  treatmentList[-1, c(3, 5:7)] %>%
    filter(tname != "Another method" & tname != "Drone brood removal") %>%
    mutate(
      across(is.numeric, ~ ifelse(.x == 0, "-", .x %>% round(2) %>% ft(nsmall = 2)))
    ),
  booktabs = TRUE,
  row.names = FALSE,
  align = c("l", rep("r", 3)),
  # format = "markdown",
  caption = c,
  col.names = cn,
  linesep = c("", "", "\\addlinespace"),
) %>%
  kableExtra::add_header_above(c(" " = 1, "divided by colonies" = 1, "per colony" = 2)) %>%
  footnote(threeparttable = TRUE, general = "'Synthetic methods' include Amitraz, Coumaphos, and other synthetic methods") %>%
  kable_styling(latex_options = "HOLD_position")
tab
rm(c, cn)
```

```{=tex}
\begin{equation}
  \begin{aligned}
    \text{estimate/colony} = \frac{\text{investment}}{\text{total colonies}} + \text{material} + \text{consumable} * \text{months}
  \end{aligned}
  (\#eq:estimate-calculation)
\end{equation}
```

## Data Cleanup

Data cleaning was performed for the analysis and exploration which included expenses as variable. As for the motivational analysis, all participants which did answer the question were used without any filtering. 

### Newspaper and Paper Submission

The short version of the survey, which was printed in a beekeeping journal and distributed as paper handout submissions, did not include the question about expenses and, therefore, was excluded. That concerns a total of `r sum(mmList$submitted_all %>% filter(submitted != "Internet") %>% pull(n)) %>% ft(.)` submission from `r sum(mmList$submitted_all$n) %>% ft(.)` over the three years of surveys. <!-- With one exception, as one person did send a filled out paper handout of the online survey. -->

### No Answers on Expenses

Not all participants answered the question on expenses, even when they could do so. These participants were therefore excluded from the expenses analysis. In numbers these are in 2018/19 a total of `r mmList$internet_no_answer %>% filter(year == "18/19") %>% pull(nn)` (`r mmList$internet_no_answer %>% filter(year == "18/19") %>% pull(np)`%); in 2019/20 a total of `r mmList$internet_no_answer %>% filter(year == "19/20") %>% pull(nn)` (`r mmList$internet_no_answer %>% filter(year == "19/20") %>% pull(np)`%) and in the last survey year 2020/21 a total of `r mmList$internet_no_answer %>% filter(year == "20/21") %>% pull(nn)` (`r mmList$internet_no_answer %>% filter(year == "20/21") %>% pull(np)`%) participants.

### No Treatment
<!-- r nrow(mmList$no_treatment %>% filter(costs == 0)) -->
<!-- r nrow(mmList$no_treatment %>% filter(costs > 0)) -->

Only two participants performed no treatment against the *Varroa* mite and answered with zero expenses. In addition, two participants answered that they did not perform any treatment but did include expenses nonetheless. These illogical entries and the before-mentioned no treatment answers, which would be not enough for a separate statistical analysis, were removed.

### No Treatment Methods
<!-- r nrow(mmList$no_method) -->

In total six participants did not specify their treatment methods but said they treated against *Varroa* and answered on *Varroa* treatment expenses (min. EUR&nbsp;`r fr(min(mmList$no_method$costs))`, max. EUR&nbsp;`r fr(max(mmList$no_method$costs))` per colony). I decided to discard them from the dataset, as it was only a small number and I could not use it for statistics which involved the treatment methods.

### Zero Expenses
<!-- r length(mmList$cost_zero$id_sponsor) -->

In total `r nrow(mmList$cost_zero$data)` participants said their treatment expenses were zero. Only one answer was kept, as the treatment was a combination of hyperthermia and biotechnical, and I cannot make an assumption about the real expenses of the participant. Of course the equipment for the hyperthermia does cost anything, but I do not have more details, as she/he could already own the equipment for a long time. A total of `r length(mmList$cost_zero$id_remove)-length(mmList$cost_zero$id_sponsor)` participants were removed as they stated zero expenses, but in fact did treat with organic acids, which is illogical and consumes material during each treatment. Additionally, four participants received a sponsorship (beekeeping club or local community) for their treatment expenses and therefore had zero expenses, which were also removed, as the treatment expenses are only shifted to the sponsor.

### Outlier Detection and Manipulation

To detect outliers in the dataset, two methods were used. The first method is based on upper extreme outliers only. For this I decided to filter the dataset based on the arbitrary decision if the expenses per colony is more or equal than double of the whole dataset upper extreme outlier point, which was an upper limit of EUR&nbsp;`r mmList$cost_upper$upper_limit`, calculated with equation \@ref(eq:extreme-outliers). This did detect `r nrow(mmList$cost_upper$data)` upper outliers based on the maximum limit and without any information on the given treatment method.

```{=tex}
\begin{equation}
\begin{aligned}
x >= (Q3 + 3 * IQR) * 2
\end{aligned}
(\#eq:extreme-outliers)
\end{equation}
```

The second approach was model based with the isolation forest approach [@liu2008], implemented in the R package 'isotree' [@isotree2021]. The input to the model is the expenses per colony and a dummy column encoded for each treatment method combination without drone brood removal. The parameters were set to contain subsamples on 256 rows, build 10 binary trees, and split on the combined two columns. I defined the standardized outlier score values above or equal to 0.6 as outliers [@liu2008]. This resulted in a list of `r nrow(mmList$cost_upper$forest)` anomaly values (Fig.&nbsp;\@ref(fig:isolation-forest)).

The two approaches did mostly overlap in their results and the lists were combined, which resulted in `r nrow(mmList$cost_upper$combined_list)` possible outliers. Further investigation of these outliers revealed that most of them did probably report total expenses per operation and not expenses per colony as asked in the survey. I removed participants from the outliers list who used 'Hyperthermia', which could involve considerable investment, especially if it was only used for one year. This was also stated by some beekeepers as comments in the survey. In addition participants were also removed from the outliers list if plausible reasons in the comments were given. As for the rest of the extreme outliers (`r nrow(mmList$cost_upper$new_data)`), I divided the expenses by the reported number of colonies going into winter. This means the manipulative impact was more notable for participants which included high expenses above the limit and also reported a large number of colonies and less for small operations, were the high expenses are possibly true (Appendix Fig.&nbsp;\@ref(fig:manipulated-costs)). The new calculated expenses were inspected to be plausible, e.g. not too low, and were updated in the cleaned dataset for further analysis.

```{r isolation-forest, include=T, fig.cap="Isolation forest standardized outlier score to expenses per colony of the whole dataset. Black points are within the defined average range and pink coloured points indicate possible outliers. Red background represents the area above the set average limit of >= 0.6 outlier score."}
include_custom("output/figs/isolation_forest.png")
```

## Final Data Count

After the aforementioned data cleaning steps, which filtered survey responses for answers on treatment expenses per colony and removed illogical answers, the number of responses was reduced by around 21-24% for the expense data analysis (Table&nbsp;\@ref(tab:reports)).

```{r reports, include=T}
cn <- c("Year", "Total [n]", "After Filtering [n]", "Percent [%]")
c <- "Final data count after filtering of survey responses on the question for estimated treatment expenses per colony. In comparison to the total number of survey participants."
tab <- knitr::kable(
  mmList$reports %>%
    mutate(
      year = paste0("20", year),
      across(is.numeric, ft)
    ),
  align = c("c", rep("r", 3)),
  booktabs = TRUE,
  row.names = FALSE,
  col.names = cn,
  caption = c
) %>%
  kable_styling(latex_options = "HOLD_position")
tab
rm(cn, c, tab)
```

## Statistics

The analysis of the data was carried out with R [@r2020], plotting, and data exploration was done mainly with the tidyverse package collection [@tidyverse2019]. The code, including a listing of the remaining packages not mentioned here, is available on GitHub under the MIT licence[^mm-2]. As the particular goal of the study is an exploratory data analysis (EDA) I try to rely predominantly on descriptive statistics and data visualisation. The use of inference based statistics and model generation is merely used for exploratory purposes and no focus on the *p*-value alone is given. Reported *p*-values are uncorrected for alpha error accumulation, this could lead to incorrectly rejecting the null hypothesis. Hypothesis confirmation statistics should normally only be used for a predefined hypothesis, which can be falsifiable, and experimental design directly related to the research question, which is untrue for this analysis. As I look at each observation/report multiple times, by definition I perform an exploratory analysis.

[^mm-2]: Github: <https://github.com/HannesOberreiter/treatment-expenses-austria>

### Operational Factor Analysis

The operational factor analysis (Migratory Beekeeper and Certified Organic Beekeeper) did include 'Yes', 'No' and 'Uncertain' as possible answers. The category 'Uncertain' and participants who did not answer the question were removed in the statistical two level factor analysis (Table&nbsp;\@ref(tab:factor-removed)). The *p*-value is generated from the median difference between groups in comparison to the permutation generated null distribution (5,000 permutations), significant results are *p* $\leq$ 0.05. For the median point estimate, a confidence interval (95%) was calculated based on 1,000 bootstrap resamples. The visual statistical test results of the difference between groups are attached in the appendix section.

### Multiple Regression

To evaluate if and how well the selected questions could be used for prediction and explain the treatment expenses, a simple linear multiple regression model with treatment expenses per colony as dependent variable was fitted. The model creation and evaluation was done with the tidymodels package [@tidymodels2020]. Data preprocessing included normalization with $\log_{10}$ transformation of number of colonies wintered and expenses of treatment to remove the effect of the right skewed data distribution, whereas beforehand one single zero expenses answer was removed (Fig.&nbsp;\@ref(fig:model-qq)). Same as in the two level operational factor analysis in the multiple regression models, only participants which did answer 'Yes' or 'No' for the two operational factor questions were used in the model to simplify it (Table&nbsp;\@ref(tab:factor-removed)). The levels were dummy transformed to 0 (False) for 'No' and 1 (True) for 'Yes'. Treatment methods were binary transformed, if the participants did practice it in any month to 1 (True) otherwise if not at all applied to 0 (False). After preprocessing and filtering a total of `r nrow(r_model$data) %>% ft()` answers were left. The model fitting on the dataset was done with 90% and out-of-sample evaluation of accuracy was done with 10%. Best model selection was conducted based on lowest akaike information criterion (AIC) and bayesian information criterion (BIC).

(ref:model-qq) QQ-Plot of $\log_{10}$ transformed for normalization of (A) treatment expenses per colony and (B) number of honey bee colonies wintered.
```{r model-qq, include=T, fig.cap="(ref:model-qq)"}
include_custom("output/figs/model-qq.png")
```

```{r factor-removed, include=T}
# cn <- c("Question", "Year", "Participants [n]", "Answer")
cn <- c("Year", "Participants [n]", "Answer")
c <- "Number of participants answering with 'Uncertain' or not at all for the operational questions. These beekeepers were removed in the statistical two factor analysis and multiple regression model."
tab <- knitr::kable(
  mmList$single_factor_drop %>%
    mutate(
      year = paste0("20", year),
      year = ifelse(lag(year, default = "0") == year, "", year)
    ) %>%
    # select(name, year, n, value),
    select(year, n, value),
  booktabs = TRUE,
  row.names = FALSE,
  col.names = cn,
  align = c("c", "r", "l"),
  caption = c
) %>%
  pack_rows("Certified Organic Beekeeper", 1, 6) %>%
  pack_rows("Migratory Beekeeper", 7, 9) %>%
  kable_styling(latex_options = "HOLD_position")
tab
rm(cn, c, tab)
```

### Treatment Methods and Loss Rates

The treatment methods were selected in accordance with COLOSS, and the same list of methods is used in every participating country [@vanderzee2013]. In the survey, a total of 18 different *Varroa* treatment methods were given as choices with time of application as month. The treatment methods included 'Drone brood removal', 'Oxalic acid' (sublimation and mixture or pure trickling), 'Formic acid' (long or short term), 'Another biotechnical method' (excluding drone brood removal and hyperthermia), 'Thymol', 'Hyperthermia', 'Lactic acid', 'Another method' and 'Synthetic methods' which are a combination of the treatment categories 'Amitraz (in strips or sublimation), Coumaphos (Perizin or Checkmite+) or Flumethrin (Bayvarol, Polyvar), Tau-fluvalinate (Apistan)', because of the limited number of replies in Austria concerning these methods [@oberreiter2020].

To generate groups of treatment method combinations, the individual months of the treatment applied were grouped into the following seasons 'Spring' (April-May), 'Summer' (June-October) and 'Winter' (November-March). In addition to reducing the number of possible different treatment method combinations, the method 'Drone brood removal' was excluded, as it led to too many combinations and should at least not influence the expenses.

The observed total loss rate was calculated based on equation \@ref(eq:lossrate). The reported 95% confidence intervals (CI) were created for each analysed treatment combination from a separate null model, which was calculated using a Generalised Linear Model (GZLM) with quasi-binomial distribution and the link function \enquote{logit}. The procedure was carried out according to scientifically established methods [@vanderzee2013].

```{=tex}
\begin{equation}
\begin{aligned}
\frac
    {
      \sum{\text{Loss of colonies}}-\sum{\text{Loss due to natural hazards}}
    }
    {
      \sum{\text{Colonies wintered}}
    }
      *100
\end{aligned}
(\#eq:lossrate)
\end{equation}
```


### Effectiveness & Economic

To test if *Varroa* treatment combinations, with at least 30 answers across all three years, are effective and also economic, the observed loss rates and reported mean expenses per treatment combination were plotted and grouped into nine categories. The categories comprise the combination of low, intermediate or high losses (observed colony winter losses) and low, intermediate or high expenses (expenses per colony) for each method combination (Fig.&nbsp;\@ref(fig:efficient-economic-mm)). The calculation and assignment to these categories are based on the value for each combination in comparison to the quantile distribution of the observed loss rate and mean expense per colony of all combinations. If the mean expense for a method is equal or between the middle 33-57% quantile range, it will get the category 'Intermediate'. If it is lower it will be declared as 'Low-Expense' and if higher 'High-Expense'. The same method was used for the the observed loss rate with the categories 'Intermediate', 'Low-Loss' and 'High-Loss'. The separation for each year was performed because the loss rate does change between years [@brodschneider2019a;@oberreiter2020]. Although, the expenses could probably be summarised over all three years, as they are not expected to change drastically in this timeframe. However, I can not make assumption as to whether the same beekeepers employed the same methods and participated in more than one survey, which could bias the results.

<!--
# we could do the geometric mean for costs to reduce effects of outliers, but the result is not straightforward interpretable and the mean is maybe more true in this case, as we already to a lot of pre filtering to reduce the outliers.
-->

(ref:efficient-economic-mm) Graphical explanation of the assignment of effectiveness and economic categories for the treatment method combinations. This assignment was done for each year separately.
```{r efficient-economic-mm, include=T, fig.cap="(ref:efficient-economic-mm)"}
include_custom("output/figs/efficient-economic-mm.png")
```