# Material and Methods

```{r mm-load, warning=F, echo=F, include=F}
source("partials/mm-cleanup.R", local = knitr::knit_global())
```

## Survey

The survey was a citizen science survey which is done each year in spring since 2008 and is a combination of online, beekeeping journal and paper submission. The main focus and question of the survey are the yearly colony losses over winter for the Austrian Beekeeper. It is part of the COLOSS questionnaire [@vanderzee2013]. The questions are simple and relate to possible risk factors which may influence overwintering success and as seen in other studies [@oberreiter2020]. The survey is not mandatory but beekeepers who want to get more official support they must participate. 

As most questions are related to winter colony losses we are only interested in an subset of questions: Number of colonies wintered (size of operation), Number of colonies lost over winter period (due natural or queen loss, without losses due to natural disaster), month(s) and type of treatment used, estimated treatment expenses per colony without labor costs. The last question about the expenses, which is our main focus, was only available in the online survey as the questionnaire printed in the beekeeping journal and handed out paper submissions are a shorter version. The survey was anonymous and private data like IP addresses were only used in the first step to help identify double entries. If voluntary contact details were given, the data was anonymized prior analysis and only used if further enquire on questionable answers was necessary.

As we analyzed two continuous survey years and the survey is anonymous we have no assumption if the same beekeepers participated in both surveys. Therefore, we can not say that both years are independent or not and analyzed them separately. The advantages to compare two years is, that we may have a better overview if the survey is representative, as there should not much change in the expenses and distribution of beekeepers apiary location and size of operations in the two surveys.

## Expenses Estimation

Expenses estimation was carried out for data control of given answers in the survey and to check if a logical expenses estimation of products yield the same result as the participants answers. To calculate estimates we researched online store prices in 2020 in Austria for official registered medication^[Austrian medicinal product index: https://aspregister.basg.gv.at/, accessed November 2020] and used common application amount and methods in Austria from beekeeping school materials and experience from previous surveys to calculate an estimate for each treatment method an expenses per colony, see table \@ref(tab:standard-cost) for our calculated estimate. The calculation with average month of usage and estimated 11 colonies can be seen in \@ref(tab:common). After that for each survey participants an estimate was calculated with the given answer for treatment method (or combination), number of colonies and number of months used.

```{r standard-cost, include=T}
c <- "Our calculated estimates for treatment expenses per method. Investments can be used long term (we calculate 10 years, eg. Hyperthermia) and are divided by the number of colonies. Material can also be used long term (7 years, eg. Queen Cages) but needs to be bought for each colony. Consumables are used for each colony (eg. Formic Acid Long Term, 200ml)."
cn = c("Methode", "Investment", "Material", "Consumables")
kable(
  treatmentList[-1, c(3, 5:7)],
  booktabs = TRUE, row.names = FALSE,
  caption = c, col.names = cn
) %>%
  footnote(general = "Synthetic methods are combined and include Amitraz, Coumaphos, and other synthetic methods'; Other biotechnical methods do not include drone brood removal or hypothermia; and Another method excludes synthetic methods.")

rm(c, cn)
```

```{r common, include=T}
# Get median number of colonies to calculate an estimate
m_colonies = median(DATA$hives_winter)

col <- DATA %>% 
    select(!!(paste0(treatmentList$ttotal, "12")))
# Get number of participates using the single treatment methods
sin <- col %>% 
    mutate(across(.cols = everything(), as.logical)) %>% colSums()
# get the average amount for number of months
nu <- col %>% na_if(0) %>% colMeans(na.rm=T)

common_single <- tibble(
  tname = treatmentList$tname,
  n = sin,
  p = round(sin*100/nrow(DATA),2),
  nu = round(nu, 2),
  est = round(treatmentList$investment / m_colonies + treatmentList$material + treatmentList$consumables*nu, 2)
) %>% arrange(desc(p))

c <- paste0("Most used Methods in the two survey years, average months indicating the average of months this methods was used. Estimated costs are calculated with ", m_colonies, " colonies, which represents the median number of colonies in our survey, and the average months of usage.")

cn = c("Method", "Total [n]", "Percent [%]", "Average Months[n]", "Estimated Costs")
kable(
  common_single[-1,],
  booktabs = TRUE, row.names = FALSE,
  col.names = cn,
  caption = c
) # %>%   kable_classic()
rm(cn, c, common_single, nu, sin, col, m_colonies)
```


## Data Cleanup

### Newspaper and Paper Submission

The short version of the survey, which was printed in a beekeeping journal and paper handout submissions, which we received via postal service, did not include the question about expenses and, therefore, were excluded these answers from the analysis. That concern a total of `r sum(COUNT_PAPER$n[COUNT_PAPER$submitted=="Paper" | COUNT_PAPER$submitted=="Newspaper"])` answers from `r sum(COUNT_PAPER$n[COUNT_PAPER$submitted=="Paper" | COUNT_PAPER$submitted=="Newspaper"]) + nrow(RAW)` over the two years of survey.

### No Treatment

In the survey `r nrow(NO_TREATMENT)` participants did answer "No treatment against varroa" and did give no answers for any treatment method, but included expenses for varroa treatment. These `r nrow(NO_TREATMENT)` illogical entries were removed from our data analysis.

### No Methods

Because we want to analyze treatment methods in combination with the costs, we did also remove `r nrow(NO_METHOD)` participants which did answer treatment expenses in the survey, but were not specific which treatment methods they used. 

### Zero Expenses

In total `r nrow(COST_ZERO)` participants said their treatment expenses were zero. Only `r length(id_keep)` answer were kept, as the treatment was a combination of hyperthermia and biotechnical and we can make no assumption about the real expenses of the participant. Ofcourse the equipment for the hyperthermia do cost something, but we don't know more details, as he could already have the equipment for a long time. `r length(id_remove)-length(id_sponsor)` were removed as they said zero expenses but actually did treatment with organic acids, which is illogical as the organic acids are consumed and not for free. In addition `r length(id_sponsor)` participants got a sponsorship (beekeeping club or local community) for their treatment expenses, which were also removed.

### High Expenses

We defined an upper limit cost outliner with the 3rd quantile and multiplied it by two which resulted in a cost limit of `r upper_limit`. In total `r nrow(COST_UPPER_NO_ESTIMATE)` participants did answers above this limit. To remove possible plausible answers, we did filter the resulting samples to only include expenses which are double of our previous calculated estimated expense for the individual beekeeper. This did reduce the survey answers in question to `r nrow(COST_UPPER)`. Upon investigating the `r nrow(COST_UPPER)` answers, we found out that most of them did answer total expenses per operation and not expenses per colony. Therefore, we divided the expenses by the number of colonies going into winter. Only `r length(id_nochange)` participants were not changed, as one only had one colony and did hyperthermia treatment and the second one did explicit explain the high expenses as comment.

## Statistics

The analysis of the data was carried out with R [@r2020], plotting and data manipulation was done mainly with Tidyverse [@tidyverse2019]. Packages are version controlled with Renv [@renv2020]. For full reproducibility the code and a list of all other packages are available on GitHub^[Github: https://github.com/HannesOberreiter/treatment-expenses-austria]. The confidence interval (CI) for expenses per colony was generated with adjusted bootstrap percentile (BCa) method. Loss rate and corresponding CI were compute with a quasibinominal generalized linear model (GZLM) and link “logit” function [@vanderzee2013]. To minimize the variation in the varroa control methods and their combination, all months were grouped together. 


For the analysis of the combination of different varroa control methods, a usage histogram for each method was generated and grouped into spring, summer, and winter. After that, a vector of all applied methods with at least 15 answers and without drone brood removal was generated to minimize calculation time and to generate statistically relevant results. 

Mapping of participants main wintering apiary location was done with modified k-means cluster search method [@oberreiter2020].


```{r}
# Cleanup Env
rm(list=ls()[
  !(ls() %in%
  c(
    "RAW", "DATA", "fBootMean", 
    "fBootMedian", "colorBlindBlack8", "stats_aut", 
    "treatmentList", "fMapCluster"
    )
  )])
```




