# Material and Methods

```{r mm-load, warning=F, echo=F, include=F}
source("partials/mm-cleanup.R", local = knitr::knit_global())
```

## Survey

The survey was a citizen science survey which is done each year in spring since 2008 and is a combination of online, beekeeping journal and paper submission. The main focus and question of the survey are the yearly colony losses over winter for the Austrian Beekeeper. It is part of the COLOSS questionnaire [@vanderzee2013]. The questions are simple and relate to possible risk factors which may influence overwintering success and as seen in other studies [@oberreiter2020]. The survey is not mandatory but beekeepers who want to get more official support they must participate. 

As most questions are related to winter colony losses we are only interested in an subset of questions: Number of colonies wintered (size of operation), Number of colonies lost over winter period (due natural or queen loss), time (in months) and type of treatment used, estimated treatment cost per colony without labour. The last question about the cost, which is our main focus, was only available in the online survey as the questionnaire printed in the keeping journal and handed out paper submission are a shorter version.

The survey is anonyme if contact data was present it was removed prior analysis and only used if further investigation was necessary.

As we analyse two continuous survey years and the survey is anonyme we have no assumption if same beekeepers participated in both survey. Therefore, we analyse both year separately as we can not say that both years are independent or not. The good part is that with two years with have a better overview if the survey is representative as there should not be much change in the costs question and distribution of operations.

## Cost Estimation

For data control of answers and to check if logical price estimation of products can be used to calculate the beekeepers expenses. We used current prices in Austrian and German Beekeeping Online Stores for registered Medication and common application amount and methods in Austria to calculate for each treatment methods a cost per colony and operation. For each survey participants an estimate was calculated with the given answer for method, number of colonies and number of months used. Results of these Estimations in Comparison to the survey, please check section \@ref(r-estimations).

## Data Cleanup

### Newspaper and Paper Submission

The short version of the survey, which was printed a beekeeping journal and were a paper print out which was send via postal service, did not include the question about costs and therefore were excluded from the analysis. That concern a total of `r sum(COUNT_PAPER$n[COUNT_PAPER$submitted=="Paper" | COUNT_PAPER$submitted=="Newspaper"])` answers from `r sum(COUNT_PAPER$n[COUNT_PAPER$submitted=="Paper" | COUNT_PAPER$submitted=="Newspaper"]) + nrow(RAW)` over the two years of survey.

### No Treatment

In the survey `r nrow(NO_TREATMENT)` participants did answer "No Treatment against Varroa" and did give no answers for any treatment method, but included costs for Varroa treatment. These `r nrow(NO_TREATMENT)` illogical entries were removed from our data analysis.

### No Methods

Because we want to analyze treatment methods in combination with the costs, we did also remove `r nrow(NO_METHOD)` participants which did answer treatment cost in the survey, but were not specific which methods they used. 

### Zero Costs

In total `r nrow(COST_ZERO)` participants said their treatment costs were zero. Only `r length(id_keep)` answers were kept, as the treatment was a combination of hyperthermia and biotechnical. We have no assumption. Ofcourse the equipment for the hyperthermia do costs something, but we don't know more details. `r length(id_remove)-length(id_sponsor)` were removed as they said zero costs but actually did treatment with organic acids. In addition `r length(id_sponsor)` participants got a sponsorship (eg from beekeeping club or local community) for their treatment costs, which were also removed from the cost analysis dataset.

### High Costs

We defined upper limit cost outliner with 3rd quantile and multiplied by 2 which resulted in a cost limit of `r upper_limit`. In total `r nrow(COST_UPPER_NO_ESTIMATE)` participants did answers above this limit. To remove possible plausible answers, we did filter the resulting samples to only include costs which are double of our estimated costs, this did reduce the survey answers in question to `r nrow(COST_UPPER)`. Upon investigating the `r nrow(COST_UPPER)` answers, we found out that most of them did answer a total cost and no cost by colony. Therefore, we divided the costs by the number of colonies going into winter. Only `r length(id_nochange)` participants were not changed, as one only had one colony and did hyperthermia treatment and the second one did explicit explain the high costs as comment.


## Statistics

The analysis of the data was carried out with R [@r2020] and plotting and data manipulation was done mainly with Tidyverse [@tidyverse2019]. Packages are version controlled with Renv [@renv2020]. For full reproducibility the code and a list of all other packages are available on GitHub LINK.

The confidence interval (CI) for cost estimation were generated with adjusted bootstrap percentile (BCa) method. Loss rate and corresponding CI were computed ere computed with a quasibinominal generalized linear model (GZLM) link “logit” function [@vanderzee2013].

To minimize the variation in the varroa control methods, all months were grouped together 

For the analysis of the combination of different varroa control methods, a usage histogram for each method was generated and grouped into spring, summer, and winter. After that, a vector of all applied methods with at least 15 answers and without drone brood removal was generated to minimize calculation time and to generate statistically relevant results. 


Aggregation of apiaries on the map was conducted with k-means cluster search method. To improve the accuracy in areas with lower density, the resulting clusters, which included only a single apiary or a low number of apiaries and high within cluster sum of squares, were removed and the original geolocation of the affected apiaries were used for plotting.

```{r}
# Cleanup Env
rm(list=ls()[
  !(ls() %in%
  c(
    "RAW", "DATA", "fBootMean", 
    "fBootMedian", "colorBlindBlack8", "stats_aut", 
    "treatmentList", "fMapCluster"
    )
  )])
```




