# Material and Methods

```{r load-files, warning=F, echo=F, include=F}
source("src/setup.R", local = knitr::knit_global())
source("src/mm-cleanup.R", local = knitr::knit_global())
```

```{r, include=FALSE}
if (file.exists("output/r-model.rds")) {
  r_model <- readRDS("output/r-model.rds")
} else {
  source("src/r-model.R", local = knitr::knit_global())
}
```


## Survey

The survey is a citizen science survey which is done each year in spring since 2008. It is a combination of online, beekeeping journal and paper submission. The main focus and question of the survey is the yearly colony losses over winter for the Austrian beekeepers'. The survey is part of the COLOSS questionnaire and most questions are in accordance with other participating countries [@vanderzee2013]. The questions are simple and relate to possible risk factors which may influence overwintering success as seen in previous studies in Austria [@oberreiter2020;@brodschneider2010a]. The survey is only obligatory for beekeepers who want to participate in the Austrian bee health programme. The confirmation of participation is one of the options which allows a beekeeper to apply and/or get a higher rate of government aid for investments in technical equipment to improve production quality and hygiene in their beekeeping operation.^[AgrarMarkt Austria, 06/2021 - Version 04]

As most questions are related to winter colony losses I am only interested in an subset of questions for my research question: number of colonies wintered (size of operation), number of colonies lost over the winter period (due natural or queen loss, without losses due to natural disaster), state and district of main wintering apiary (as possible covariate), month(s) and type of treatment used. In addition two operational questions are also explored, if the beekeeper migrates with his colonies and if he/she is an organic certified beekeeper, with the possible answers "Yes", "No" and "Unsure". The answers of these questions are explored in combination with the answer on the question about estimated treatment expenses per colony without labor costs. Participants could answer this questions with any positive number from zero and without any upper limit. The treatment expenses questions was only available in the Austrian COLOSS questionnaire and only in the online survey as the questionnaire which are printed in the beekeeping journal and handed out paper submissions are shorter versions.

To further understand why participants choose their varroa control method, the last survey year (winter 2020/21) did also include the question: "Which main motivations are most likely to apply to you when choosing the Varroa control method (max. 5)?". The possible answers, which were arbitrary defined, were displayed in a randomized order and included: "Habit", "Economic costs", "Marketing", "Subsidies", "Prescription", "Philosophy", "Efficacy", "Method applicable with and without brood", "Time of season where they can be applied", "Colony strength", "Colony infestation level (Varroa)", "Environmental conditions", "Recommendations from beekeeping associations, extension workers, veterinarians, magazines, etc.", "Legal status (regulatory framework)", "Availability of products", "Resistances", "Residues",  "Side effects on bees", "Level of difficulty and needed  equipment", "Time and number of applications", "Other". This question as only available in the online survey. A total of `r length(mmList$motivation$no_answer)` participants did choose not to answer this question or did by accident choose more than five answers, which leaves `r mmList$motivation$valid_answers` valid answers to this new question.

The survey was anonymous and private data like IP addresses were only used in the first step to help identify possible double entries. In order to protect the personal data of the participants, a privacy statement was drawn up between the main persons involved in the evaluation. If voluntary contact details was given, it was only used for enquiries to correct missing or incorrect data. All personal data was anonymised prior analysis. 

As I analysed three continuous survey years and the survey is anonymous I have no assumption if the same beekeepers participated in more than one survey. Only for the handful of beekeepers who did always give the same contact details. Therefore, I can not say that the years are independent or not and analysed them separately or use year as random effect. The advantages to compare multiple years is, that I may have a better overview if the survey is representative, as there should not be much change in the expenses and distribution of beekeepers apiary location and size of operations over the survey years. Furthermore, overall winter loss rates fluctuate between years of survey. This means I can check if a given treatment method did result in all survey years in low or high winter mortality or not.

## Expenses Estimation

Expenses estimation was carried out to check if there are discrepancies between participants answers and estimate. To calculate the estimates I researched online store prices in 2020 in Austria for official registered medication[^mm-1] and used common application amount from beekeeping school materials and manufacturer instructions (Tab. \@ref(tab:standard-cost)). An example calculation, based on \@ref(eq:estimate-calculation), with the given estimates, with average month of usage and an estimated of 10 colonies can be seen in table \@ref(tab:common). 

```{=tex}
\begin{equation}
\begin{aligned}
  \text{estimate/colony} = (\text{investment} / \text{total colonies}) + \text{material} + \text{consumable} * \text{months}
\end{aligned}
(\#eq:estimate-calculation)
\end{equation}
```

With this method an estimate was calculated for each survey participants with the given answer for treatment method (or combination), number of colonies and number of months used.

[^mm-1]: Austrian medicinal product index: <https://aspregister.basg.gv.at/>, accessed November 2020

```{r standard-cost, include=T}
c <- "The calculated estimates for treatment expenses per method. Investments can be used long term (we calculate 10 years, eg. hyperthermia equipment) and are divided by the number of colonies. Material can also be used long term (7 years, eg. queen cages) but needs to be bought for each colony. Consumables are used for each colony (eg. formic acid long term was calculated with 200ml use of formic acid)."
cn <- c("Methode", "Investment", "Material", "Consumables")
kable(
  treatmentList[-1, c(3, 5:7)],
  booktabs = TRUE,
  row.names = FALSE,
  format = "markdown",
  caption = c,
  col.names = cn
) %>%
  footnote(threeparttable = T, general = "Synthetic methods are combined and include Amitraz, Coumaphos, and other synthetic methods; Other biotechnical methods do not include drone brood removal or hypothermia; Another method excludes synthetic methods.")

rm(c, cn)
```

```{r common, include=T}
# colonies based for estimate
m_colonies <- 10
col <- dfData %>%
  select(!!(paste0(treatmentList$ttotal, "12")))
# Get number of participates using the single treatment methods
sin <- col %>%
  mutate(across(.cols = everything(), as.logical)) %>%
  colSums()
# get the average amount for number of months
nu <- col %>%
  na_if(0) %>%
  colMeans(na.rm = T)

common_single <- tibble(
  tname = treatmentList$tname,
  n = sin,
  p = round(sin * 100 / nrow(dfData), 2),
  nu = round(nu, 2),
  est = round(treatmentList$investment / m_colonies + treatmentList$material + treatmentList$consumables * nu, 2)
) %>% arrange(desc(p))

c <- paste0("Treatment methods ordered by usage over the survey years, months indicating the average of months the corresponding method was used. Estimated costs per colony are calculated with ", m_colonies, " colonies, which represents the median number of colonies in the survey, and the average months of usage.")

cn <- c("Method", "Total [n]", "Percent [%]", "Average Months [n]", "Estimated Costs")
kable(
  common_single[-1, ],
  booktabs   = TRUE,
  row.names = FALSE,
  col.names  = cn,
  format = "markdown",
  caption    = c
) # %>% kable_classic()
rm(cn, c, common_single, nu, sin, col, m_colonies)
```

## Data Cleanup

### Newspaper and Paper Submission

The short version of the survey, which was printed in a beekeeping journal and paper handout submissions, did not include the question about expenses and, therefore, were excluded. That concern a total of `r sum(mmList$submitted_all %>% filter(submitted != "Internet") %>% pull(n))` submission from `r sum(mmList$submitted_all$n)` over the three years of survey. Though there is one exception which was kept to this, as one person did send a filled out paper handout of the online survey.

### No Answers on Expenses

Not all participants did answers the question on expenses even when they had the possibility to do so. These participates were also excluded as this question is the main focus of the study. In numbers these are in 2018/19 a total of `r mmList$internet_no_answer %>% filter(year == "18/19") %>% pull(nn)` (`r mmList$internet_no_answer %>% filter(year == "18/19") %>% pull(np)`%), 2019/20 a total of `r mmList$internet_no_answer %>% filter(year == "19/20") %>% pull(nn)` (`r mmList$internet_no_answer %>% filter(year == "19/20") %>% pull(np)`%) and in the last survey year 2020/21 a total of `r mmList$internet_no_answer %>% filter(year == "20/21") %>% pull(nn)` (`r mmList$internet_no_answer %>% filter(year == "20/21") %>% pull(np)`%) answers.

### No Treatment

Only `r nrow(mmList$no_treatment %>% filter(costs == 0))` participants did no treatment against the varroa mite and answered with zero expenses. In addition `r nrow(mmList$no_treatment %>% filter(costs > 0))` participants answered that they did no treatment but have included expenses for varroa treatment. This illogical entries and the participants with zero costs and the before mentioned no treatment answers were removed from the data analysis. 

### No Methods

In total `r nrow(mmList$no_method)` participants did answer varroa expenses (min. `r min(mmList$no_method$costs)`, max. `r max(mmList$no_method$costs)` Euro per colony) and said they treated against varroa but did not specify their treatment method. I decided to remove them from the dataset, as it was only a small number and I could not further validate the expenses answers or use it for the statistics which involved the treatment methods.

### Zero Expenses

In total `r nrow(mmList$cost_zero$data)` participants said their treatment expenses were zero. Only one answer were kept, as the treatment was a combination of hyperthermia and biotechnical and I can make no assumption about the real expenses of the participant. Ofcourse the equipment for the hyperthermia do cost something, but I do not know more details, as he could already have the equipment for a long time. A total of `r length(mmList$cost_zero$id_remove)-length(mmList$cost_zero$id_sponsor)` were removed as they said zero expenses but actually did treatment with organic acids, which is illogical as the organic acids are consumed and not for free. Additionally `r length(mmList$cost_zero$id_sponsor)` participants got a sponsorship (beekeeping club or local community) for their treatment expenses and had therefore zero expenses, which were also removed.

### Outlier detection and imputation

To detect outliers in the dataset two methods were used. For the upper limit I decided to filter the dataset based on the arbitrary decision if the expenses per colony is equal or more than double of the whole dataset extreme outlier (`r mmList$cost_upper$upper_limit` Euro), as defined in \@ref(eq:extreme-outliers).

```{=tex}
\begin{equation}
\begin{aligned}
  x >= (Q3 + 3 * IQR) * 2
\end{aligned}
(\#eq:extreme-outliers)
\end{equation}
```

This did detect `r nrow(mmList$cost_upper$data)` anomaly values based on the upper limit and without any information of the given treatment method. The second one is model based with the isolation forest approach [@liu2008], implemented in the R package [@isotree2021]. Input for the model was the expenses per colony and a dummy column which was encoded for each treatment method combination without drone brood removal. The parameters were set to take all rows of our dataset, 100 binary trees and split on a single variable. I defined values above 0.6 in the resulting standardized outlier score, as described in @liu2008, as anomaly values, which resulted in a list of `r nrow(mmList$cost_upper$forest)`. The two approaches did mostly overlap in their results and the lists were combined, which resulted in `r nrow(mmList$cost_upper$combined_list)` possible outliers. Further investigating of the outlier answers revealed that most of them did answer total expenses per operation and not expenses per colony as asked in the survey. As this is only an assumption, I removed from the outliers list participants which used Hyperthermia as treatment method, which could involve big investment, especially if it was only used for one year. This was also stated by some beekeepers as comments in the survey. As for the rest of the extreme outliers (*n* = `r nrow(mmList$cost_upper$new_data)`), I divided for each answers the expenses by the reported number of colonies going into winter. This means the manipulative impact was higher for participants which had high expenses above the limit and also reported a large number of colonies and less for small operations, were the high expenses are possible true. The new calculated costs were inspected to be plausible, eg. not too low, and were updated in the cleaned dataset for further analysis.

```{r isolation-forest, include=T, fig.cap="Isolation forest standardized outlier score to expenses per colony of the whole dataset. Black points are within the defined average range and pink coloured points indicating possible outliers. Red background area above the set average limit of >= 0.6 outlier score."}
include_custom("output/figs/isolation_forest.png")
```

### Final Data Count

After the afore mentioned data cleaning steps, which filtered survey responses with valid answers for expenses per colony, the number of answer was reduced by around 21-24% (Tab. \@ref(tab:reports)).

```{r reports, include=T}
cn <- c("Year", "Total [n]", "Valid Answers [n]", "Percent [%]")
c <- "Number of participants with valid answers on the question for estimated treatment expenses per colony and the total number of survey participants."
tab <- knitr::kable(
  mmList$reports %>%
    mutate(
      year = paste0("20", year)
    ),
  booktabs = TRUE,
  row.names = FALSE,
  col.names = cn,
  format = "markdown",
  caption = c
)
tab
rm(cn, c, tab)
```

## Statistics

The analysis of the data was carried out with R [@r2020], plotting and data manipulation was done mainly with the tidyverse package collection [@tidyverse2019]. For full reproducibility the code and a list of all other packages are available on GitHub[^mm-2]. As the main goal of the study is an exploratory data analyses (EDA) I try to rely mostly on descriptive statistics and visualise the data behind it. The use of inference based statistics is only used for this explorative purpose and no focus on the *p*-value is given, but as it is still common the *p*-value will be reported. Inference based statistics should normally only be used for a predefined hypothesis and experimental design directly related to the research question. 

[^mm-2]: Github: <https://github.com/HannesOberreiter/treatment-expenses-austria>

### Single Factor Analysis

The operational factor question analysis (Migratory Beekeeper and Certified Organic Beekeeper) did include as possible answers "Yes", "No" and "Uncertain". The category "Uncertain" and participants which did not answers the question were not used in the single factor statistical analysis (Tab. \@ref(tab:factor-removed)). The *p*-value is generated from the median difference between groups in comparison to the permutation generated null distribution (5,000 permutations). For the median point estimate a confidence interval (95%) was calculated based on 1,000 bootstrap resamples. The visual statistical test results of the the difference between groups is attached in the appendix section.

### Multiple Regression - Prediction

To evaluate if and how well the selected questions can be used for prediction of the treatment expenses a simple linear multiple regression model with treatment expenses per colony as dependent variable was fitted. The model creation and evaluation was done with the tidymodels package [@tidymodels2020]. Data preprocessing included normalization with $log_10$ transform of number of colonies wintered and expenses of treatment, whereas beforehand one single zero expenses answer was removed (Fig. \@ref(fig:model-qq)). Same as in the single factor analysis to simplify the model interpretation only participants which did answer "Yes" or "No" for the two operational factor questions were used in the model and dummy transformed to 0 (False) if the answer was no and 1 (True) if the answer was yes (Tab. \@ref(tab:factor-removed)).

(ref:model-qq) QQ-Plot of $\log_10$ transformed normalization of (A) treatment expenses per colony and (B) number of honeybee colonies wintered.
```{r model-qq, include=T, fig.cap="(ref:model-qq)"}
include_custom("output/figs/model-qq.png")
```

Treatment methods were binary transformed, if the participants did use it in any month to 1 (True) otherwise if not at all applied to 0 (False). Model fitting was done with 90% of the original dataset (n = `r nrow(r_model$train_data)`) and out of sample evaluation of accuracy was done with 10% (n = `r nrow(r_model$test_data)`. Best model selection was selected based on lowest AIC and BIC value. Importance plot of variables based on @vip2020.

```{r factor-removed, include=T}
cn <- c("Question", "Year", "Participants [n]", "Answer")
c <- "Number of participants which were removed in the single factor analysis and multiple regression model."
tab <- knitr::kable(
  mmList$single_factor_drop %>%
    mutate(
      year = paste0("20", year)
    ) %>%
    select(name, year, n, value),
  booktabs = TRUE,
  row.names = FALSE,
  col.names = cn,
  align = c("r", "c"),
  caption = c
) %>%
  kableExtra::collapse_rows(columns = 1:2, valign = "top")
tab
rm(cn, c, tab)
```

### Loss Rate

Loss rate and corresponding CI were compute with a quasibinominal generalized linear model (GZLM) and link "logit" function [@vanderzee2013], which corresponds to the observed total loss rate in each group. To minimize the variation in the varroa control methods and their combination, the drone brood removal was removed, as it led to only low common combinations. For the analysis of the combination of different varroa control methods, a usage histogram for each method was generated and grouped into spring, summer, and winter. After that, a vector of all applied methods with at least 15 answers and without drone brood removal was generated to minimize calculation time and to generate statistically relevant results. Mapping of participants main wintering apiary location was done with modified k-means cluster search method [@oberreiter2020].

For the analysis of expenses and operation size, we grouped survey participants by the number of colonies wintered into following groups: 1-20, 21-50 and more than 50. Data distribution analysis was done with QQ-plots and boxplots. The operation factor question analysis (Migratory Beekeeper and Certified Organic Beekeeper) did include as possible answers "Yes", "No", "Uncertain". The category "Uncertain" and participants which did not answers the question were not used in statistical analysis.