# Results
## Survey Data
### Representative and Distribution of Sample

```{r, include=FALSE}
source("src/r-survey.R", local = knitr::knit_global())
```

The survey participants with valid answers including treatment expenses per colony represent between 3.5% and 4.0% of the beekeepers in Austria, when compared to the number of registered beekeeper and honey bee colonies as reported by the national beekeeping association 'Biene Österreich'.[^r-vis] In detail these are in the survey year 2018/19 with `r r_survey$response$rate_beekeeper[[1]]`% (n = `r format(r_survey$response$survey_colonies[[1]], big.mark = ",")`), survey year 2019/20 with `r r_survey$response$rate_beekeeper[[2]]`% (n = `r format(r_survey$response$survey_colonies[[2]], big.mark = ",")`) and 2020/21 with `r r_survey$response$rate_beekeeper[[3]]`% (n = `r format(r_survey$response$survey_colonies[[3]], big.mark = ",")`) (Table \@ref(tab:registered)). The distribution map with participants grouped by main winter apiary location on district level shows that the survey covers most districts in Austria in the three survey years. The distribution of beekeepers between the districts is not the same which could reflect real beekeeper density and/or a possible participation bias, although the size of districts is also not equal (Fig. \@ref(fig:distr-map)).

[^r-vis]: In Austria the registration of beekeepers, apiaries and honeybee colonies in the Consumer Health Information System (Verbrauchergesundheitsinformationssystem - VIS) has been implemented into the national Animal Identification and Registration Regulation 2009 (Tierkennzeichnungs- und Registrierungsverordnung, § 1 TKZVO 2009 idF II 193/2015) from 8th of July 2015.

```{r registered, include=T}
cn <- c("Year", "Total", "Survey", "%", "Total", "Survey", "%")
c <- "Percentage of survey participants in relation to the number of beekeepers in Austria as reported by the national beekeeping association 'Biene Österreich'."
tab <- knitr::kable(
  r_survey$response %>%
    select(year, beekeeper, survey_beekeeper, rate_beekeeper, colonies, survey_colonies, rate_colonies),
  booktabs = TRUE,
  row.names = FALSE,
  col.names = cn,
  caption = c
) %>%
  kableExtra::add_header_above(c("", "Beekeeper [n]" = 3, "Colonies [n]" = 3))
tab
rm(tab, c, cn)
```

```{r distr-map, include=T, fig.cap="Distribution heatmap of the participants main winter apiary on district level. Darker colours are more participants in corresponding district and lighter colours are less. Shapefiles under Creative Commons https://www.data.gv.at/"}
include_custom("output/figs/map.png")
```

The reported expenses per colony arithmetic mean (`r paste(range(r_survey$summary$survey_table$mean), collapse = "-")` Euro) and median (`r paste(range(r_survey$summary$survey_table$median), collapse = "-")` Euro) is similar between the three survey years (Table \@ref(tab:central)). The arithmetic mean is higher than the median, due to the right skewed data distribution with some high values above 100 Euro (Fig. \@ref(fig:distr-year)). The distribution density of the reported expenses are similar between the three survey years and show that participant did answer in round full numbers eg. 5.00, 10.00, 15.00 Euros (Fig. \@ref(fig:distr-year)).


(ref:distr-year) Distribution of the expenses for each survey year, values above 50 are not visible. First graph column represents the actual data points, second column a boxplot with median and interquantile range (IQR, $25^{th}$ - $75^{th}$ quantile) and vertical lines 1.5*IQR. Last column is density distribution of the data points. Different years are colour coded (orange = 2018/19, green = 2019/20, blue = 2020/21).

```{r distr-year, out.width="100%", include=T, messages=F, warnings=F, fig.cap="(ref:distr-year)"}
include_custom("output/figs/distr-year.png")
```

```{r central, include=T}
cn <- c("Year", "Minimum", "Mean", "Geom. Mean", "Median", "Max", "n")
c <- "Survey central tendency values for treatment expenses per colony for all years and additionally over all years. Column 'Mean' is the arithmetic mean, the 'Geom. Mean' (geometrical mean) was added as alternative to reduce the effect of positive outliers."
tab <- knitr::kable(
  dplyr::bind_rows(r_survey$summary$survey_table, r_survey$summary$survey_total) %>%
    select(-type) %>%
    mutate(
      year = ifelse(year == "total", "All Years", paste0("20", year))
    ),
  align = c("l", rep("r", 5)),
  booktabs = TRUE,
  row.names = FALSE,
  col.names = cn,
  # format = "markdown",
  caption = c
) %>%
  kableExtra::row_spec(3, hline_after = TRUE)
tab
rm(tab, c, cn)
```

### Distribution of Participants for Operational Questions {#operational-survey}

As overview for correlation and distribution of the included operational factors of interest, the participants where grouped into bins of combination of these factors (Fig. \@ref(fig:distr-groups)). The two survey questions 'Migratory Beekeeper' and 'Certified Organic Beekeeper' did include as possible answers 'Yes', 'No' and 'Uncertain', whereas 'Uncertain' and participants which did not answer on these questions were ignored. The answer on operation size (number colonies going into winter, any positive integer number was possible) was grouped into the two levels for factor analysis. These levels are 'Recreational Beekeeper' with 25 or less colonies and 'Professional Beekeeper' with more colonies. Based on these three operational questions the most common and dominating group in the survey are recreational beekeepers, who are neither migratory or certified organic with 63-65%. Second most are professional beekeepers with more than 25 colonies with 8-11%. All other combinations of these operational factors are below 10% in each survey year. A small decline in recreational beekeepers which report to migrate their colonies can be seen over the three survey years (2018/19: 7,7%; 2019/20: 6.2%; 2020/21: 5.4%) (Fig. \@ref(fig:distr-groups)).

```{r distr-groups, include=T, messages=F, warnings=F, fig.cap="Count and distribution of participants grouped into operational factors and their combination for each year. The three operational factors and combination bins are below the bar plots, whereas black dots indicate 'Yes'/'True' for given factor. Percentage reported above the bars stands respectively for the survey years 2018/19, 2019/20 and 2020/21. Recreational Beekeepers are participants with 25 or less colonies going into winter.", out.width="100%"}
include_custom("output/figs/distr-groups.png")
```
