# Results

## Survey Data

```{r, include=FALSE}
RAW_REPORTS <- RAW %>% group_by(year) %>% 
  summarize(
    n = n(),
    mean_colonies   = round(mean(hives_winter),2),
    median_colonies = median(hives_winter)
  )

DATA_REPORTS <- DATA %>% group_by(year) %>% 
  summarize(
    n = n(),
    colonies = sum(hives_winter),
    mean_colonies   = round(mean(hives_winter),2),
    median_colonies = median(hives_winter)
  )

REPORTS <- tibble(
  Year    = RAW_REPORTS$year,
  Total   = RAW_REPORTS$n,
  Cost    = DATA_REPORTS$n,
  Percent = round(DATA_REPORTS$n/RAW_REPORTS$n*100,1)
)

NO_COST <- RAW %>% filter(is.na(costs)) %>% group_by(year) %>% summarise(n = n())

stats_aut = stats_aut %>% add_column(
  survey_beekeeper = DATA_REPORTS$n,
  survey_colonies = DATA_REPORTS$colonies
) %>% mutate(
  rate_beekeeper = round(survey_beekeeper / beekeeper * 100, 1),
  rate_colonies = round(survey_colonies / colonies * 100, 1)
)

rm(DATA_REPORTS, RAW_REPORTS)
```

The number of reports which did include costs and removed invalid or illogical answers. The cost answer was left empty by `r NO_COST$n[NO_COST$year == '18/19']` participants in 2018/19 survey and `r NO_COST$n[NO_COST$year == '19/20']` participants in 2019/20. The number of naswers was similar for the survey of 2018/19 with `r REPORTS$Percent[1]`% and for 2019/20 with `r REPORTS$Percent[2]`, see table \@ref(tab:reports). Compared to the number of registered beekeeper and honey bee colonies with the national beekeeping association "Biene Österreich", the number of participation with valid answers including costs for survey year 2018/19 with `r stats_aut$rate_beekeeper[stats_aut$year == '2018']`% and for the survey year 2019/20 with `r stats_aut$rate_beekeeper[stats_aut$year == '2019']`%, see table \@ref(tab:registered). The distribution map of participants with valid answers shows good coverage all over Austria (Figure \@ref(fig:distr-map)). 

```{r reports, include=T}
cn = c("Year", "Total [n]", "Answered [n]", "Percent [%]")
c = "Number of participants answering the question for estimated costs per colony in the survey two survey years."
knitr::kable(
  REPORTS,
  booktabs = TRUE, row.names = FALSE,
  col.names = cn, format = "markdown", caption = c
)
rm(REPORTS, cn, NO_COST)
```

```{r registered, include=T}
cn = c("Year", "Beekeeper [n]", "Colonies [n]", "Survey Beekeeper [n]", "Survey Colonies [n]")
c = "Percentage of survey participants in relation to the amount of registered beekeepers with the national beekeeping association Biene Österreich (2018 and 2019)."
knitr::kable(
  stats_aut[,c(1:5)],
  booktabs = TRUE, row.names = FALSE,
  col.names = cn, format = "markdown", caption = c
)
rm(c, cn)
```

```{r r-map, warning=F, echo=F, include=F}
source("partials/r-map.R", local = knitr::knit_global())
```

```{r distr-map, include=T, fig.cap="Distribution map of approximnate main winter apiary position for our survey participants in the survey year 2018/19 and 2019/20. Shapefiles under Creative Commons https://www.data.gv.at/"}
p
rm(p, MAP1819, MAP1920, MAP_POINTS, MF_DISTRICTS, MF_STATES)
```




## Price Estimations {#r-estimations}