## Expenses Extrapolation

```{r, include=FALSE}
if (file.exists("output/r-extrapolation.rds")) {
    r_extrapolation <- readRDS("output/r-extrapolation.rds")
} else {
    source("src/r-extrapolation.R", local = knitr::knit_global())
    saveRDS(r_extrapolation, "output/r-extrapolation.rds")
}
```

The rough extrapolation of the survey mean expenses with the total sum of honeybee colonies in Austria is based on the reported numbers by the national beekeeping association 'Biene Österreich'[^r-vis].
This extrapolation with the arithmetic mean sums up in 
2018/19 to `r format(r_extrapolation$data %>% pull(extrapolation_estimate)%>% .[1], big.mark = ",")` Euro (95% CI: `r format(r_extrapolation$data %>% pull(extrapolation_estimate_lower)%>% .[1], big.mark = ",")` - `r format(r_extrapolation$data %>% pull(extrapolation_estimate_upper)%>% .[1], big.mark = ",")`), 
2019/20 to `r format(r_extrapolation$data %>% pull(extrapolation_estimate)%>% .[2], big.mark = ",")` Euro (95% CI: `r format(r_extrapolation$data %>% pull(extrapolation_estimate_lower)%>% .[2], big.mark = ",")` - `r format(r_extrapolation$data %>% pull(extrapolation_estimate_upper)%>% .[2], big.mark = ",")`) and 
2020/21 to `r format(r_extrapolation$data %>% pull(extrapolation_estimate)%>% .[3], big.mark = ",")` Euro (95% CI: `r format(r_extrapolation$data %>% pull(extrapolation_estimate_lower)%>% .[3], big.mark = ",")` - `r format(r_extrapolation$data %>% pull(extrapolation_estimate_upper)%>% .[3], big.mark = ",")`) (Table \@ref(tab:extrapolation-tab), Fig. \@ref(fig:extrapolation)).

In addition the same extrapolation was calculated with the geometric mean from the natural log ($\ln$). For this calculation the single zero value in the dataset was replaced with 0.01 Euro. The geometric mean is less influenced by outliers and typically used as point estimate for a log-normal distribution. The results are always lower than the arithmetic mean and the difference in the extrapolation is on average 1,258,377 Euro.
<!-- https://www.graphpad.com/guides/prism/latest/statistics/stat_the_geometric_mean_and_geometr.htm -->
Over the three survey years the extrapolation with the geometric mean sums up in 2018/19 to `r format(r_extrapolation$data %>% pull(extrapolation_estimate)%>% .[4], big.mark = ",")` (95% CI: `r format(r_extrapolation$data %>% pull(extrapolation_estimate_lower) %>% .[4], big.mark = ",")` - `r format(r_extrapolation$data %>% pull(extrapolation_estimate_upper)%>% .[4], big.mark = ",")`), 
2019/20 to `r format(r_extrapolation$data %>% pull(extrapolation_estimate) %>% .[5], big.mark = ",")` (95% CI: `r format(r_extrapolation$data %>% pull(extrapolation_estimate_lower)%>% .[5], big.mark = ",")` - `r format(r_extrapolation$data %>% pull(extrapolation_estimate_upper)%>% .[5], big.mark = ",")`) and 
2020/21 to `r format(r_extrapolation$data %>% pull(extrapolation_estimate) %>% .[6], big.mark = ",")` (95% CI: `r format(r_extrapolation$data %>% pull(extrapolation_estimate_lower)%>% .[6], big.mark = ",")` - `r format(r_extrapolation$data %>% pull(extrapolation_estimate_upper)%>% .[6], big.mark = ",")`) (Table \@ref(tab:extrapolation-tab), Fig. \@ref(fig:extrapolation)).

```{r extrapolation-tab, include=T}
c <- "Multiplication of the average treatment expenses per colony with the total number of honeybee colonies wintered from the survey participants and extrapolation for Austria with the number of colonies as reported from the national beekeeping association 'Biene Österreich'. Confidence intervals for point estimates are calculated with 1,000 bootstrap resamples. Two methods applied arithmetic mean and geometric mean."
cn <- c("Year", "Mean [Euro]", "CI Lower", "CI Higher", "Colonies [n]", "Mean Sum [Euro]", "Colonies [n]", "Mean Sum [Euro]")
kable(
    r_extrapolation$data %>%
        select(year, statistic, conf.low, conf.high, colonies_survey, survey_estimate, colonies, extrapolation_estimate) %>%
        mutate(
            across(statistic:conf.high, round, digits = 2),
            across(colonies_survey:extrapolation_estimate, round)
        ),
    align = c("c", rep("r", 7)),
    booktabs = TRUE,
    row.names = FALSE,
    col.names = cn,
    caption = c,
    escape = TRUE,
) %>%
    add_header_above(c(" " = 4, "Survey" = 2, "Extrapolation" = 2)) %>%
    pack_rows("Arithmetic Mean", 1, 3) %>%
    pack_rows("Geometric Mean", 4, 6) %>%
    kable_styling(latex_options = "scale_down")
rm(c, cn)
```

(ref:extrapolation) Multiplication of the average treatment expenses per colony with the total number of honeybee colonies wintered from the survey participants and extrapolation for Austria with the number of colonies as reported from the national beekeeping association 'Biene Österreich'. Confidence intervals for point estimates are calculated with 1,000 bootstrap resamples. Two methods applied arithmetic mean and geometric mean. Basis of the calculation are the mean expenses for one colony, see Table \@ref(tab:extrapolation-tab).
```{r extrapolation, include=T, fig.cap="(ref:extrapolation)", out.width="100%"}
include_custom("output/figs/extrapolation.png")
```