
## Expenses Estimations {#r-estimations}

```{r, include=FALSE}
if (file.exists("output/r_estimate.rds")) {
    r_estimate <- readRDS("output/r_estimate.rds")
} else {
    source("src/r-estimate.R", local = knitr::knit_global())
    saveRDS(r_estimate, "output/r_estimate.rds")
}
```

The self calculated estimated expenses, based on online research of product prices, central tendency values are similar to the survey responses in each year. The estimate mean is on average EUR&nbsp;`r round(mean(as.numeric(r_estimate$table$mean[1:3]) - as.numeric(r_estimate$table$mean[4:6])), 2)` higher than the survey mean and the estimate median is on average EUR&nbsp;`r round(mean(as.numeric(r_estimate$table$median[1:3]) - as.numeric(r_estimate$table$median[4:6])), 2)` higher than the survey median (Table&nbsp;\@ref(tab:table-estimates)).

```{r table-estimates, include=T}
c <- "Comparison of central tendency values for estimated expenses and the expenses as reported by participants in the survey for each year separately. All values are in Euro."
cn <- c("", "Min.", "Mean", "Median", "Max")

tab <- kable(
    r_estimate$table %>% select(-type),
    align = c("c", rep("r", 4)),
    booktabs = TRUE,
    row.names = FALSE,
    col.names = cn,
    caption = c
) %>%
    pack_rows("Estimate", 1, 3) %>%
    pack_rows("Survey", 4, 6)
tab
rm(c, cn, tab)
```

The comparison between the estimate and reported survey expenses not separated by year but for the different treatment method combinations (excluding drone brood removal) shows that the total overestimation is not true for all treatment methods, as some are also underestimated. This comparison can be visualized as differences plot (Fig.&nbsp;\@ref(fig:estimates-bland)). Typically this is done as a statistical method of agreement between methods [@bland1986]. On average the difference between treatment method combinations with at least 15 participants is close to zero (EUR&nbsp;`r round(r_estimate$ba$mean.diffs, 2)`), indicating in summary no over- or underestimation (Fig.&nbsp;\@ref(fig:estimates-bland), yellow line). Only one combination is outside the 95% confidence interval of the upper and lower two times the standard deviations from the mean difference. This combination of summer and winter oxalic acid trickling of readymade mixture ('SU-Ox-Mix & Wi-Ox-Mix') was underestimated on average by EUR&nbsp;15, although the high difference is mainly driven by extreme outliers in this group (*n*=22, mean expenses survey = EUR&nbsp;19, median expenses survey = EUR&nbsp;9). Overestimation of the estimated expenses are shown for treatment combinations which include biotechnical methods excluding drone brood removal or hypothermia in summer ('SU-Biot') and oxalic acid sublimation in winter ('WI-Ox'). The three most common treatment combinations, formic acid long term in summer ('SU-Fa-LT') and winter oxalic treatment ('WI-Ox-sub', 'WI-Ox-pure', 'WI-Ox-mix'), show similar values for mean estimate and mean reported survey expenses (Fig.&nbsp;\@ref(fig:estimates-bland)). For the top 11 treatment method combinations and the comparison of the mean and median expenses of the estimate and as reported from the participants see Table \@ref(tab:common-comb).

<!--
The values are $\log_2$ converted to minimizes effect of high expense differences between treatment methods (Fig.&nbsp;\@ref(fig:estimates-bland)). The $\log_2$ change can be seen as follows, one represent double of the original and minus one is half of the original on the transformed scale.
Treatment method combinations which include 'formic acid short-term' (SU-FA-ST) are below the lower two times standard deviation limit, which means the survey responses are higher roughly ~1.7 times higher that the calculated estimates.
The $\log_2$ change can be seen as follows, one represent a double of the original and minus one is half of the original on the transformed scale. Each treatment combination is inside the 95% confidence interval of the upper and lower two times standard deviations from the mean difference. Combinations which include formic acid short-term are above the upper two times standard deviation limit, which means the survey responses are higher roughly ~1.7 times higher that the calculated estimates. The two combinations with the higher overestimating and outside of the two times standard deviation limit are both including a biotechnical treatment method in summer.
-->

(ref:estimated-bland) Difference plot between survey and estimated expenses of treatment method combinations with at least 15 beekeepers using it, over the three years of survey. Negative values on the vertical axis are underestimates and positive values are overestimates of the estimated expenses. Lines indicating mean difference (yellow) and lower and upper two times standard deviation (blue) each with 95% confidence interval. Point size is the sample count of each treatment method combination. Combinations inside the two times standard deviations CI range and the top 3 combinations based on number of participants are colour coded and labelled. SP/SU/WI = Spring/Summer/Winter || Fa-LT/ST = Formic acid - long/short term || Ox-sub/pure/mix = Oxalic acid - sublimation, pure or mixture || Biot = Another biotechnical methods.

```{r estimates-bland, include=T, fig.cap="(ref:estimated-bland)", out.width="100%"}
include_custom("output/figs/bland-altman.png")
```

```{r common-comb, include=T}
c <- "Top 11 most used treatment method combination for all three survey years. Total and Percent correspond to participants. Mean and median expenses from the survey and in comparison to the calculated estimate in Euro."
cn <- c("Method", "Total [n]", "Percent [%]", "Mean", "Median", "Mean", "Median")
kable(
    r_estimate$data %>% dplyr::slice_max(n, n = 10) %>% select(c_short_od:med_estimate),
    booktabs = TRUE,
    row.names = FALSE,
    digits = 1,
    col.names = cn,
    caption = c,
    linesep = c("", "", "", "", "\\addlinespace"),
) %>%
    add_header_above(c(" " = 3, "Survey" = 2, "Estimate" = 2)) %>%
    footnote(threeparttable = T, general = "SP/SU/WI = Spring/Summer/Winter || AS-LT/ST = Formic acid - long/short term || Ox-sub/pure/mix = Oxalic acid - sublimation, pure or mixture || Biot = Another biotechnical methods.") %>%
    kable_styling(font_size = 8)
rm(c, cn, d)
```

