## Motivation choosing Varroa control method

```{r, include=FALSE}
source("src/r-motivation.R", local = knitr::knit_global())
```

The question "Which main motivations are most likely to apply to you when choosing the Varroa control method (max. 5)?" was only asked in the survey 2020/21, with a total of `r format(r_motivation$answers, big.mark = ",")` valid answers. Even though it was possible to pick up to 5 categories from a total selection of 21, the most common answer was the single answer of 'Recommendations from beekeeping associations, extension workers, veterinarians, magazines, etc.' (*n* = 16). All other answer combinations in the top 10 had in common that they included 'Efficacy' as one of their answers, with third position only the single answer of 'Efficacy' (*n*=10) (Table \@ref(tab:motivation-comb)). A longer list with the top 25 answers is available in the appendix (Table \@ref(tab:motivation-comb-full)).

The total count of how often one category was picked as motivation shows the dominance of 'Efficacy', which was picked by 67.3% (*n* = 844) of the participants, who did answer on the question. Second most was with 49% (*n* = 615) the category 'Side effects on bees', followed up by 'Residues' (42.9%, n = 538) and 'Infestation Level' (37%, *n* = 464). The answer 'Economic Costs', which is the main topic of this research, was picked by 6.8% (*n* = 80) beekeeper (Fig. \@ref(fig:motivation-count)).

To see if there are geographically differences in the picked categories the top 5 categories where evaluated for the nine federal states in Austria separately. The category 'Efficacy' was still the most common motivation in all states. Starting with second place the relative amount of answers for each category is different between the states. Whereas 'Residues' was second most for Styria and all other states with the category 'Side effects on bees'. Followed by the third most common category 'Residues' for Lower and Upper Austria, Burgenland, Salzburg, Styria and Tyrol, the category 'Legal status' for Vienna and Carinthia and the category 'Colony Infestation Level (Varroa)' in Vorarlberg (Fig. \@ref(fig:motivation-state)). One interesting difference was the placing of the category 'Residues', which was for most states at the third position but for Carinthia at 5$^{th}$ and for Vienna outside of the top 5 at the 7$^{th}$ position (Appendix Table \@ref(tab:motivation-state-rank)).

```{r motivation-comb, include=T}
cn <- c("Motivation", "Answers", "Motivations")
c <- "Top 10 motivation answers combinations for choosing varroa control method given by participants. Participants could select out of a pool of 21 categories and where able to pick up to 5 categories maximum. 'Answers' is the total number this combination was picked and 'Motivations' are the number of different categories in this combination, whereas 1 would indicate a single answer."
tab <- knitr::kable(
    r_motivation$comb_list %>% select(motivation, n, motivations) %>% dplyr::slice_max(n, n = 10),
    align = c("l", rep("r", 2)),
    booktabs = TRUE,
    row.names = FALSE,
    col.names = cn,
    caption = c
) %>%
    kable_styling(latex_options = "scale_down")
tab
rm(tab, c, cn)
```

```{r motivation-count, include=T, messages=F, warnings=F, fig.cap="Total count of all 21 motivation answers for choosing varroa control method. Percentage is the relative amount of participants which did pick this motivation category as one of their answers, number in brackets is the absolute count how many times the answer was picked. Order from most at the top and least at the bottom.", out.width="100%"}
include_custom("output/figs/motivation-count.png")
```

```{r motivation-state, include=T, messages=F, warnings=F, fig.cap="Top five motivation answers for choosing varroa control method. Stacked bar plot representing answers for each federal state in Austria in relative amount to the total answers for each motivation category. Numbers inside stacked bars are the relative number of answers in percent for each state. White background numbers indicate highest and lowest for given motivation category. States are color coded and sorted alphabetically.", out.width="100%"}
include_custom("output/figs/motivation-state.png")
```