# Results
## Survey Data
### Representative and distribution of sample

```{r, include=FALSE}
source("src/r-survey.R", local = knitr::knit_global())
```

The survey participants with valid answers including treatment expenses per colony represent between 3.5% and 4.0% of the beekeepers in Austria, when compared to the number of registered beekeeper and honey bee colonies as reported by the national beekeeping association "Biene Österreich"[^r-vis]. In detail these are in the survey year 2018/19 with `r r_survey$response$rate_beekeeper[[1]]`% (n = `r format(r_survey$response$survey_colonies[[1]], big.mark = ",")`), survey year 2019/20 with `r r_survey$response$rate_beekeeper[[2]]`% (n = (n = `r format(r_survey$response$survey_colonies[[2]], big.mark = ",")`) and 2020/21 with `r r_survey$response$rate_beekeeper[[3]]`% (n = `r format(r_survey$response$survey_colonies[[3]], big.mark = ",")`) (Table \@ref(tab:registered)). The distribution map with participants grouped by main winter apiary location on district level shows that the survey covers most districts in Austria in the three survey years. The distribution is not the same which could reflect real beekeeper density or a possible participation bias, although the size of districts is also not equal (Fig. \@ref(fig:distr-map)).

[^r-vis]: In Austria the registration of beekeepers, apiaries and honeybee colonies in the Consumer Health Information System (Verbrauchergesundheitsinformationssystem - VIS) has been implemented into the national Animal Identification and Registration Regulation 2009 (Tierkennzeichnungs- und Registrierungsverordnung, § 1 TKZVO 2009 idF II 193/2015) from 8th of July 2015.

```{r registered, include=T}
cn <- c("Year", "Total", "Survey", "%", "Total", "Survey", "%")
c <- "Percentage of survey participants in relation to the number of registered beekeepers with the national beekeeping association Biene Österreich."
tab <- knitr::kable(
  r_survey$response %>%
    select(year, beekeeper, survey_beekeeper, rate_beekeeper, colonies, survey_colonies, rate_colonies),
  booktabs = TRUE,
  row.names = FALSE,
  col.names = cn,
  caption = c
) %>%
  kableExtra::add_header_above(c("", "Beekeeper [n]" = 3, "Colonies [n]" = 3))
tab
rm(tab, c, cn)
```

```{r distr-map, include=T, fig.cap="Distribution map of the main winter apiary on district level for our survey participants, which did answer give information about their district. Shapefiles under Creative Commons https://www.data.gv.at/"}
include_custom("output/figs/map.png")
```

The over all survey years the mean equals to `r r_survey$summary$survey_total$mean` euro and the median to `r r_survey$summary$survey_total$median`, see table \@ref(tab:central). The distribution of treatment expenses per colony is right skewed with high values above 100 euro. The years are similar in their distribution, see figure \@ref(fig:distr-year). The distribution shows that participants preferable answer in round estimates eg. 5.00, 10.00, 15.00 Euros.

```{r distr-year, out.width="100%", include=T, messages=F, warnings=F, fig.cap="Disitrubtion of the expenses for each survey year, values above 50 are not visible."}
include_custom("output/figs/distr-year.png")
```

```{r central, include=T}
cn <- c("Year", "Minimum", "Mean", "Geom. Mean", "Median", "Max", "n")
c <- "Survey central tendency values for treatment expenses per colony for all years and additionally over all years. Geometrical mean (Geom. Mean) added to reduce effect of outliers for the mean."
tab <- knitr::kable(
  dplyr::bind_rows(r_survey$summary$survey_table, r_survey$summary$survey_total) %>%
    select(-type) %>%
    mutate(
      year = ifelse(year == "total", "All Years", paste0("20", year))
    ),
  align = c("l", rep("r", 5)),
  booktabs = TRUE,
  row.names = FALSE,
  col.names = cn,
  # format = "markdown",
  caption = c
) %>%
  kableExtra::row_spec(3, hline_after = TRUE)
tab
rm(tab, c, cn)
```

### Distribution of Participants in groups of interest

The most common group in the survey are recreational beekeepers, which I defined as participants with 25 or less colonies (figure \@ref(fig:distr-groups)). Second most are beekeepers with more than 25 colonies. Only a small portion of participants are certified organic or migratory beekeepers and even less both combined. The observed distribution is similar over all three survey years, a small decline in recreational beekeepers which are also report to migrate their colonies can be seen.

```{r distr-groups, include=T, messages=F, warnings=F, fig.cap="Count and distribution of participants in groups of interest for each year. Percentage reported stands respectivley for the survey years 2018/19, 2019/20 and 2020/21. Recreational Beekeepers are participants with 25 or less colonies going into winter.", out.width="100%"}
include_custom("output/figs/distr-groups.png")
```
