## Treatment Network {#network}

```{r}
source("src/r-treatment-network.R", local = knitr::knit_global())
```

To better understand which treatment methods are actually commonly reported a adjacency matrix of all treatment methods with at least 30 participants reporting it over all three survey years was generated. The follow up network analysis of the treatment methods over all three survey years showed the dominance of oxalic treatment methods in winter. In spring the biomechinal method drone brood removal (*n*=1,833) is the most dominant reported varroa treatment and very often in combination with the summer variant (*n*=1,638, combination=1,138). In summer the formic acid treatment variants are the most common (long=2,004, short=1,639) followed by the oxalic acid treatments (sublimation=994, pure trickling or spraying=582, ready-made mixture trickling=560). The treatment method 'Another biotechnical Method' (without drone brood removal or hyperthermia) is in summer also frequently reported with *n*=950. The other treatment methods in summer with lower number of participants are 'Thymol' (*n*=328), 'Hyperthermia' (*n*=160), 'Lactic acid' (*n*=92) and 'Synthetic Methods' (*n*=86) (Fig.&nbsp;\@ref(fig:treatment-network)).

(ref:treatment-network) Network graph of the treatment methods over all three survey years. Node sizes are the number the method were reported. Node colours are the seasons 'Spring' (April-May; Green), 'Summer' (June-October; Black) and 'Winter' (November-January; Blue). Edge connections colour and width are the reported amount of each combination.  Abbrevations: Drone = Drone brood removal || AS-LT/ST = Formic acid - long/short term || Ox-sub/pure/mix = Oxalic acid - sublimation, pure or mixture || Biot = Another biotechnical methods || Thy = Thymol || Hyp = Hyperthermia || syn Met = synthetic Methods. Categories with less than 30 reported answers were not included.

```{r treatment-network, include=T, fig.cap="(ref:treatment-network)"}
include_custom("output/figs/treatment-network.png")
```