
## Expenses Estimations {#r-estimations}

```{r, include=FALSE}
source("src/r-estimate.R", local = knitr::knit_global())
```

The calculated estimated expenses mean and median are higher but similar in each year, than in the survey given by participants, see table \@ref(tab:table-estimates). The most common combinations, without drone brood removal, over the survey years and the corresponding estimates can be seen in table \@ref(tab:common-comb). For the tables the estimated expenses are calculated for each beekeeper in the survey and then the average was taken as with the survey expenses.

```{r table-estimates, include=T}
c <- "Comparison of estimated expenses and the expenses given by participants in the survey. All values are in Euro."
cn <- c("", "Min.", "Mean", "Median", "Max")
tab <- kable(
    r_estimate$table %>% select(-type),
    align = c("c", rep("r", 4)),
    booktabs = TRUE,
    row.names = FALSE,
    col.names = cn,
    caption = c
) %>%
    pack_rows("Estimate", 1, 3) %>%
    pack_rows("Survey", 4, 6)
tab
rm(c, cn, tab)
```

```{r common-comb, include=T}
c <- paste0("Top 10 most used treatment combination for all three survey years. Total and Percent correspond to participants. Mean and median expenses from the survey and in comparison to the estimate in Euro. Drone brood removal method is excluded from the combinations.")
cn <- c("Method", "Total [n]", "Percent [%]", "Mean", "Median", "Mean", "Median")
kable(
    r_estimate$data %>% dplyr::slice_max(n, n = 10) %>% select(c_short_od:med_estimate),
    booktabs = TRUE,
    row.names = FALSE,
    digits = 1,
    col.names = cn,
    caption = c,
) %>%
    add_header_above(c(" " = 3, "Survey" = 2, "Estimate" = 2)) %>%
    footnote(threeparttable = T, general = "SP/SU/WI = Spring/Summer/Winter || AS-LT/ST = Formic acid - long/short term || Ox-sub/pure/mix = Oxalic acid - sublimation, pure or mixture || Biot = Another biotechnical methods")
rm(c, cn, d)
```

Most of our estimates for common treatment with at least 10 participants using the combination of methods are overestimated. This can be seen if we plot the mean differences, of survey expenses versus estimated expenses as statistical method of agreement [@bland1986]. $\log_2$ converted to minimizes effect of high expense differences between treatment methods, see figure \@ref(fig:estimates-bland).

The $\log_2$ change can be seen as follows, one represent a double of the original and minus one is half of the original on the transformed scale. Each treatment combination is inside the 95% confidence interval of the upper and lower two times standard deviations from the mean difference. Combinations which include formic acid short-term are above the upper two times standard deviation limit, which means the survey responses are higher roughly ~1.7 times higher that the calculated estimates. Overall there are more overestimates of calculated estimates and the results coming from the survey are lower on average lower, which is indicate by the negative mean difference (figure \@ref(fig:estimates-bland), yellow line). The two combinations with the higher overestimating and outside of the two times standard deviation limit are both including a biotechnical treatment method in summer.

(ref:estimated-bland) $\log_2$ transformed mean difference between survey and estimated expenses of the treatment combinations with at least 10 beekeepers using it, over the three years of survey. Lines indicating mean and lower and upper two times standard deviations each with 95% confidence interval. Point size is the frequency count of each treatment method. Treatment combination above the two times standard deviations are colour coded and labeled, the label meaning is explained in \@ref(tab:common-comb).

```{r estimates-bland, include=T, fig.cap="(ref:estimated-bland)"}
include_custom("output/figs/bland-altman.png")
```
