## Operational factor analysis
```{r, include=FALSE}
if (file.exists("output/r-operational.rds")) {
    r_operational <- readRDS("output/r-operational.rds")
} else {
    source("src/r-operational.R", local = knitr::knit_global())
}
```

The three selected operational factors of interest are shown in this analysis as single factor, although the groups interact and cohort sizes are different [Section \@ref(operational-survey)]. The factors include the two survey questions 'Migratory Beekeeper' and 'Certified Organic Beekeeper' did include as possible answers 'Yes', 'No' and 'Uncertain', whereas 'Uncertain' and participants which did not answer on these questions were ignored. 

### Operation size

The number of honeybee colonies managed by participants follow a heavy tailed count distribution (Mean = 21-23 Colonies, Median = 11 Colonies), with most participants managing less than 20 colonies and a few big operation with over one hundred colonies (Fig. \@ref(fig:size-distr)). The participants were grouped into two levels. Beekeeper with 25 or less colonies (<=25) as 'Recreational Beekeeper'. Over the survey years the 'Recreational Beekeeper' made up the majority of the survey responses with 78% in 2018/19, 80% in 2019/20 and 76% in 2020/21. The second level which is the minority are the 'Professional Beekeeper' with more than 25 colonies (>25) (Table \@ref(tab:size-table)). 

Comparing these two levels and the reported expenses per colony on treatment, shows that the 'Recreational Beekeeper' spend on average and in median more per colony as the 'Professional Beekeeper' with more than 25 colonies. The absolute median difference between the groups is between `r min(r_operational$size$statistic_perm$point_estimate %>% unlist())` and `r max(r_operational$size$statistic_perm$point_estimate %>% unlist())` Euro per colony. Inference based statistic based on the median difference with permutation test (R = 5,000) shows significant difference in all years (*p* < 0.01) and 95% CI (bootstrap, R = 1,000; 2018/19: 4.1-5.9; 2019/20: 2.6-5.4; 2020/21: 2.6-5.4 Euro) is not overlapping with zero difference in any survey year (Fig. \@ref(fig:size-stats)).

(ref:size-distr) Distribution of operation size / number of colonies over the three survey years. Numbers are based on the reported colonies going into winter, as reported by survey participants. Different years are colour coded (orange = 2018/19, green = 2019/20, blue = 2020/21). Total number of participants and colonies are reported in each years plot header.

```{r size-distr, include=T, fig.cap="(ref:size-distr)", out.width="100%"}
include_custom("output/figs/size-distr.png")
```

```{r size-table, include=T}
c <- "Operation size / number of colonies distribution and measures of central tendency over the three survey years. Operation size with 25 or less colonies ('<=25') are 'Recreational Beekeeper' and with more are 'Professional Beekeeper' ('>25')."
cn <- c("Operation Size", "Beekeeper [n]", "Colonies [n]", "Mean [Euro]", "Median [Euro]")
tab <- kable(
    r_operational$size$table %>% select(-year),
    booktabs = TRUE,
    col.names = cn,
    digits = 1,
    align = c("c", rep("r", 4)),
    caption = c
) %>%
    pack_rows("2018/19", 1, 2) %>%
    pack_rows("2019/20", 3, 4) %>%
    pack_rows("2020/21", 5, 6) %>%
    kable_styling(latex_options = "scale_down")
tab
rm(cn, c, tab)
```

(ref:size-stats) Boxplot with median and interquantile range (IQR, $25^{th}$ - $75^{th}$ quantile) and vertical lines indicate 1.5\*IQR. Pink rhombus represents the arithmetic mean. Pairwise difference of the mean between groups are given as labels above the boxplots. Gray points are the actual data points. Reported inference statistic results on median difference between the given operation size groups, with *p*-values from permutation test (R = 5,000) and 95% CI (bootstrap, R = 1,000). Exact *p*-values and test results are available in the Appendix Fig. \@ref(fig:size-permutation). Survey answers above 80 euros per colony are not shown.  

```{r size-stats, include=T, fig.cap="(ref:size-stats)", out.width="100%"}
include_custom("output/figs/size-stats.png")
```

### Certified organic beekeeping operation

Only `r r_operational$organic$summary$np[[2]]`% of the survey participants are certified organic, see table \@ref(tab:organic-table). Participants who are a certified organic operation have signifikant lower treatment expenses per colony on average and median than participants without certificate \@ref(fig:organic-stats). The absolute median difference between the groups is between `r min(r_operational$organic$statistic_perm$point_estimate %>% unlist())` and `r max(r_operational$organic$statistic_perm$point_estimate %>% unlist())` Euro per colony.

```{r organic-table, include=T}
c <- "Certified organic beekeeping operation distribution and measures of central tendency over the three survey years."
cn <- c("Certified Organic", "Beekeeper [n]", "Colonies [n]", "Mean [Euro]", "Median [Euro]")
tab <- kable(
    r_operational$organic$table %>% select(-year),
    booktabs = TRUE,
    col.names = cn,
    digits = 1,
    align = c("c", rep("r", 4)),
    caption = c
) %>%
    pack_rows("2018/19", 1, 2) %>%
    pack_rows("2019/20", 3, 4) %>%
    pack_rows("2020/21", 5, 6) %>%
    kable_styling(latex_options = "scale_down")
tab
rm(cn, c, tab)
```

(ref:organic-stats) Boxplot and test results of expenses in if survey participant is certified organic or not, with median difference based on permutation test and bootstrap confidence interval of the difference. Exact *p*-value and test results see fig. \@ref(fig:size-permutation). Gray points indicating individual datapoints, survey answers above 80 euros per colony are not shown. Pink rhombus indicating mean of sample group. Pairwise comparison of mean between groups are given as label above the boxplots.

```{r organic-stats, include=T, fig.cap="(ref:organic-stats)", out.width="100%"}
include_custom("output/figs/organic-stats.png")
```

### Migratory beekeeper

The beekeepers participating in our study were asked whether they transported their bees to sources of honey yields or for pollination service (this excludes movements in the course of breeding or splits). A total of `r r_operational$migratory$summary$np[[2]]`% of the survey participants are migrating with their colonies, see table \@ref(tab:migratory-table). Participants who migrate their colonies have signifikant lower treatment expenses per colony on average and median than non-migratory beekeepers \@ref(fig:migratory-stats). The absolute median difference between the groups is between `r min(r_operational$migratory$statistic_perm$point_estimate %>% unlist())` and `r max(r_operational$migratory$statistic_perm$point_estimate %>% unlist())` Euro per colony.

```{r migratory-table, include=T}
c <- "Migratory beekeeper distribution and measures of central tendency over the three survey years."
cn <- c("Migratory Beekeeper", "Beekeeper [n]", "Colonies [n]", "Mean [Euro]", "Median [Euro]")
tab <- kable(
    r_operational$migratory$table %>% select(-year),
    booktabs = TRUE,
    col.names = cn,
    digits = 1,
    align = c("c", rep("r", 4)),
    caption = c
) %>%
    pack_rows("2018/19", 1, 2) %>%
    pack_rows("2019/20", 3, 4) %>%
    pack_rows("2020/21", 5, 6) %>%
    kable_styling(latex_options = "scale_down")
tab
rm(cn, c, tab)
```

(ref:migratory-stats) Boxplot and test results of expenses in if migratory beekeeping operation or not, with median difference based on permutation test and bootstrap confidence interval of the difference. Exact *p*-value and test results see fig. \@ref(fig:size-permutation). Gray points indicating individual datapoints, survey answers above 80 euros per colony are not shown. Pink rhombus indicating mean of sample group. Pairwise comparison of mean between groups are given as label above the boxplots.

```{r migratory-stats, include=T, fig.cap="(ref:migratory-stats)", out.width="100%"}
include_custom("output/figs/migratory-stats.png")
```
