## Treatment Network {#network}

```{r}
# Note the graph count is lower than the heatmap as for the graph the season entry is only counted once eg. drone brood removal in spring 1 even though participants answered for march and april with 1
source("src/r-treatment-network.R", local = knitr::knit_global())
```

The application of treatment methods follows a seasonal pattern with 'Drone brood removal' in spring, 'Formic acid' mainly in summer and 'Oxalic acid' in winter (Fig.&nbsp;\@ref(fig:treatment-heat)). To get an overview on which treatment methods are commonly reported together a adjacency matrix of all methods grouped into the seasons with at least 30 participants reporting it over all three survey years (*n*=`r nrow(dfClean) %>% ft()`) was generated. The follow up network analysis of the treatment methods over all three survey years showed the mentioned dominance of oxalic treatment methods in winter. In spring the biomechanical method drone brood removal (52.8%) is the most dominant reported varroa treatment and very often in combination with the summer variant (47.2%, combination=32.8%). In summer the formic acid treatment variants are the most common (long=57.8%, short=47.2%) followed by the oxalic acid treatments (sublimation=28.7%, pure trickling or spraying=16.8%, ready-made mixture trickling=16.1%). The treatment method 'Another biotechnical Method' (without drone brood removal or hyperthermia) is in summer also frequently reported with 27.4%. The other treatment methods in summer with lower number of participants are 'Thymol' (9.5%), 'Hyperthermia' (4.6%), 'Lactic acid' (2.7%) and 'Synthetic Methods' (2.5%) (Fig.&nbsp;\@ref(fig:treatment-network)).

(ref:treatment-network) Network graph of the treatment methods over all three survey years (*n*=`r nrow(dfClean) %>% ft()`). Node sizes are the number the method were reported. Node colours are the seasons 'Spring' (April-May; Green), 'Summer' (June-October; Black) and 'Winter' (November-March; Blue). Edge connections colour and width are the reported amount of each combination.  Abbrevations: Drone = Drone brood removal || Fa-LT/ST = Formic acid - long/short term || Ox-sub/pure/mix = Oxalic acid - sublimation, pure or mixture || Biot = Another biotechnical methods || Thy = Thymol || Hyp = Hyperthermia || syn Met = synthetic Methods. Categories with less than 30 reported answers were not included.

(ref:treatment-heat) Absolute numbers of reported treatments in each month over the three survey years (*n*=`r nrow(dfClean) %>% ft()`). Colour coded as heatmap with darker colours representing more answers and no answers in white.

```{r treatment-heat, include=T, fig.cap="(ref:treatment-heat)", out.width="70%"}
include_custom("output/figs/treatment-heat.png")
```

```{r treatment-network, include=T, fig.cap="(ref:treatment-network)", out.width="70%"}
include_custom("output/figs/treatment-network.png")
```



