## Motivation choosing *Varroa* Control Method

```{r, include=FALSE}
source("src/r-motivation.R", local = knitr::knit_global())
```

The question "What are your main motivations in choosing your *Varroa* control method (max. 5)?" was only asked in the survey 2020/21, with a total of `r format(r_motivation$answers, big.mark = ",")` valid answers. Even though it was possible to pick up to five categories from 21 available options, the most common answer was the single answer of 'Recommendations from beekeeping associations, extension workers, veterinarians, magazines, etc.' (16 participants). All other answer combinations in the top 10 had in common that they included 'Efficacy' as one of their answers, or even as single answer (10 participants) (Table&nbsp;\@ref(tab:motivation-comb)). A more extensive list with the top 25 answers is available in the appendix (Table&nbsp;\@ref(tab:motivation-comb-full)).

The total count of how often one category was picked as motivation (*n*=`r format(r_motivation$answers, big.mark = ",")`) shows the dominance of 'Efficacy', which was picked by 67.3% of the participants. Second most was the category 'Side effects on bees' with 49%, followed up by 'Residues' (42.9%) and 'Infestation Level' (37.0%). The answer 'Economic Costs', which reflects is one of the key topics of this research, was picked by 6.8% beekeepers (Fig.&nbsp;\@ref(fig:motivation-count)).

```{r motivation-comb, include=T}
cn <- c("Motivation", "Answers", "Motivations")
c <- "Top 10 answer combinations for the motivation of choosing a certain \\textit{Varroa} control method given by participants (\\textit{n}=1,255). Participants could select out of a pool of 21 categories and where able to pick up to 5 categories maximum. 'Answers' is the total number this combination was picked and 'Motivations' represents the number of different categories in this combination, whereas 1 would indicate a single answer."
tab <- knitr::kable(
    r_motivation$comb_list %>% select(motivation, n, motivations) %>% dplyr::slice_max(n, n = 10),
    align = c("l", rep("r", 2)),
    booktabs = TRUE,
    row.names = FALSE,
    col.names = cn,
    caption = c
) %>%
    kable_styling(latex_options = "scale_down")
tab
rm(tab, c, cn)
```

(ref:motivation-count) Total count of all 21 motivation answers for choosing a *Varroa* control method. Percentage is the relative amount of participants (*n*=`r format(r_motivation$answers, big.mark = ",")`) which did pick this motivation category as one of their answers, number in brackets is the absolute count how many times the answer was picked. Order from most at the top to least at the bottom.

```{r motivation-count, include=T, messages=F, warnings=F, fig.cap="(ref:motivation-count)", out.width="100%"}
include_custom("output/figs/motivation-count.png")
```

To see if there are geographical differences in the picked categories, the top five categories where evaluated for the nine federal states in Austria separately. The category 'Efficacy' was again the most common motivation across all states. Starting with second place the relative number of answers for each category differs between the states. 'Residues' was second most picked for Styria and 'Side effects on bees' for the others states. The third most common category was 'Residues' for Lower and Upper Austria, Burgenland, Salzburg, Styria and Tyrol, the category 'Legal status' for Vienna and Carinthia and the category 'Colony Infestation Level (*Varroa*)' in Vorarlberg (Fig.&nbsp;\@ref(fig:motivation-state)). One interesting difference was the placing of the category 'Residues', which for most states took third position but  5$^{th}$ for Carinthia and 7$^{th}$ for Vienna, outside of the top five (Appendix Table \@ref(tab:motivation-state-rank)).

(ref:motivation-state) Top five motivation answers for choosing *Varroa* control method. Stacked bar plot representing answers for each federal state in Austria in relative amount to the total answers for each motivation category. Numbers inside stacked bars are the relative number of answers in percent for each state. White background numbers indicate highest and lowest for given motivation category. States are colour coded and sorted alphabetically.

```{r motivation-state, include=T, messages=F, warnings=F, fig.cap="(ref:motivation-state)", out.width="100%"}
include_custom("output/figs/motivation-state.png")
```

As motivation could also be driven by the operation size, the results were split into beekeepers with 25 or less colonies (<=25) representing 'Recreational Beekeepers' and 'Professional Beekeepers' with more than 25 colonies (>25) (Fig. \@ref(fig:motivation-operation)). 'Efficacy' was the most picked category for both groups (67%/68%). The category 'Residues' was more often picked by 'Professional Beekeepers' (58%/39%), whereas 'Side effects on bees' was the second most common for 'Recreational Beekeepers' (50%/44%). The categories 'Time and number of applications' (25%/34%) and 'Resistances' (14%/28%) showed a 1.5 - 2x higher rate for 'Professional Beekeepers'. For 'Recreational Beekeepers' the categories 'Recommendations from beekeeping associations, extension workers, veterinarians, magazines, etc.' (31%/10%) and 'Level of difficulty and needed equipment' (23%/13%) were more important and more often chosen in comparison to the bigger operations. The category 'Economic costs' (6%/9%) showed a higher relative count for 'Professional Beekeepers' (Fig. \@ref(fig:motivation-operation)).

(ref:motivation-operation) Relative count of all 21 answers regarding the motivation for the selection of a certain *Varroa* control method separated into operation sizes. Operation size with 25 or less colonies ('<=25') are denoted as 'Recreational Beekeeper' (blue) and with more are subsumed as 'Professional Beekeeper' ('>25') (green). The percentage is the relative amount of participants which did pick this motivation category as one of their answers, number in brackets is the absolute count how many times the answer was picked. The data is ordered by the sum of both percentages from most at the top and least at the bottom.

```{r motivation-operation, include=T, messages=F, warnings=F, fig.cap="(ref:motivation-operation)", out.width="100%"}
include_custom("output/figs/motivation-operation.png")
```
