## Operational factor analysis
```{r, include=FALSE}
source("src/r-operational.R", local = knitr::knit_global())
```

Most questions in the survey are related to winter colony losses but two operational questions could give more insight into the expenses spend by beekeepers for their treatment method. This two are if the beekeeper migrates with his colonies and if it is an organic certified beekeeping operation, with possible answers "Yes", "No". If survey participants did answer "Unsure" or did not answer at all it was not used for the factor analysis. Additionally I look at the operation size, which can also be seen as operational questions.

### Operation size

Operation sizes and number of participants in our survey are not normally distributed, with most participants managing less than 20 colonies and a few big operation with over one hundred colonies, see figure \@ref(fig:size-distr). Further on the participants were grouped into beekeepers with 25 or less colonies, which should represent recreational beekeepers, and the ones with more than 25 colonies, which should represent side-line and professional operations (table \@ref(tab:size-table)).

(ref:size-distr) Distribution of operation size over the three survey years. Colonie numbers based on the reported colonies going into winter.

```{r size-distr, include=T, fig.cap="(ref:size-distr)"}
include_custom("output/figs/size-distr.png")
```

```{r size-table, include=T}
c <- "Operation size distribution and measures of central tendency over the three survey years."
cn <- c("Operation Size", "Beekeeper [n]", "Colonies [n]", "Mean [Euro]", "Median [Euro]")
tab <- kable(
    r_operational$size$table %>% select(-year),
    booktabs = TRUE,
    col.names = cn,
    digits = 1,
    align = c("c", rep("r", 4)),
    caption = c
) %>%
    pack_rows("2018/19", 1, 2) %>%
    pack_rows("2019/20", 3, 4) %>%
    pack_rows("2020/21", 5, 6)
tab
rm(cn, c, tab)
```

The given operation size groups show statistical significant difference in all survey years, see figure \@ref(fig:size-stats).

(ref:size-stats) Boxplot and test results of expenses in the two given operation size groups, with median difference based on permutation test and bootstrap confidence interval of the difference. Exact *p*-value and test results see fig. \@ref(fig:size-permutation). Gray points indicating individual datapoints, survey answers above 80 euros per colony are not shown. Pink rhombus indicating mean of sample group. Pairwise comparison of mean between groups are given as label above the boxplots.

```{r size-stats, include=T, fig.cap="(ref:size-stats)"}
include_custom("output/figs/size-stats.png")
```

### Certified organic beekeeping operation

Only `r r_operational$organic$summary$np[[2]]`% of the survey participants are certified organic, see table \@ref(tab:organic-table). There is a statistical significant difference between organic and non-organic beekeepers in all survey years, see figure \@ref(fig:organic-stats).

```{r organic-table, include=T}
c <- "Certified organic beekeeping operation distribution and measures of central tendency over the three survey years."
cn <- c("Certified Organic", "Beekeeper [n]", "Colonies [n]", "Mean [Euro]", "Median [Euro]")
tab <- kable(
    r_operational$organic$table %>% select(-year),
    booktabs = TRUE,
    col.names = cn,
    digits = 1,
    align = c("c", rep("r", 4)),
    caption = c
) %>%
    pack_rows("2018/19", 1, 2) %>%
    pack_rows("2019/20", 3, 4) %>%
    pack_rows("2020/21", 5, 6)
tab
rm(cn, c, tab)
```

(ref:organic-stats) Boxplot and test results of expenses if the beekeeping operation is certified organic or not, with approximative Kruskal-Wallis test results and effect size. Gray points indicating individual datapoints, survey answers above 80 euros per colony are not shown. Pink rhombus indicating mean of sample group. Pairwise comparison of median between groups are given as label above the boxplots.

```{r organic-stats, include=T, fig.cap="(ref:organic-stats)"}
include_custom("output/figs/organic-stats.png")
```

### Migratory beekeeper

The beekeepers participating in our study were asked whether they transported their bees to sources of honey yields or for pollination service (this excludes movements in the course of breeding or splits). A total of `r r_operational$migratory$summary$np[[2]]`% of the survey participants are migrating with their colonies, see table \@ref(tab:migratory-table). There is a statistical significant difference between migratory beekeeper and non-migrating beekeepers in all survey years, see figure \@ref(fig:migratory-stats).

```{r migratory-table, include=T}
c <- "Migratory beekeeper distribution and measures of central tendency over the three survey years."
cn <- c("Migratory Beekeeper", "Beekeeper [n]", "Colonies [n]", "Mean [Euro]", "Median [Euro]")
tab <- kable(
    r_operational$organic$table %>% select(-year),
    booktabs = TRUE,
    col.names = cn,
    digits = 1,
    align = c("c", rep("r", 4)),
    caption = c
) %>%
    pack_rows("2018/19", 1, 2) %>%
    pack_rows("2019/20", 3, 4) %>%
    pack_rows("2020/21", 5, 6)
tab
rm(cn, c, tab)
```

(ref:migratory-stats) Boxplot and test results of expenses if migratory beekeeper or not, with approximative Kruskal-Wallis test results and effect size. Gray points indicating individual datapoints, survey answers above 80 euros per colony are not shown. Pink rhombus indicating mean of sample group. Pairwise comparison of median between groups are given as label above the boxplots.

```{r migratory-stats, include=T, fig.cap="(ref:migratory-stats)"}
include_custom("output/figs/migratory-stats.png")
```
