\newpage
## Efficacy & Economical
```{r, include=FALSE}
if (file.exists("output/r-treatment.rds")) {
    r_treatment <- readRDS("output/r-treatment.rds")
} else {
    source("src/r-treatment.R", local = knitr::knit_global())
    saveRDS(r_treatment, "output/r-treatment.rds")
}
```

The total observed loss rate over the three survey years was in 2018/19 `r round(r_treatment$year_loss$year_loss[1], 1)`% (95% CI: `r round(r_treatment$year_loss$loss_lower_ci[1], 1)`-`r round(r_treatment$year_loss$loss_upper_ci[1], 1)`%), in 2019/20 `r round(r_treatment$year_loss$year_loss[2], 1)`% (95% CI: `r round(r_treatment$year_loss$loss_lower_ci[2], 1)`-`r round(r_treatment$year_loss$loss_upper_ci[2], 1)`%) and in 2020/21 `r round(r_treatment$year_loss$year_loss[3], 1)`% (95% CI: `r round(r_treatment$year_loss$loss_lower_ci[3], 1)`-`r round(r_treatment$year_loss$loss_upper_ci[3], 1)`%) (Fig.&nbsp;\@ref(fig:efficient-economic), blue dashed vertical lines). Over the three survey years there were 17 different treatment method combinations (without drone brood removal) which had in sum at least 30 participants employing it (Table&nbsp;\@ref(tab:ee-table)). The most common treatment combination is long term formic acid in summer followed by oxalic acid trickling in winter (**A** 'SU-Fa-LT & WI-Ox-pure'). It showed in each year a mean low loss rate compared to the total loss rate but only in one year a low expenses category. In comparison the application of short term formic acid treatment in summer and oxalic acid sublimation in summer and winter (**K** 'SU-Fa-ST & SU/WI-Ox-Sub') was observed with a low loss rate category in each year and low price category in two out of three years. A total of three combinations showed in all three years in mean a higher loss rate than the yearly total loss rate, which are formic acid short term in summer followed by oxalic acid trickling pure or ready made mixture in winter (**E** 'SU-Fa-ST & WI-Ox-pure'; **L** 'SU-Fa-ST & WI-Ox-mix') and the unique application of formic acid long term method in summer, without winter treatment, which also showed high expenses in each year (**M** 'SU-Fa-LT'). The three combinations with biotechnical method (without drone brood removal or hyperthermia) in summer showed low expenses in each year (**G** 'SU-Biot. & SU-Ox-sub & WI-Ox-sub'; **P** 'SU-Biot. & SU-Fa-LT & WI-Ox-pure'; **Q** 'SU-Biot. & SU-Fa-ST & WI-Ox-pure'). Additionally the treatment method of oxalic acid sublimation in summer and winter resulted in low mean expenses each year (**F** 'SU-Ox-sub & WI-Ox-sub') (Fig.&nbsp;\@ref(fig:efficient-economic)).

(ref:efficient-economic) Efficiency and economical plot for treatment method combinations with at least *n* = 30 as sum over the three survey years. Letters corresponding to different treatment methods, see Table \@ref(tab:ee-table). Vertical axis the mean expenses for each treatment combination with standard error (SE). Horizontal axis the mean observed winter colony loss rates for each treatment combination with 95% confidence interval (CI). The blue dashed line indicating means total expenses (horizontal) and means total observed winter colony loss rates (vertical) per year. Categorizing of treatment combinations based on location in relation to the blue dashed lines (Fig.&nbsp;\@ref(fig:efficient-economic-mm)). The four categories are colour coded as seen in the plot legend.

```{r efficient-economic, include=T, fig.cap="(ref:efficient-economic)"}
include_custom("output/figs/efficient-economic.png")
```

(ref:ee-table) Treatment combinations with at least *n*=30 as sum over the three survey years. Ordered by number of participants applying the combinations. 'Efficient' and 'Economical' are colour coded based on low (teal) and high (orange), categorized based to total mean loss rates and total mean expenses (Fig.&nbsp;\@ref(fig:efficient-economic)). Abbrevations: SP/SU/WI = Spring/Summer/Winter || AS-LT/ST = Formic acid - long/short term || Ox-sub/pure/mix = Oxalic acid - sublimation, pure or mixture || Biot = Another biotechnical methods.

```{r ee-table, include=T}
cn <- c(" ", "Treatment", "Year", "\\textit{n}", "Median", "Mean", "SD", "Mean", "Lower CI", "Upper CI", "Efficient", "Economical")
cell_colors_price <- c("high-expense" = "cred!80", "low-expense" = "cgreen!80")
cell_colors_loss <- c("high-loss" = "cred!80", "low-loss" = "cgreen!80")

tab <- kbl(
    r_treatment$tab %>%
        mutate(
            c_short_od = str_replace_all(c_short_od, "&", " \\\\& "),
            across(letter:c_short_od, ~ ifelse(lag(as.character(.x), default = "0") == .x, "", .x)),
            across(observed_costs:observed_costs_sd, ~ ifelse(lag(.x, default = 0) == .x, "", round(.x, 2)))
        ) %>%
        mutate(
            letter = cell_spec(letter, bold = TRUE),
            efficient = cell_spec(
                efficient,
                background = cell_colors_loss[efficient]
            ),
            economical = cell_spec(
                economical,
                background = cell_colors_price[economical]
            )
        ),
    linesep = c("", "", "\\addlinespace"),
    booktabs = TRUE,
    col.names = cn,
    digits = 1,
    align = c("c", "l", "c", rep("r", 6), "c", "c"),
    caption = "(ref:ee-table)",
    escape = FALSE
) %>%
    add_header_above(c(" " = 4, "Expenses [EUR]" = 3, "Observed Loss Rate [\\%]" = 3, " " = 2)) %>%
    # landscape()
    kable_styling(latex_option = "scale_down")
tab
rm(cn, c, tab)
```
