# (APPENDIX) Appendix {-}
# Appendix

(ref:extrapolation-estimate) Multiplication of the mean estimated treatment expenses per colony with the total number of honey bee colonies wintered from the survey participants and extrapolation for Austria with the number of colonies as reported from the national beekeeping association 'Biene Österreich'. Confidence intervals for point estimates are calculated with 1,000 bootstrap resamples. Two methods applied arithmetic mean and geometric mean. Basis of the calculation are the estimated mean expenses for one colony..

```{r extrapolation-estimate, out.width="100%", include=T, fig.cap="(ref:extrapolation-estimate)"}
include_custom("output/figs/extrapolation-estimate.png")
```

(ref:size-permutation) Median difference null distribution (gray bars) based on 5,000 permutations. Red line is the point estimate for the median difference with yellow area as the 95% confidence interval. Red shadow indicating the calculated two-sided *p*-value range on the null distribution. 

(ref:size-permutation2) The *p*-values reported as "\*\*\*" would need more permutation to calculate a precise *p*-value.

```{r size-permutation, out.width="100%", include=T, fig.cap="Two level factor analysis statistical test results for 'Operation Size'. (ref:size-permutation) (ref:size-permutation2)"}
include_custom("output/figs/size-permutation.png")
```

```{r organic-permutation, out.width="100%", include=T, fig.cap="Two level factor analysis statistical test results for 'Certified Organic Beekeeper'. (ref:size-permutation)"}
include_custom("output/figs/organic-permutation.png")
```

```{r migratory-permutation, out.width="100%", include=T, fig.cap="Two level factor analysis statistical test results for 'Migratory Beekeeper'. (ref:size-permutation)"}
include_custom("output/figs/migratory-permutation.png")
```

```{r model-qq-residuals, out.width="80%", include=T, fig.cap="Distribution as QQ-plot for the model residuals after fitting the selected best model 'Base + Treatment Methods' on the training dataset."}
include_custom("output/figs/model-qq-residuals.png")
```

```{r motivation-state-rank, include=T}
cn <- c("State", "Category", "Answers [%]", "Rank")
c <- "Top 7 motivation answers for choosing varroa control method for each state separately. Participants could select out of a pool of 21 categories and where able to pick up to 5 categories maximum. 'Answers' is the relative amount the corresponding category was picked. Rank simply the decreasing order of the relative value in each state."
tab <- knitr::kable(
    r_motivation$state_rank,
    align = c("l", "l", "r", "c"),
    booktabs = TRUE,
    row.names = FALSE,
    col.names = cn,
    caption = c
) %>%
    kable_styling(latex_options = "scale_down") %>%
    kableExtra::collapse_rows(
        1,
        valign = "top",
        latex_hline = "major",
    )
tab
rm(tab, c, cn)
```

```{r motivation-comb-full, include=T}
cn <- c("Motivation", "Answers", "Motivations")
c <- "Top 25 motivation answers combinations for choosing varroa control method given by participants. Participants could select out of a pool of 21 categories and where able to pick up to 5 categories maximum. 'Answers' is the total number this combination was picked and 'Motivations' are the number of different categories in this combination, whereas 1 would indicate a single answer."
tab <- knitr::kable(
    r_motivation$comb_list %>% select(motivation, n, motivations) %>% dplyr::slice_max(n, n = 25),
    align = c("l", rep("r", 2)),
    booktabs = TRUE,
    row.names = FALSE,
    col.names = cn,
    # format = "markdown",
    caption = c
) %>%
    kable_styling(latex_options = "scale_down")
tab
rm(tab, c, cn)
```



