# (APPENDIX) Appendix {-}
# Appendix

(ref:size-permutation) Median difference null distribution (gray bars) based on 5,000 permutations. Red line is the point estimate for the median difference with yellow area as the 95% confidence interval. Red shadow indicating the calculated two-sided *p*-value range on the null distribution. 

(ref:size-permutation2) The *p*-values reported as "\*\*\*" would need more permutation to calculate a precise *p*-value.

```{r size-permutation, out.width="100%", include=T, fig.cap="Two level factor analysis statistical test results for 'Operation Size'. (ref:size-permutation) (ref:size-permutation2)"}
include_custom("output/figs/size-permutation.png")
```

```{r organic-permutation, out.width="100%", include=T, fig.cap="Two level factor analysis statistical test results for 'Certified Organic Beekeeper'. (ref:size-permutation)"}
include_custom("output/figs/organic-permutation.png")
```

```{r migratory-permutation, out.width="100%", include=T, fig.cap="Two level factor analysis statistical test results for 'Migratory Beekeeper'. (ref:size-permutation)"}
include_custom("output/figs/migratory-permutation.png")
```

```{r model-qq-residuals, out.width="80%", include=T, fig.cap="Distribution as QQ-plot for the model residuals after fitting the selected best model 'Base + Treatment Methods' on the training dataset."}
include_custom("output/figs/model-qq-residuals.png")
```

```{r motivation-state-rank, include=T}
cn <- c("State", "Category", "Answers [%]", "Rank")
c <- "Top 7 motivation answers for choosing varroa control method for each state separately. Participants could select out of a pool of 21 categories and where able to pick up to 5 categories maximum. 'Answers' is the relative amount the corresponding category was picked. Rank simply the decreasing order of the relative value in each state."
tab <- knitr::kable(
    r_motivation$state_rank,
    align = c("l", "l", "r", "c"),
    booktabs = TRUE,
    row.names = FALSE,
    col.names = cn,
    caption = c
) %>%
    kable_styling(latex_options = "scale_down") %>%
    kableExtra::collapse_rows(
        1,
        valign = "top",
        latex_hline = "major",
    )
tab
rm(tab, c, cn)
```

```{r motivation-comb-full, include=T}
cn <- c("Motivation", "Answers", "Motivations")
c <- "Top 25 motivation answers combinations for choosing varroa control method given by participants. Participants could select out of a pool of 21 categories and where able to pick up to 5 categories maximum. 'Answers' is the total number this combination was picked and 'Motivations' are the number of different categories in this combination, whereas 1 would indicate a single answer."
tab <- knitr::kable(
    r_motivation$comb_list %>% select(motivation, n, motivations) %>% dplyr::slice_max(n, n = 25),
    align = c("l", rep("r", 2)),
    booktabs = TRUE,
    row.names = FALSE,
    col.names = cn,
    # format = "markdown",
    caption = c
) %>%
    kable_styling(latex_options = "scale_down")
tab
rm(tab, c, cn)
```



