---
documentclass: scrreprt
papersize: a4
classoption:
- twoside
header-includes:
- \usepackage{hyperref}
- \usepackage{float}
- \usepackage{siunitx}
- \usepackage{threeparttable}
- \usepackage{multirow}
- \usepackage{pdflscape}
- \usepackage{colortbl}
bibliography: bib-refs.bib
biblatexoptions: [
    backend=biber,
    style=authoryear,
	sortcites=false,
	natbib=true,				
	hyperref=true,			
	backref=false,		
	isbn=false,					
	url=false,					
	doi=false,
    date=year,			
	urldate=long,
	maxbibnames=99,
	maxcitenames=2,
	uniquelist=false
    ]
link-citations: yes
---
```{r, warning=F, echo=F, include=F}
# special case for PDF and table as they otherwise wont work
if (knitr::is_latex_output()) {
    options(knitr.table.format = "latex")
} else {
    options(knitr.table.format = "html")
}
knitr::opts_chunk$set(fig.pos = "pht", out.extra = "")
```

<!--chapter:end:index.Rmd-->

# Introduction

The economic and ecology value and the importance of feeding mankind by pollinating crops of wild bees and honeybees, *Apis mellifera*, cannot be denied [@kearns1998;@klein2007;@gallai2009]. Furthermore managed honey bee colonies improve the quality and amount of harvest for various crops, as example oilseed rape, sunflowers and brassica vegetables [@krause1981;@sushil2013]. In an economic research a market analysis from England showed the value of pollination, honey, and beeswax estimated by £&nbsp;800 per colony in the year 1992 [@carreck1998]. If one adds the society benefit from pollination as done in a study in the U.S. the worth of honeybees was estimated with USD 1.6-8.3 billion [@southwick1992]. Additionally, @southwick1992 made it clear in their discussion that we do not only profit from honeybees but also from wild pollinators and should improve their environment and create conservation zones. There may be a lot more benefits which are often overlooked, for example the bee colonies fertilizing effect of bee frass in foraging areas, which at least under study control did show a significant release of inorganic nitrogen into the soil [@mishra2013]. Even though we learn each day more about the importance of insects in our ecosystem an alarming decline of wild pollinators can be seen over the last years [@hallmann2017;@sanchez-bayo2019]. The reasons for the decline seem manifold: parasites; pathogens; environmental pollution; rapid climate change, loss of insect friendly habitats etc. [eg. @biesmeijer2006;@potts2010;@goulson2015;@woodcock2017].

With the economic and ecology value of honey bees in mind I want to look at the expenses and treatment efficiency of varroa mite control methods in Austria. The varroa mite, *Varroa destructor*, is worldwide with some exceptions, an active threat to apiculture [eg. @dahle2010;@rosenkranz2010;@genersch2010;@morawetz2019;@noel2020;@roth2020]. The pest lead to the increase of costs in labour and material for beekeeping operations by expenses on treatment and replacement of lost colonies due to the varroa [@nationalresearchcouncil2007;@vanengelsdorp2010a]. In the US, were pollination services are a big industry for beekeepers, the spread of varroa mites and colony collapse disorder (CCD) did came at the same time as higher prices paid by farmers for pollination service [@carman2011;@rucker2012]. Indicating in real world examples that the introduced varroa mite and colony losses to not only impact economically beekeepers but also the food production industry and making the importance of pollination services of insects visible [@kearns1998;@klein2007;@gallai2009].

In Austria and most other central european countries the beekeepers have to cope with this imported aggressor since the 1980s [@griffiths1981;@potts2010a]. Beekeepers need to treat their colonies to prevent unacceptable level of colony losses, as most untreated colonies would die off within 2-3 years [@rosenkranz2010]. The first years after the mite domination in Europa and the resulting treating effort in combination with other socio-economic factors may have led to a sharp decline of beekeepers [@potts2010a;@brodschneider2011]. Although in Austria the number of beekeeper and honeybee colonies is steadily growing over the past 7 years.[^biene-österreich] The control of the mite in Austria is mainly done with organic acids (eg. formic acid, oxalic acid), essential oils (thymol), synthetic acrazides or biotechnical methods [@oberreiter2020]. Each of these treatment methods has it advantages and disadvantages and most of the time beekeepers use a combination of them [@rosenkranz2010;@noel2020]. Though, the treatment alone does not guaranteed the survival of colonies, studies show that the education, experience and correct application and timing of treatment plays an important role [@brodschneider2013;@vanderzee2014;@jacques2017;@morawetz2019].

Additionally I want to look at the main motivations for choosing a treatment method. This empirical approach to evaluate the motivations behind management practices against the varroa mites was never done before in Austria. Somewhat similar studies from the U.S. did identify different types of philosophies in management practices [@thoms2019;@underwood2019]. The focus of these studies was quite novel to the already rich literature of most topics in beekeeping. As the U.S. is quite different in their beekeeping structure, it is important to create similar socioecological studies in Austria.

In the foreground of this work is an exploratory analysis of a citizen science survey from the years 2018/19, 2019/20 and 2020/21. The survey is carried out since 2008 by the Institute of Biology at the University of Graz to collect data on winter losses of managed honeybee colonies [@brodschneider2013]. In the past three years the participants could also answers an optional question about their expenses of treatment per honeybee colony, without labor costs and in the last year an additional question why they chose their varroa control methods. These two question are the main focus of this study, as such a descriptive analysis of treatment expenses and motivational reasons has never been carried out in Austria in this form before. Furthermore, the selected treatment methods and their efficiency are compared in combination with their expenses to give valuable input for beekeepers in their decision making process on which methods are a good option for their operation.

[^biene-österreich]: National beekeeping association 'Biene Österreich' \url{https://www.biene-oesterreich.at/}, accessed: 12/2020.

<!--chapter:end:01-intro.Rmd-->

# Material and Methods

```{r load-files, warning=F, echo=F, include=F}
source("src/setup.R", local = knitr::knit_global())
source("src/mm-cleanup.R", local = knitr::knit_global())
```

```{r, include=FALSE}
if (file.exists("output/r-model.rds")) {
  r_model <- readRDS("output/r-model.rds")
} else {
  source("src/r-model.R", local = knitr::knit_global())
}
```


## Survey

The survey is a citizen science survey which is done each year in spring since 2008. It is a combination of online, beekeeping journal and paper submission. The main focus and question of the survey are the yearly winter honeybee colony losses in Austria. The survey is part of the COLOSS questionnaire and most questions are in accordance with other participating countries [@vanderzee2013]. The questions are direct and relate to possible risk factors which may influence overwintering success as seen in previous studies in Austria [@brodschneider2010a;@oberreiter2020]. The survey is only obligatory for beekeepers who want to participate in the Austrian bee health programme. The confirmation of participation is one of the options which allows a beekeeper to apply and/or get a higher rate of government aid for investments in technical equipment to improve production quality and hygiene in their beekeeping operation.^[AgrarMarkt Austria, 06/2021 - Version 04]

As most questions are related to winter colony losses I am only interested in a subset of questions for my research question: number of colonies wintered (size of operation), number of colonies lost over the winter period (due to natural or queen loss, without losses due to natural disaster), state and district of the main wintering apiary (as possible covariate), month(s) and type of treatment method used. In addition two operational questions are also explored, if the beekeeper migrates with her/his colonies and if she/he is an organic certified beekeeper, with the possible answers 'Yes', 'No' and 'Unsure'. The answers of these questions of interested are explored in combination with the answer on the question about estimated treatment expenses per colony without labor costs. Participants could answer this question with any positive number from zero and without any upper limit. The treatment expenses question was only available in the Austrian COLOSS questionnaire and only in the online survey as the questionnaire which is printed in the beekeeping journal and handed out paper submissions are more compact versions.

To further comprehend why participants choose their varroa control method, the last survey year 2020/21 did additionally include the question: 'Which main motivations are most likely to apply to you when choosing the Varroa control method (max. 5)?'. The possible answers, which were arbitrary defined, were displayed in a randomized order and included: 'Habit', 'Economic costs', 'Marketing', 'Subsidies', 'Prescription', 'Philosophy', 'Efficacy', 'Method applicable with and without brood', 'Time of season where they can be applied', 'Colony strength', 'Colony infestation level (Varroa)', 'Environmental conditions', 'Recommendations from beekeeping associations, extension workers, veterinarians, magazines, etc.', 'Legal status (regulatory framework)', 'Availability of products', 'Resistances', 'Residues', 'Side effects on bees', 'Level of difficulty and needed equipment', 'Time and number of applications', 'Other'. Again, this question was only available in the online survey. A total of `r length(mmList$motivation$no_answer)+12` participants did choose not to answer this question or did by accident choose more than five answers, which leaves `r mmList$motivation$valid_answers` valid answers to this new question.
<!-- the +12 are participants who could not finish the survey as they did answer more than 5 categories and had to be manually imported -->

The survey was anonymous and private data like IP addresses were only used in the first step to help identify possible double entries. In order to protect the personal data of the participants, a privacy statement was drawn up between the primary persons involved in the evaluation. If voluntary contact details were given, it was only used for enquiries to correct missing or incorrect data. All personal data were anonymised prior analysis.

As I analysed three continuous survey years and the survey is anonymous I have no assumption if the same beekeepers participated in more than one survey. Therefore, I can not tell which survey replies are over the years independent or not and analysed them separately. Only for a handful of beekeepers who did consistently give the same contact details this would be possible. The advantages to compare multiple years are, that I may have a more thorough overview if the survey is robust in the results, as there should not be much change in the expenses and distribution of beekeepers apiary location and size of operations over the survey years. Furthermore, overall winter loss rates fluctuate between years of the surveys. This means I can check if a given treatment method did result in all survey years in low or high winter mortality or not.

## Expenses Estimation

Expenses estimation was carried out to check if there are discrepancies between participants answers and estimates. To calculate the estimates, I researched online store prices in 2020 in Austria for official registered medication[^mm-1] and used common application amount from beekeeping school materials and manufacturer instructions (Table&nbsp;\@ref(tab:standard-cost)).

[^mm-1]: Austrian medicinal product index: <https://aspregister.basg.gv.at/>, accessed November 2020

```{r standard-cost, include=TRUE}
c <- "The calculated estimates for treatment expenses per method. Investments can be used long term (I calculate 10 years, eg. hyperthermia equipment) and are divided by the number of colonies. Material can also be used long term (7 years, eg. queen cages) but needs to be bought for each colony. Consumables are used for each colony (eg. formic acid long term was calculated with 200ml use of formic acid)."
cn <- c("Methode", "Investment", "Material", "Consumables")
tab <- kbl(
  treatmentList[-1, c(3, 5:7)],
  booktabs = TRUE,
  row.names = FALSE,
  # format = "markdown",
  caption = c,
  col.names = cn,
  linesep = c("", "", "\\addlinespace"),
) %>%
  footnote(threeparttable = T, general = "'Synthetic methods' include Amitraz, Coumaphos, and other synthetic methods; 'Another methods' excludes synthetic methods.") %>%
  kable_styling(latex_options = "HOLD_position")
tab
rm(c, cn)
```

An example calculation, based on equation \@ref(eq:estimate-calculation) with the given estimates and the average month of usage for each treatment for a total of 10 colonies can be seen in Table&nbsp;\@ref(tab:common). The 10 colonies are the median number of colonies in the survey. With this equation an estimate was calculated for each survey participant with the given answer for treatment method (or combination), number of colonies and number of months used.

```{=tex}
\begin{equation}
  \begin{aligned}
    \text{estimate/colony} = (\text{investment} / \text{total colonies}) + \text{material} + \text{consumable} * \text{months}
  \end{aligned}
  (\#eq:estimate-calculation)
\end{equation}
```

```{r common, include=T}
# colonies based for estimate
m_colonies <- 10
col <- dfData %>%
  select(!!(paste0(treatmentList$ttotal, "12")))
# Get number of participates using the single treatment methods
sin <- col %>%
  mutate(across(.cols = everything(), as.logical)) %>%
  colSums()
# get the average amount for number of months
nu <- col %>%
  na_if(0) %>%
  colMeans(na.rm = T)

common_single <- tibble(
  tname = treatmentList$tname,
  n = sin %>% ft(),
  p = round(sin * 100 / nrow(dfData), 2),
  nu = round(nu, 2),
  est = round(treatmentList$investment / m_colonies + treatmentList$material + treatmentList$consumables * nu, 2)
) %>% arrange(desc(p))

c <- paste0("Treatment methods ordered by usage over the survey years, months indicating the average of months the corresponding method was used. Estimated expenses per colony are calculated with ", m_colonies, " colonies, which represents the median number of colonies in the survey. The average months are based on the average number of reported months for each treatment method separately.")

cn <- c("Method", "Total [n]", "Percent [%]", "Average Months [n]", "Estimated expenses [EUR]")
kable(
  align = c("l", rep("r", 4)),
  common_single[-1, ],
  booktabs   = TRUE,
  row.names = FALSE,
  col.names  = cn,
  # format = "markdown",
  caption    = c,
  linesep = c("", "", "\\addlinespace"),
) %>%
  kable_styling(latex_options = c("HOLD_position", "scale_down"))

rm(cn, c, common_single, nu, sin, col, m_colonies)
```

## Data Cleanup

### Newspaper and Paper Submission

The compressed version of the survey, which was printed in a beekeeping journal and paper handout submissions, did not include the question about expenses and, therefore, were excluded. That concerns a total of `r sum(mmList$submitted_all %>% filter(submitted != "Internet") %>% pull(n)) %>% ft(.)` submission from `r sum(mmList$submitted_all$n) %>% ft(.)` over the three years of surveys. With one exception, as one person did send a filled out paper handout of the online survey.

### No Answers on Expenses

Not all participants did answers the question on expenses even when they could do so. These participates were also excluded as this question represent the key focus of the study. In numbers these are in 2018/19 a total of `r mmList$internet_no_answer %>% filter(year == "18/19") %>% pull(nn)` (`r mmList$internet_no_answer %>% filter(year == "18/19") %>% pull(np)`%), in 2019/20 a total of `r mmList$internet_no_answer %>% filter(year == "19/20") %>% pull(nn)` (`r mmList$internet_no_answer %>% filter(year == "19/20") %>% pull(np)`%) and in the last survey year 2020/21 a total of `r mmList$internet_no_answer %>% filter(year == "20/21") %>% pull(nn)` (`r mmList$internet_no_answer %>% filter(year == "20/21") %>% pull(np)`%) answers.

### No Treatment

Only `r nrow(mmList$no_treatment %>% filter(costs == 0))` participants performed no treatment against the varroa mite and answered with zero expenses. In addition `r nrow(mmList$no_treatment %>% filter(costs > 0))` participants answered they did no treatment but have included expenses for varroa treatment. This illogical entries and the participants with zero costs and the before mentioned no treatment answers were removed from the data analysis.

### No Treatment Methods

In total `r nrow(mmList$no_method)` participants did not specify their treatment methods but said they treated against varroa and answered on varroa treatment expenses (min. EUR&nbsp;`r min(mmList$no_method$costs)`, max. EUR&nbsp;`r max(mmList$no_method$costs)` per colony). I decided to discard them from the dataset, as it was only a small number and I could not further validate the expenses answers or use it for statistics which involved the treatment methods.

### Zero Expenses

In total `r nrow(mmList$cost_zero$data)` participants said their treatment expenses were zero. Only one answer was kept, as the treatment was a combination of hyperthermia and biotechnical and I cannot make an assumption about the real expenses of the participant. Of course the equipment for the hyperthermia does cost something, but I do not know more details, as she/he could already have the equipment for a long time. A total of `r length(mmList$cost_zero$id_remove)-length(mmList$cost_zero$id_sponsor)` were removed as they said zero expenses but in fact did treatment with organic acids, which is illogical as the organic acids are consumed and not for free. Additionally `r length(mmList$cost_zero$id_sponsor)` participants received a sponsorship (beekeeping club or local community) for their treatment expenses and had therefore zero expenses, which were also removed.

### Outlier Detection and Imputation

To detect outliers in the dataset, two methods were used. For the maximum limit I decided to filter the dataset based on the arbitrary decision if the expenses per colony were equal or more than double of the whole dataset extreme outlier (EUR&nbsp;`r mmList$cost_upper$upper_limit`), as defined in equation \@ref(eq:extreme-outliers).

```{=tex}
\begin{equation}
\begin{aligned}
x >= (Q3 + 3 * IQR) * 2
\end{aligned}
(\#eq:extreme-outliers)
\end{equation}
```

This did detect `r nrow(mmList$cost_upper$data)` anomaly values based on the maximum limit and without any information of the given treatment method. The second one is model based with the isolation forest approach [@liu2008], implemented in the R package [@isotree2021]. Input for the model represent the expenses per colony and a dummy column which was encoded for each treatment method combination without drone brood removal. The parameters were set to contain all rows of the dataset, 100 binary trees and split on a single variable. I defined values above or equal to 0.6 in the resulting standardized outlier score, as described in @liu2008, as anomaly values, which resulted in a list of `r nrow(mmList$cost_upper$forest)`.

The two approaches did mostly overlap in their results and the lists were combined, which resulted in `r nrow(mmList$cost_upper$combined_list)` possible outliers. Further investigating of these outlier answers revealed that most of them did probably answer total expenses per operation and not expenses per colony as asked in the survey. I removed from the outliers list participants which used Hyperthermia as treatment method, which could involve considerable investment, especially if it was only used for one year. This was also stated by some beekeepers as comments in the survey. In addition participants were also removed from the outliers list if plausible reasons in the comments were given. As for the rest of the extreme outliers (*n*=`r nrow(mmList$cost_upper$new_data)`), I divided for each answers the expenses by the reported number of colonies going into winter. This means the manipulative impact was more notable for participants which included high expenses above the limit and also reported a large number of colonies and less for small operations, were the high expenses are possible true. The new calculated costs were inspected to be plausible, e.g. not too low, and were updated in the cleaned dataset for further analysis.

```{r isolation-forest, include=T, fig.cap="Isolation forest standardized outlier score to expenses per colony of the whole dataset. Black points are within the defined average range and pink coloured points indicating possible outliers. Red background area above the set average limit of >= 0.6 outlier score."}
include_custom("output/figs/isolation_forest.png")
```

### Final Data Count

After the afore mentioned data cleaning steps, which filtered survey responses with valid answers for expenses per colony, the number of answer was reduced by around 21-24% (Table&nbsp;\@ref(tab:reports)).

```{r reports, include=T}
cn <- c("Year", "Total [n]", "Valid Answers [n]", "Percent [%]")
c <- "Number of participants with valid answers on the question for estimated treatment expenses per colony and the total number of survey participants."
tab <- knitr::kable(
  mmList$reports %>%
    mutate(
      year = paste0("20", year),
      across(is.numeric, ft)
    ),
  align = c("c", rep("r", 3)),
  booktabs = TRUE,
  row.names = FALSE,
  col.names = cn,
  caption = c
) %>%
  kable_styling(latex_options = "HOLD_position")
tab
rm(cn, c, tab)
```

## Statistics

The analysis of the data was carried out with R [@r2020], plotting and data manipulation was done mainly with the tidyverse package collection [@tidyverse2019]. The code, including a listing of the remaining packages not mentioned here, is available on GitHub under MIT licence[^mm-2]. As the particular goal of the study is an exploratory data analysis (EDA) I try to rely predominantly on descriptive statistics and visualise the data behind it. The use of inference based statistics and model generation is merely used for explorative purpose and no focus on the *p*-value alone is given. Reported *p*-values are uncorrected for alpha error accumulation, this could lead to incorrectly rejecting the null hypothesis. Hypothesis confirmation statistics should normally only be used for a predefined hypothesis, which can be falsifiable, and experimental design directly related to the research question, which is untrue for this analysis. As I look at each observation/report multiple times I do by definition exploratory analysis.

[^mm-2]: Github: <https://github.com/HannesOberreiter/treatment-expenses-austria>

### Operational Factor Analysis

The operational factor analysis (Migratory Beekeeper and Certified Organic Beekeeper) did include as possible answers 'Yes', 'No' and 'Uncertain'. The category 'Uncertain' and participants which did not answers the question were removed in the statistical two levels factor analysis (Table&nbsp;\@ref(tab:factor-removed)). The *p*-value is generated from the median difference between groups in comparison to the permutation generated null distribution (5,000 permutations), significant results are *p* $\leq$ 0.05. For the median point estimate a confidence interval (95%) was calculated based on 1,000 bootstrap resamples. The visual statistical test results of the difference between groups are attached in the appendix section.

### Multiple Regression

To evaluate if and how well the selected questions can be used for prediction and explain the treatment expenses a simple linear multiple regression model with treatment expenses per colony as dependent variable was fitted. The model creation and evaluation was done with the tidymodels package [@tidymodels2020]. Data preprocessing included normalization with $\log_{10}$ transform of number of colonies wintered and expenses of treatment to remove the effect of the right skewed data distribution, whereas beforehand one single zero expenses answer was removed (Fig.&nbsp;\@ref(fig:model-qq)). Same as in the two level operational factor analysis in the multiple regression models only participants which did answer 'Yes' or 'No' for the two operational factor questions were used in the model to simplify it (Table&nbsp;\@ref(tab:factor-removed)). The levels were dummy transformed to 0 (False) for 'No' and 1 (True) for 'Yes'.

(ref:model-qq) QQ-Plot of $\log_{10}$ transformed for normalization of (A) treatment expenses per colony and (B) number of honeybee colonies wintered.
```{r model-qq, include=T, fig.cap="(ref:model-qq)"}
include_custom("output/figs/model-qq.png")
```

Treatment methods were binary transformed, if the participants did practise it in any month to 1 (True) otherwise if not at all applied to 0 (False). The model fitting was done with 90% of the original dataset (*n*=`r nrow(r_model$train_data)%>%ft()`) and out of sample evaluation of accuracy was done with 10% (*n*=`r nrow(r_model$test_data)`). Best model selection was selected based on lowest AIC and BIC value.

```{r factor-removed, include=T}
# cn <- c("Question", "Year", "Participants [n]", "Answer")
cn <- c("Year", "Participants [n]", "Answer")
c <- "Number of participants answering with 'Uncertain' or not at all for the operational questions. These beekeepers were removed in the statistical two factor analysis and multiple regression model."
tab <- knitr::kable(
  mmList$single_factor_drop %>%
    mutate(
      year = paste0("20", year),
      year = ifelse(lag(year, default = "0") == year, "", year)
    ) %>%
    # select(name, year, n, value),
    select(year, n, value),
  booktabs = TRUE,
  row.names = FALSE,
  col.names = cn,
  align = c("c", "r", "l"),
  caption = c
) %>%
  pack_rows("Certified Organic Beekeeper", 1, 6) %>%
  pack_rows("Migratory Beekeeper", 7, 9) %>%
  kable_styling(latex_options = "HOLD_position")
tab
rm(cn, c, tab)
```

### Treatment Methods and Loss Rates

The treatment methods are selected in accordance with COLOSS, and the same list of treatment methods are used in every participating country [@vanderzee2013]. In the survey a total of 18 different treatment methods as varroa control were selectable by the participants and in addition as binary question the time of application. The treatment methods included 'Drone brood removal', 'Oxalic acid' (sublimation and mixture or pure trickling), 'Formic acid' (long or short term), 'Other biotechnical methods' (excluding drone brood removal and hyperthermia), 'Thymol', 'Hyperthermia', 'Lactic acid', 'Another method' and 'Synthetic methods' which are a combination of the treatment categories 'Amitraz (in strips or sublimation), Coumaphos (Perizin or Checkmite+) or Flumethrin (Bayvarol, Polyvar), Tau-fluvalinat (Apistan)', because of the limited number of replies in Austria concerning these methods.

To generate groups of treatment method combinations the individual months of the treatment applied were grouped into following seasons 'Spring' (April-May), 'Summer' (June-October) and 'Winter' (November - January), as done previously [@oberreiter2020]. The months February and March of the next year were not used in the combinations. In addition to reduce the number of possible different treatment method combinations the method 'Drone brood removal' was excluded, as it led to too many combinations and should at least not influence the expenses.

The observed total loss rate for these treatment combinations was calculated based on equation \@ref(eq:lossrate). The reported 95% confidence interval (CI) were created for each analysed treatment combination from a separate null model, which was calculated using a Generalised Linear Model (GZLM) with quasi-binomial distribution and the link function \enquote{logit}. The procedure was carried out according to scientifically established methods \citep{vanderzee2013}.

```{=tex}
\begin{equation}
\begin{aligned}
\frac
    {
      \sum{\text{Loss of colonies}}-\sum{\text{Loss due to natural hazards}}
    }
    {
      \sum{\text{Colonies wintered}}
    }
      *100
\end{aligned}
(\#eq:lossrate)
\end{equation}
```


### Efficacy & Economical

To test if varroa treatment combinations, with at least 30 answers over all three years, are efficient and also economical the observed loss rates and reported mean expenses were plotted and grouped into four categories. The categories comprise the combination of low or high losses (colony winter losses) and low or high expense (expenses per colony) for each method combination (Fig. \@ref(fig:efficient-economic-mm)). The calculation and assignment to these categories is based on the value for each treatment combination in comparison to the total observed loss rate and total mean expenses per colony for each year. If the mean expenses for a method is lower or equal to the average, it will get the category 'Low-Expense' and if higher 'High-Expense'. The identical method applies to the loss rate, but here the treatment methods loss rates are compared to the total observed loss rate. The separation for each year was performed because the total loss rate does change between years [@brodschneider2019a;@oberreiter2020]. As for the expenses per colony, the total means value could change over the years due to different frequency of applied treatment methods, for example an increase of biotechnical methods over traditional formic acid treatment [@oberreiter2020]. Although, the expenses could probably be summarised over all three years as they will not change drastically in this timeframe, but I also have no assumption if the same beekeepers employed the same methods and participated in more than one survey, which could bias the result.

<!--
# we could do the geometric mean for costs to reduce effects of outliers, but the result is not straightforward interpretable and the mean is maybe more true in this case, as we already to a lot of pre filtering to reduce the outliers.
-->

(ref:efficient-economic-mm) Graphical explanation of the assignment of efficacy and economical categories for the treatment method combinations. This assignment was done for each year separately.
```{r efficient-economic-mm, include=T, fig.cap="(ref:efficient-economic-mm)"}
include_custom("output/figs/efficient-economic-mm.png")
```

<!--chapter:end:02-mm.Rmd-->

# Results
## Survey Data
### Representative and Distribution of Sample

```{r, include=FALSE}
source("src/r-survey.R", local = knitr::knit_global())
```

The survey participants with valid answers including treatment expenses per colony represent between 3.5% and 4.0% of the beekeepers in Austria. When compared to the number of registered beekeeper and honey bee colonies as reported by the national beekeeping association 'Biene Österreich'.[^r-vis] In detail these are in the survey year 2018/19 with `r r_survey$response$rate_beekeeper[[1]]`% (*n*=`r format(r_survey$response$survey_colonies[[1]], big.mark = ",")`), survey year 2019/20 with `r r_survey$response$rate_beekeeper[[2]]`% (*n*=`r format(r_survey$response$survey_colonies[[2]], big.mark = ",")`) and 2020/21 with `r r_survey$response$rate_beekeeper[[3]]`% (*n*=`r format(r_survey$response$survey_colonies[[3]], big.mark = ",")`) (Table&nbsp;\@ref(tab:registered)). The distribution map with participants grouped by main winter apiary location on district level shows that the survey covers most districts in Austria in the three survey years. The distribution of beekeepers between the districts is not the same which could reflect real beekeeper density and/or a possible participation bias, although the size of districts is also unequal (Fig.&nbsp;\@ref(fig:distr-map)).

[^r-vis]: In Austria the registration of beekeepers, apiaries and honeybee colonies in the Consumer Health Information System (Verbrauchergesundheitsinformationssystem - VIS) has been implemented into the national Animal Identification and Registration Regulation 2009 (Tierkennzeichnungs- und Registrierungsverordnung, § 1 TKZVO 2009 idF II 193/2015) from 8th of July 2015.

```{r registered, include=T}
cn <- c("Year", "Total", "Survey", "%", "Total", "Survey", "%")
c <- "Percentage of survey participants in relation to the number of beekeepers in Austria as reported by the national beekeeping association 'Biene Österreich'."
tab <- knitr::kable(
  r_survey$response %>%
    select(year, beekeeper, survey_beekeeper, rate_beekeeper, colonies, survey_colonies, rate_colonies) %>%
    mutate(
      across(-year, ft)
    ),
  booktabs = TRUE,
  row.names = FALSE,
  col.names = cn,
  caption = c
) %>%
  kableExtra::add_header_above(c("", "Beekeeper [n]" = 3, "Colonies [n]" = 3))
tab
rm(tab, c, cn)
```

```{r distr-map, include=T, fig.cap="Distribution heatmap of the participants main winter apiary on district level. Darker colours are more participants in corresponding district and lighter colours are less. Shapefiles under Creative Commons https://www.data.gv.at/"}
include_custom("output/figs/map.png")
```

The reported expenses per colony arithmetic mean (EUR&nbsp;`r paste(range(r_survey$summary$survey_table$mean), collapse = "-")`) and median (EUR&nbsp;`r paste(sort(range(r_survey$summary$survey_table$median), decreasing = TRUE), collapse = "-")`) is similar between the three survey years (Table&nbsp;\@ref(tab:central)). The arithmetic mean is higher than the median, due to the right skewed data distribution with some high values above EUR&nbsp;100 (Fig.&nbsp;\@ref(fig:distr-year)). The distribution density of the reported expenses are similar between the three survey years and show that participant did answer in round full numbers eg. EUR&nbsp;5.00, 10.00, 15.00 (Fig.&nbsp;\@ref(fig:distr-year)).


(ref:distr-year) Distribution of the expenses for each survey year, values above EUR&nbsp;50 are not visible. First graph column represents the actual data points, second column, a boxplot with median and interquartile range (IQR, $25^{th}$ - $75^{th}$ quantile) and vertical lines 1.5*IQR. Last column is density distribution of the data points. Different years are colour coded (orange = 2018/19, green = 2019/20, blue = 2020/21).

```{r distr-year, out.width="100%", include=T, messages=F, warnings=F, fig.cap="(ref:distr-year)"}
include_custom("output/figs/distr-year.png")
```

```{r central, include=T}
cn <- c("Year", "Minimum", "Mean", "Geom. Mean", "Median", "Max", "n")
c <- "Survey central tendency values for treatment expenses per colony for all years and additionally over all years. Column 'Mean' is the arithmetic mean, the 'Geom. Mean' (geometrical mean) was added as alternative to reduce the effect of high outliers."
tab <- knitr::kable(
  dplyr::bind_rows(r_survey$summary$survey_table, r_survey$summary$survey_total) %>%
    select(-type) %>%
    mutate(
      year = ifelse(year == "total", "All Years", paste0("20", year)),
      n = ft(n)
    ),
  align = c("l", rep("r", 5)),
  booktabs = TRUE,
  row.names = FALSE,
  col.names = cn,
  # format = "markdown",
  caption = c
) %>%
  kableExtra::row_spec(3, hline_after = TRUE)
tab
rm(tab, c, cn)
```

### Distribution of Participants for Operational Questions {#operational-survey}

As overview for correlation and distribution of the included operational factors of interest, the participants where grouped into bins of combination of these factors (Fig.&nbsp;\@ref(fig:distr-groups)). The two survey questions 'Migratory Beekeeper' and 'Certified Organic Beekeeper' did include as possible answers 'Yes', 'No' and 'Uncertain', whereas 'Uncertain' and participants which did not answer to these questions were ignored. The answer on operation size (number colonies going into winter, any positive integer number was possible) was grouped into two levels for factor analysis. These levels are 'Recreational Beekeeper' with 25 or less colonies and 'Professional Beekeeper' with more colonies. Based on these three operational questions the most common and dominating group in the survey are 'Recreational Beekeeper', who are neither migratory or certified organic with 63-65%. Second most are 'Professional beekeeper' with more than 25 colonies with 8-11%. All other combinations of these operational factors are below 10% in each survey year. A slight decline in 'Recreational beekeeper' which report to migrate their colonies can be seen over the three survey years (2018/19: 7,7%; 2019/20: 6.2%; 2020/21: 5.4%) (Fig.&nbsp;\@ref(fig:distr-groups)).

```{r distr-groups, include=T, messages=F, warnings=F, fig.cap="Count and distribution of participants grouped into operational factors and their combination for each year. The three operational factors and combination bins are below the bar plots, whereas black dots indicate 'Yes'/'True' for the given factor. Percentage reported above the bars stands respectively for the survey years 2018/19, 2019/20 and 2020/21. Recreational Beekeepers are participants with 25 or less colonies going into winter.", out.width="100%"}
include_custom("output/figs/distr-groups.png")
```

<!--chapter:end:03-results-1-survey.Rmd-->

## Motivation choosing Varroa Control Method

```{r, include=FALSE}
source("src/r-motivation.R", local = knitr::knit_global())
```

The question "Which main motivations are most likely to apply to you when choosing the Varroa control method (max. 5)?" was only asked in the survey 2020/21, with a total of `r format(r_motivation$answers, big.mark = ",")` valid answers. Even though it was possible to pick up to five categories from a total selection of 21, the most common answer was the single answer of 'Recommendations from beekeeping associations, extension workers, veterinarians, magazines, etc.' (*n*=16). All other answer combinations in the top 10 had in common that they included 'Efficacy' as one of their answers, with third position only the sole answer of 'Efficacy' (*n*=10) (Table&nbsp;\@ref(tab:motivation-comb)). A more extensive list with the top 25 answers is available in the appendix (Table&nbsp;\@ref(tab:motivation-comb-full)).

The total count of how often one category was picked as motivation shows the dominance of 'Efficacy', which was picked by 67.3% (*n*=844) of the participants. Second most was with 49% (*n*=615) the category 'Side effects on bees', followed up by 'Residues' (42.9%, n = 538) and 'Infestation Level' (37%, *n* = 464). The answer 'Economic Costs', which remain the key topic of this research, was picked by 6.8% (*n*=80) beekeeper (Fig.&nbsp;\@ref(fig:motivation-count)).

```{r motivation-comb, include=T}
cn <- c("Motivation", "Answers", "Motivations")
c <- "Top 10 motivation answers combinations for choosing varroa control method given by participants. Participants could select out of a pool of 21 categories and where able to pick up to 5 categories maximum. 'Answers' is the total number this combination was picked and 'Motivations' are the number of different categories in this combination, whereas 1 would indicate a single answer."
tab <- knitr::kable(
    r_motivation$comb_list %>% select(motivation, n, motivations) %>% dplyr::slice_max(n, n = 10),
    align = c("l", rep("r", 2)),
    booktabs = TRUE,
    row.names = FALSE,
    col.names = cn,
    caption = c
) %>%
    kable_styling(latex_options = "scale_down")
tab
rm(tab, c, cn)
```

```{r motivation-count, include=T, messages=F, warnings=F, fig.cap="Total count of all 21 motivation answers for choosing varroa control method. Percentage is the relative amount of participants which did pick this motivation category as one of their answers, number in brackets is the absolute count how many times the answer was picked. Order from most at the top and least at the bottom.", out.width="100%"}
include_custom("output/figs/motivation-count.png")
```

To see if there are geographically differences in the picked categories the top 5 categories where evaluated for the nine federal states in Austria separately. The category 'Efficacy' was even so the most common motivation in all states. Starting with second place the relative number of answers for each category is different between the states. Whereas 'Residues' was second most for Styria and all other states with the category 'Side effects on bees'. Followed by the third most common category 'Residues' for Lower and Upper Austria, Burgenland, Salzburg, Styria and Tyrol, the category 'Legal status' for Vienna and Carinthia and the category 'Colony Infestation Level (Varroa)' in Vorarlberg (Fig.&nbsp;\@ref(fig:motivation-state)). One interesting difference was the placing of the category 'Residues', which was for most states at the third position but for Carinthia at 5$^{th}$ and for Vienna outside of the top five at the 7$^{th}$ position (Appendix Table \@ref(tab:motivation-state-rank)).

```{r motivation-state, include=T, messages=F, warnings=F, fig.cap="Top five motivation answers for choosing varroa control method. Stacked bar plot representing answers for each federal state in Austria in relative amount to the total answers for each motivation category. Numbers inside stacked bars are the relative number of answers in percent for each state. White background numbers indicate highest and lowest for given motivation category. States are color coded and sorted alphabetically.", out.width="100%"}
include_custom("output/figs/motivation-state.png")
```

As motivation could also be driven by the operation size, the results were split into beekeepers with 25 or less colonies (<=25) as 'Recreational Beekeepers' and 'Professional Beekeepers' with more than 25 colonies (>25) (Fig. \@ref(fig:motivation-operation)). 'Efficacy' was the most picked category for both groups (68%/67%). The category 'Residues' was more often picked by 'Professional Beekeepers' (58%/39%), whereas 'Side effects on bees' was the second most common for 'Recreational Beekeepers' (44%/50%). The categories 'Time and number of applications' (34%/25%) and 'Resistances' (28%/14%) showed a 1.5 - 2x higher rate for 'Professional Beekeepers'. For 'Recreational Beekeepers' the categories 'Recommendations from beekeeping associations, extension workers, veterinarians, magazines, etc.' (10%/31%) and 'Level of difficulty and needed equipment' (13%/23%) were in comparison to the bigger operations more important and more often chosen. The category 'Economic costs' (9%/6%) showed a higher relative count for 'Professional Beekeepers' (Fig. \@ref(fig:motivation-operation)).

```{r motivation-operation, include=T, messages=F, warnings=F, fig.cap="Relative count of all 21 motivation answers for choosing varroa control method separated into operation sizes. Operation size with 25 or less colonies ('<=25') are 'Recreational Beekeeper' (blue) and with more are 'Professional Beekeeper' ('>25') (green). The percentage is the relative amount of participants which did pick this motivation category as one of their answers, number in brackets is the absolute count how many times the answer was picked. Order by the sum of both percentages from most at the top and least at the bottom.", out.width="100%"}
include_custom("output/figs/motivation-operation.png")
```

<!--chapter:end:03-results-2-motivation.Rmd-->


## Expenses Estimations {#r-estimations}

```{r, include=FALSE}
if (file.exists("output/r_estimate.rds")) {
    r_estimate <- readRDS("output/r_estimate.rds")
} else {
    source("src/r-estimate.R", local = knitr::knit_global())
    saveRDS(r_estimate, "output/r_estimate.rds")
}
```

The self calculated estimated expenses, based on online research of product prices, central tendency values are similar to the survey responses in each year. The estimate mean is on average EUR&nbsp;`r round(mean(as.numeric(r_estimate$table$mean[1:3]) - as.numeric(r_estimate$table$mean[4:6])), 2)` higher than the survey mean and the estimate median is on average EUR&nbsp;`r round(mean(as.numeric(r_estimate$table$median[1:3]) - as.numeric(r_estimate$table$median[4:6])), 2)` higher than the survey median (Table&nbsp;\@ref(tab:table-estimates)).

```{r table-estimates, include=T}
c <- "Comparison of central tendency values for estimated expenses and the expenses as reported by participants in the survey for each year separately. All values are in Euro."
cn <- c("", "Min.", "Mean", "Median", "Max")

tab <- kable(
    r_estimate$table %>% select(-type),
    align = c("c", rep("r", 4)),
    booktabs = TRUE,
    row.names = FALSE,
    col.names = cn,
    caption = c
) %>%
    pack_rows("Estimate", 1, 3) %>%
    pack_rows("Survey", 4, 6)
tab
rm(c, cn, tab)
```

The comparison between the estimate and reported survey expenses not separated by year but for the different treatment method combinations (excluding drone brood removal) shows that the total overestimation is not true for all treatment methods, as some are also underestimated. This comparison can be visualized as differences plot (Fig.&nbsp;\@ref(fig:estimates-bland)). Typically this is done as a statistical method of agreement between methods [@bland1986]. On average the difference between treatment method combinations with at least 15 participants is close to zero (EUR&nbsp;`r round(r_estimate$ba$mean.diffs, 2)`), indicating in summary no over- or underestimation (Fig.&nbsp;\@ref(fig:estimates-bland), yellow line). Only one combination is outside the 95% confidence interval of the upper and lower two times the standard deviations from the mean difference. This combination of summer and winter oxalic acid trickling of readymade mixture ('SU-OX-Mix & Wi-Ox-Mix') was underestimated on average by EUR&nbsp;15, although the high difference is mainly driven by extreme outliers in this group (*n*=22, mean expenses survey = EUR&nbsp;19, median expenses survey = EUR&nbsp;9). Overestimation of the estimated expenses are shown for treatment combinations which include biotechnical methods excluding drone brood removal or hypothermia in summer ('SU-Biot') and oxalic acid sublimation in winter ('WI-Ox'). The three most common treatment combinations, formic acid long term in summer ('SU-Fa-LT') and winter oxalic treatment ('WI-Ox-sub', 'WI-Ox-pure', 'WI-Ox-mix'), show similar values for mean estimate and mean reported survey expenses (Fig.&nbsp;\@ref(fig:estimates-bland)). For the top 11 treatment method combinations and the comparison of the mean and median expenses of the estimate and as reported from the participants see Table \@ref(tab:common-comb).

<!--
The values are $\log_2$ converted to minimizes effect of high expense differences between treatment methods (Fig.&nbsp;\@ref(fig:estimates-bland)). The $\log_2$ change can be seen as follows, one represent double of the original and minus one is half of the original on the transformed scale.
Treatment method combinations which include 'formic acid short-term' (SU-FA-ST) are below the lower two times standard deviation limit, which means the survey responses are higher roughly ~1.7 times higher that the calculated estimates.
The $\log_2$ change can be seen as follows, one represent a double of the original and minus one is half of the original on the transformed scale. Each treatment combination is inside the 95% confidence interval of the upper and lower two times standard deviations from the mean difference. Combinations which include formic acid short-term are above the upper two times standard deviation limit, which means the survey responses are higher roughly ~1.7 times higher that the calculated estimates. The two combinations with the higher overestimating and outside of the two times standard deviation limit are both including a biotechnical treatment method in summer.
-->

(ref:estimated-bland) Difference plot between survey and estimated expenses of treatment method combinations with at least 15 beekeepers using it, over the three years of survey. Negative values on the vertical axis are underestimates and positive values are overestimates of the estimated expenses. Lines indicating mean difference (yellow) and lower and upper two times standard deviation (blue) each with 95% confidence interval. Point size is the sample count of each treatment method combination. Combinations inside the two times standard deviations CI range and the top 3 combinations based on number of participants are colour coded and labelled. SP/SU/WI = Spring/Summer/Winter || AS-LT/ST = Formic acid - long/short term || Ox-sub/pure/mix = Oxalic acid - sublimation, pure or mixture || Biot = Another biotechnical methods.

```{r estimates-bland, include=T, fig.cap="(ref:estimated-bland)", out.width="100%"}
include_custom("output/figs/bland-altman.png")
```

```{r common-comb, include=T}
c <- "Top 11 most used treatment method combination for all three survey years. Total and Percent correspond to participants. Mean and median expenses from the survey and in comparison to the calculated estimate in Euro."
cn <- c("Method", "Total [n]", "Percent [%]", "Mean", "Median", "Mean", "Median")
kable(
    r_estimate$data %>% dplyr::slice_max(n, n = 10) %>% select(c_short_od:med_estimate),
    booktabs = TRUE,
    row.names = FALSE,
    digits = 1,
    col.names = cn,
    caption = c,
    linesep = c("", "", "", "", "\\addlinespace"),
) %>%
    add_header_above(c(" " = 3, "Survey" = 2, "Estimate" = 2)) %>%
    footnote(threeparttable = T, general = "SP/SU/WI = Spring/Summer/Winter || AS-LT/ST = Formic acid - long/short term || Ox-sub/pure/mix = Oxalic acid - sublimation, pure or mixture || Biot = Another biotechnical methods.") %>%
    kable_styling(font_size = 8)
rm(c, cn, d)
```


<!--chapter:end:03-results-3-estimate.Rmd-->

## Expenses Extrapolation

```{r, include=FALSE}
if (file.exists("output/r-extrapolation.rds")) {
    r_extrapolation <- readRDS("output/r-extrapolation.rds")
} else {
    source("src/r-extrapolation.R", local = knitr::knit_global())
    saveRDS(r_extrapolation, "output/r-extrapolation.rds")
}
```

The rough extrapolation of the survey means expenses with the total sum of honeybee colonies in Austria are based on the reported numbers by the national beekeeping association 'Biene Österreich'[^r-vis].
This extrapolation with the arithmetic mean sums up in
2018/19 to EUR&nbsp;`r format(r_extrapolation$data %>% pull(extrapolation_estimate)%>% .[1], big.mark = ",")` (95% CI: `r format(r_extrapolation$data %>% pull(extrapolation_estimate_lower)%>% .[1], big.mark = ",")` - `r format(r_extrapolation$data %>% pull(extrapolation_estimate_upper)%>% .[1], big.mark = ",")`),
2019/20 to EUR&nbsp;`r format(r_extrapolation$data %>% pull(extrapolation_estimate)%>% .[2], big.mark = ",")` (95% CI: `r format(r_extrapolation$data %>% pull(extrapolation_estimate_lower)%>% .[2], big.mark = ",")` - `r format(r_extrapolation$data %>% pull(extrapolation_estimate_upper)%>% .[2], big.mark = ",")`) and
2020/21 to EUR&nbsp;`r format(r_extrapolation$data %>% pull(extrapolation_estimate)%>% .[3], big.mark = ",")` (95% CI: `r format(r_extrapolation$data %>% pull(extrapolation_estimate_lower)%>% .[3], big.mark = ",")` - `r format(r_extrapolation$data %>% pull(extrapolation_estimate_upper)%>% .[3], big.mark = ",")`) (Table&nbsp;\@ref(tab:extrapolation-tab), Fig. \@ref(fig:extrapolation)).

In addition the same extrapolation was calculated with the geometric mean from the natural log. For this calculation the single zero value in the dataset was replaced with EUR&nbsp;0.01. The geometric mean is less influenced by outliers and typically used as point estimate for a log-normal distribution. The results are always lower than the arithmetic mean and the difference in the extrapolation is on average EUR&nbsp;1,258,377.

Over the three survey years the extrapolation with the geometric mean sums up in 2018/19 to `r format(r_extrapolation$data %>% pull(extrapolation_estimate)%>% .[4], big.mark = ",")` (95% CI: `r format(r_extrapolation$data %>% pull(extrapolation_estimate_lower) %>% .[4], big.mark = ",")` - `r format(r_extrapolation$data %>% pull(extrapolation_estimate_upper)%>% .[4], big.mark = ",")`),
2019/20 to `r format(r_extrapolation$data %>% pull(extrapolation_estimate) %>% .[5], big.mark = ",")` (95% CI: `r format(r_extrapolation$data %>% pull(extrapolation_estimate_lower)%>% .[5], big.mark = ",")` - `r format(r_extrapolation$data %>% pull(extrapolation_estimate_upper)%>% .[5], big.mark = ",")`) and
2020/21 to `r format(r_extrapolation$data %>% pull(extrapolation_estimate) %>% .[6], big.mark = ",")` (95% CI: `r format(r_extrapolation$data %>% pull(extrapolation_estimate_lower)%>% .[6], big.mark = ",")` - `r format(r_extrapolation$data %>% pull(extrapolation_estimate_upper)%>% .[6], big.mark = ",")`) (Table&nbsp;\@ref(tab:extrapolation-tab), Fig. \@ref(fig:extrapolation)). In the survey interpolation these two methods are compared to the total sum as is in fact seen in the survey, where the total expenses for each participant was calculated by the reported expenses multiplied by colonies going into winter. It showed that the real survey sum is in-between the arithmetic and geometric mean (Table&nbsp;\@ref(tab:extrapolation-tab) - 'Real Sum [EUR]').


```{r extrapolation-tab, include=T}
c <- "Multiplication of the average treatment expenses per colony with the total number of honeybee colonies wintered from the survey participants and extrapolation for Austria with the number of colonies as reported from the national beekeeping association 'Biene Österreich'. Column 'Real Sum' is the observed total sum from the survey, where the total expenses for each participant was calculated by reported expenses multiplied by colonies going into winter and afterwards summarised for each year. Confidence intervals for point estimates are calculated with 1,000 bootstrap resamples. Two methods applied arithmetic mean and geometric mean."
cn <- c("Year", "Mean [EUR]", "CI Lower", "CI Higher", "Colonies [n]", "Mean Sum [EUR]", "Real Sum [EUR]", "Colonies [n]", "Mean Sum [EUR]")
kable(
    r_extrapolation$data %>%
        select(year, statistic, conf.low, conf.high, colonies_survey, survey_estimate, survey_real, colonies, extrapolation_estimate) %>%
        mutate(
            across(statistic:conf.high, round, digits = 2),
            across(colonies_survey:extrapolation_estimate, ~ .x %>%
                round() %>%
                ft())
        ),
    align = c("c", rep("r", 7)),
    booktabs = TRUE,
    row.names = FALSE,
    col.names = cn,
    caption = c,
    escape = TRUE,
) %>%
    add_header_above(c(" " = 4, "Survey" = 3, "Extrapolation" = 2)) %>%
    pack_rows("Arithmetic Mean", 1, 3) %>%
    pack_rows("Geometric Mean", 4, 6) %>%
    kable_styling(latex_options = "scale_down")
rm(c, cn)
```

(ref:extrapolation) Multiplication of the average treatment expenses per colony with the total number of honeybee colonies wintered from the survey participants and extrapolation for Austria with the number of colonies as reported from the national beekeeping association 'Biene Österreich'. Confidence intervals for point estimates are calculated with 1,000 bootstrap resamples. Two methods applied arithmetic mean and geometric mean. Basis of the calculation are the mean expenses for one colony, see Table&nbsp;\@ref(tab:extrapolation-tab).
```{r extrapolation, include=T, fig.cap="(ref:extrapolation)", out.width="100%"}
include_custom("output/figs/extrapolation.png")
```

<!--chapter:end:03-results-4-extrapolation.Rmd-->

## Operational Factor Analysis {#two-level}
```{r, include=FALSE}
if (file.exists("output/r-operational.rds")) {
    r_operational <- readRDS("output/r-operational.rds")
} else {
    source("src/r-operational.R", local = knitr::knit_global())
}
```

The separate analysis of each of the three selected operational questions are shown in following factor analysis with two levels each for each year. Interpretation of the results need to factor in possible interactions and the different cohort sizes in the survey [Section \@ref(operational-survey)].

### Operation Size {#operational-size-single}

The number of honeybee colonies managed by participants follow a heavy tailed count distribution (Mean = 21-23 Colonies, Median = 11 Colonies), with most participants managing less than 20 colonies and a few extensive operation with over one hundred colonies (Fig.&nbsp;\@ref(fig:size-distr)). The participants were grouped into two levels for factor analysis. The 'Professional Beekeeper' with more than 25 colonies (>25) and beekeepers with 25 or less colonies (<=25) as 'Recreational Beekeeper'. Over the survey years the 'Recreational Beekeeper' made up the majority of the survey responses with 78% in 2018/19, 80% in 2019/20 and 76% in 2020/21 (Table&nbsp;\@ref(tab:size-table)).

Comparing these two levels and the reported expenses per colony on treatment, shows that the 'Recreational Beekeeper' spend on average and in median more per colony as the 'Professional Beekeeper'. The absolute median difference between the groups is between EUR&nbsp;`r min(r_operational$size$statistic_perm$point_estimate %>% unlist())` and EUR&nbsp;`r max(r_operational$size$statistic_perm$point_estimate %>% unlist())` per colony. Inference based statistic based on the median difference with permutation test (R = 5,000) shows significant difference in all years (*p* < 0.01) and 95% CI (bootstrap, R = 1,000; 2018/19: EUR&nbsp;4.1-5.9; 2019/20: EUR&nbsp;2.6-5.4; 2020/21: EUR&nbsp;2.6-5.4) is not overlapping with zero difference in any survey year (Fig.&nbsp;\@ref(fig:size-stats)).

(ref:size-distr) Distribution of operation size/number of colonies over the three survey years. Numbers are based on the reported colonies going into winter, as reported by survey participants. Different years are colour coded (orange = 2018/19, green = 2019/20, blue = 2020/21). Total number of participants and colonies are reported in each years plot header.

```{r size-distr, include=T, fig.cap="(ref:size-distr)", out.width="100%"}
include_custom("output/figs/size-distr.png")
```

```{r size-table, include=T}
c <- "Operation size / number of colonies distribution and measures of central tendency over the three survey years. Operation size with 25 or less colonies ('<=25') are 'Recreational Beekeeper' and with more are 'Professional Beekeeper' ('>25')."
cn <- c("Operation Size", "Beekeeper [n]", "Colonies [n]", "Mean [EUR]", "Median [EUR]")
tab <- kable(
    r_operational$size$table %>%
        select(-year) %>%
        mutate(
            colonies = ft(colonies)
        ),
    booktabs = TRUE,
    col.names = cn,
    digits = 1,
    align = c("c", rep("r", 4)),
    caption = c
) %>%
    pack_rows("2018/19", 1, 2) %>%
    pack_rows("2019/20", 3, 4) %>%
    pack_rows("2020/21", 5, 6) %>%
    kable_styling(latex_options = "scale_down")
tab
rm(cn, c, tab)
```

(ref:size-stats) Boxplot with median and interquantile range (IQR, $25^{th}$ - $75^{th}$ quantile) and vertical lines indicate 1.5\*IQR. Pink rhombus represents the arithmetic mean. Pairwise difference of the mean between groups are given as labels above the boxplots. Gray points are the actual data points. Reported inference statistic results on median difference between the given operation size groups, with *p*-values from permutation test (R = 5,000) and 95% CI (bootstrap, R = 1,000). Exact *p*-values and test results are available in the Appendix Fig.&nbsp;\@ref(fig:size-permutation). Survey answers above EUR 80 per colony are not shown.  

```{r size-stats, include=T, fig.cap="(ref:size-stats)", out.width="100%"}
include_custom("output/figs/size-stats.png")
```

### Certified Organic Beekeeper

The survey question 'Certified Organic Beekeeper' did include as possible answers 'Yes', 'No' and 'Uncertain', whereas 'Uncertain' and participants which did not answer to this question were ignored in the statistical two levels factor analysis. The majority of the survey participants is uncertified (`r r_operational$organic$summary$np[[2]]`%) and `r r_operational$organic$summary$np[[1]]`% report to be certified organic. Only `r r_operational$organic$summary$np[[3]]`% did answer with 'Uncertain' and `r r_operational$organic$summary$np[[4]]`% did answer not at all (Table&nbsp;\@ref(tab:organic-table)).

Certified organic beekeeping operation spend on average and in median less per colony on treatment expenses. The absolute median difference between the groups is between EUR&nbsp;`r min(r_operational$organic$statistic_perm$point_estimate %>% unlist())` and EUR&nbsp;`r max(r_operational$organic$statistic_perm$point_estimate %>% unlist())` per colony. Inference based statistic based on the median difference with permutation test (R = 5,000) shows significant difference in all years (2018/19 & 2019/20: *p* < 0.01; 2020/21: *p* = 0.03 ) and 95% CI (bootstrap, R = 1,000; 2018/19: 2.5-5.9; 2019/20: 1.9-6.1; 2020/21: EUR&nbsp;2.2-5.8) is not overlapping with zero difference in any survey year (Fig.&nbsp;\@ref(fig:organic-stats)).

```{r organic-table, include=T}
c <- "Distribution of answers on questions 'Certified Organic Beekeeper' and measures of central tendency over the three survey years."
cn <- c("Certified Organic", "Beekeeper [n]", "Colonies [n]", "Mean [EUR]", "Median [EUR]")
tab <- kable(
    r_operational$organic$table %>%
        select(-year) %>%
        mutate(
            colonies = ft(colonies)
        ),
    booktabs = TRUE,
    col.names = cn,
    digits = 1,
    align = c("l", rep("r", 4)),
    caption = c
) %>%
    pack_rows("2018/19", 1, 4) %>%
    pack_rows("2019/20", 5, 8) %>%
    pack_rows("2020/21", 9, 12) %>%
    kable_styling(latex_options = "scale_down")
tab
rm(cn, c, tab)
```

(ref:organic-stats) Boxplot with median and interquantile range (IQR, $25^{th}$ - $75^{th}$ quantile) and vertical lines indicate 1.5\*IQR. Pink rhombus represents the arithmetic mean. Pairwise difference of the mean between groups are given as labels above the boxplots. Gray points are the actual data points. Reported inference statistic results on median difference between the two level factor analysis if 'Organic Certified Beekeeper' or not, with *p*-values from permutation test (R = 5,000) and 95% CI (bootstrap, R = 1,000). Exact *p*-values and test results are available in the Appendix Fig.&nbsp;\@ref(fig:organic-permutation). Survey answers above EUR 80 per colony are not shown.  

```{r organic-stats, include=T, fig.cap="(ref:organic-stats)", out.width="100%"}
include_custom("output/figs/organic-stats.png")
```

### Migratory Beekeeper

The beekeepers participating in our study were asked whether they transported their bees to sources of honey yields or for pollination service (this excludes movements in the course of breeding or splits). The question 'Migratory Beekeeper' did include as possible answers 'Yes', 'No' and 'Uncertain', whereas 'Uncertain' answers were ignored in the statistical two levels factor analysis. A total of `r r_operational$migratory$summary$np[[1]]`% of the survey participants are migrating with their honeybee colonies but the majority is not migrating (`r r_operational$migratory$summary$np[[2]]`%). Only `r r_operational$migratory$summary$np[[3]]`% did answer 'Unsure' on this question (Table&nbsp;\@ref(tab:migratory-table)).

Migratory beekeeping operations spend on average and in the median less per colony on treatment expenses in comparison to survey participants who do not migrate. The absolute median difference between the groups is between EUR&nbsp;`r min(r_operational$migratory$statistic_perm$point_estimate %>% unlist())` and EUR&nbsp;`r max(r_operational$migratory$statistic_perm$point_estimate %>% unlist())` per colony. Inference based statistic based on the median difference with permutation test (R = 5,000) shows significant difference in all years (2018/19: *p* = 0.05; 2019/20: *p* < 0.01, 2020/21: *p* = 0.02 ) and 95% CI (bootstrap, R = 1,000; 2018/19: 0.4-5.6; 2019/20: 2.2-5.8; 2020/21: EUR&nbsp;1.1-5.9) is not overlapping with zero difference in any survey year (Fig.&nbsp;\@ref(fig:migratory-stats)).

```{r migratory-table, include=T}
c <- "Distribution of answers on questions 'Migratory Beekeeper' and measures of central tendency over the three survey years."
cn <- c("Migratory Beekeeper", "Beekeeper [n]", "Colonies [n]", "Mean [EUR]", "Median [EUR]")
tab <- kable(
    r_operational$migratory$table %>%
        select(-year) %>%
        mutate(
            beekeeper = ft(beekeeper),
            colonies = ft(colonies)
        ),
    booktabs = TRUE,
    col.names = cn,
    digits = 1,
    align = c("l", rep("r", 4)),
    caption = c
) %>%
    pack_rows("2018/19", 1, 3) %>%
    pack_rows("2019/20", 4, 6) %>%
    pack_rows("2020/21", 7, 9) %>%
    kable_styling(latex_options = "scale_down")
tab
rm(cn, c, tab)
```

(ref:migratory-stats) Boxplot with median and interquantile range (IQR, $25^{th}$ - $75^{th}$ quantile) and vertical lines indicate 1.5\*IQR. Pink rhombus represents the arithmetic mean. Pairwise difference of the mean between groups are given as labels above the boxplots. Gray points are the actual data points. Reported inference statistic results on median difference between the two level factor analysis if 'Migratory Beekeeper' or not, with *p*-values from permutation test (R = 5,000) and 95% CI (bootstrap, R = 1,000). Exact *p*-values and test results are available in the Appendix Fig.&nbsp;\@ref(fig:organic-permutation). Survey answers above EUR 80 per colony are not shown. 

```{r migratory-stats, include=T, fig.cap="(ref:migratory-stats)", out.width="100%"}
include_custom("output/figs/migratory-stats.png")
```

<!--chapter:end:03-results-5-single-factor.Rmd-->


## Multiple Regression

To better understand if the collected survey data can be used as predictor variables and the explanatory power a linear model with treatment expenses per colony as target was used. Starting with the base model including only the reported number of colonies wintered (numerical, $\log_{10}$ transformed), which has demonstrated the most notable difference in the two level factor analysis between groups [Section \@ref(operational-size-single)]. Adding the two other operational questions 'Migratory Beekeeper' and 'Certified Organic Beekeeper' (Two levels: 'Yes', 'No') did lead to no noteworthy improvement and resulting in a higher BIC value. The base model plus the treatment methods (binary encoded) was the best model ($R^2$ 0.127), as it showed the lowest AIC (`r round(r_model$fitted$AIC[5])`) and BIC (`r round(r_model$fitted$BIC[5])`) value and $R^2$ of 0.127. Adding states as additional covariate did improve the $R^2$ value (0.131) but showed higher AIC and BIC, indicating the penalization of the additional factors is higher than the gain and could lead to possible overfitting. The same is true for the addition of 'Survey Year', which did merely increase the $R^2$ value (0.128) but again showed higher AIC and BIC values (Table&nbsp;\@ref(tab:model-table)).

```{r model-table, include=T}
c <- "Summary of model fit. Base model only included number of colonies wintered. The two questions migratory beekeeper and certified organic beekeeper were added as two level factor ('Yes', 'No'). The eleven treatment methods are binary encoded, excluded is drone brood removal. The covariates 'States' and 'Survey Year' were dummy encoded. Selection of best model, indicated as bold font, based on AIC and BIC value."
cn <- c("Model", "$R^2$", "adj. $R^2$", "df", "logLik", "AIC", "BIC")
tab <- kbl(
    r_model$fitted %>%
        mutate(
            model = c("Intercept Only", "Base (Operation Size)", "Base $+$ Certified Organic", "Base $+$ Migratory Beekeeper", "Base $+$ Treatment Methods", "Base $+$ Treatment Methods $+$ States", "Base $+$ Treatment Methods $+$ Year"),
            df = ifelse(is.na(df), 0, df),
            across(logLik:BIC, ~ .x %>%
                round() %>%
                ft()),
        ) %>%
        select(model, r.squared, adj.r.squared, df:BIC),
    booktabs = TRUE,
    col.names = cn,
    digits = 3,
    linesep = "",
    align = c("l", rep("r", 9)),
    caption = c,
    escape = FALSE
) %>%
    kableExtra::row_spec(5, bold = TRUE) %>%
    kable_styling(latex_options = "scale_down")
tab
rm(cn, c, tab)
```

The residuals of the best model fit ('Base $+$ Treatment Methods') showed a normal distribution (Appendix Fig. \@ref(fig:model-qq-residuals)). The geometric mean intercept of the best model, which would represent a hypothetical beekeeper with one colony ($\log_{10} 1$) and no applied treatments is at EUR&nbsp;`r round(r_model$coeff_intercept$estimate_conv, 2)` (SE: `r round(r_model$coeff_intercept$error_conv, 2)`). All predictors are statistically significant except for the treatment method 'Lactic acid' (*p* = 0.64), which is used by 3% of the survey participants (*n*=117) (Table&nbsp;\@ref(tab:model-best)). The coefficient direction is increasing for most of the treatment methods if they are applied. Only two treatment methods are decreasing the intercept if applied, 'Lactic acid' with standard error crossing zero difference and 'Another biotechnical method' which decreases the estimated expenses by 9.5% (SE: 3.6%). The decrease for 'Number of colonies' is different to interpret as in this case both dependent and independent variables are log transformed, therefore the increase of colonies by 10% leads to a mean decrease of expenses by 1.7% (SE: 0.1%) (Fig.&nbsp;\@ref(fig:model-whisker)).


The predictive ability of the selected best model ('Base $+$ Treatment Methods') is mediocre with a relative accuracy based on correlation between observed and predicted in the training set of `r round(r_model$performance_accuracy_best$Accuracy*100, 2)`% (SE: `r round(r_model$performance_accuracy_best$SE*100)`%). The root mean square error (RMSE) for the training dataset (RMSE=`r r_model$prediction_stats %>% filter(type == "Training") %>% pull(.estimate) %>% round(., 2)`) and testing set (RMSE=`r r_model$prediction_stats %>% filter(type == "Testing") %>% pull(.estimate) %>% round(., 2)`) are similar, indicating low probability of overfitting (Fig.&nbsp;\@ref(fig:model-pred)).

<!--
https://stats.stackexchange.com/questions/18480/interpretation-of-log-transformed-predictor-and-or-response
https://data.library.virginia.edu/interpreting-log-transformations-in-a-linear-model/
# Old one if we use geometric mean:
# I write geometric mean because back transformation of log transformed data is actually the same as the geometric mean
# abstract: https://www.researchgate.net/publication/5402553_The_logarithmic_transformation_and_the_geometric_mean_in_reporting_experimental_IgE_results_What_are_they_and_when_and_why_to_use_them
-->

```{r model-best, include=T}
c <- "Summary of the coefficients of the selected best model and statistic results. Terms starting with 'T.' are treatment methods. Significant stars are labelled based on the \\textit{p}-value (*** < 0.001, ** < 0.01, * <= 0.05, n.s. (not significant) > 0.5). SE = Standard Error."
cn <- c("Term", "Estimate", "SE", "Statistic (t)", "\\textit{p}", "")
tab <- kable(
    r_model$coeff %>%
        mutate(
            tname = str_replace(tname, "T.", "T."),
            p.star = case_when(
                p.value < 0.001 ~ "***",
                p.value < 0.01 ~ "**",
                p.value <= 0.5 ~ "*",
                TRUE ~ "n.s."
            ),
            p.value = ifelse(
                p.value > 0.5,
                base::format(round(p.value, 2), digits = 3, scientific = FALSE, nsmall = 1),
                base::format(p.value, digits = 3, scientific = TRUE, nsmall = 1)
            ),
        ) %>%
        select(tname, estimate:p.value, p.star) %>%
        arrange(tname),
    booktabs = TRUE,
    col.names = cn,
    digits = 3,
    escape = FALSE,
    align = c("l", rep("r", 4), "c"),
    caption = c
) %>%
    kable_styling(latex_options = "scale_down")
tab
rm(cn, c, tab)
```

(ref:model-whisker) Dot-and-whisker plot for the best model ('Base $+$ Treatment Methods') regression coefficients. The coefficients are ordered by the mean estimate value and direction, with highest positive at the top and lowest negative at the bottom. Colour coded if regression direction is negative (green) or positive (red). Coefficients starting with 'T.' are treatment methods. The values are plotted to the mean increase or decrease of the estimate in percent, which means for the treatment methods if applied they increase/decrease the value of the target by shown percentage. In the case of 'Number of colonies' [$\log_{10}$] the effect of percentage change on the expenses is shown if the number of colonies increases by 10%.

```{r model-whisker, include=T, fig.cap="(ref:model-whisker)", out.width="100%"}
include_custom("output/figs/model-whisker.png")
```

```{r model-pred, include=T, fig.cap="Model prediction of expenses on training and testing dataset to the observed data from the survey. Orange and blue lines are the linear fit with standard error of the intercept only model without any predictors and selected best model ('Base $+$ Treatment Methods'). Red line represents the theoretical perfect fit between estimate and observed data.", out.width="100%"}
include_custom("output/figs/model-pred.png")
```

<!--chapter:end:03-results-6-model.Rmd-->

## Treatment Network {#network}

```{r}
source("src/r-treatment-network.R", local = knitr::knit_global())
```

To better understand which treatment methods are actually commonly reported a adjacency matrix of all treatment methods with at least 30 participants reporting it over all three survey years was generated. The follow up network analysis of the treatment methods over all three survey years showed the dominance of oxalic treatment methods in winter. In spring the biomechinal method drone brood removal (*n*=1,833) is the most dominant reported varroa treatment and very often in combination with the summer variant (*n*=1,638, combination=1,138). In summer the formic acid treatment variants are the most common (long=2,004, short=1,639) followed by the oxalic acid treatments (sublimation=994, pure trickling or spraying=582, ready-made mixture trickling=560). The treatment method 'Another biotechnical Method' (without drone brood removal or hyperthermia) is in summer also frequently reported with *n*=950. The other treatment methods in summer with lower number of participants are 'Thymol' (*n*=328), 'Hyperthermia' (*n*=160), 'Lactic acid' (*n*=92) and 'Synthetic Methods' (*n*=86) (Fig.&nbsp;\@ref(fig:treatment-network)).

(ref:treatment-network) Network graph of the treatment methods over all three survey years. Node sizes are the number the method were reported. Node colours are the seasons 'Spring' (April-May), 'Summer' (June-October) and 'Winter' (November - January). Edge connections colour and width are the reported amount of each combination.  Abbrevations: Drone = Drone brood removal || AS-LT/ST = Formic acid - long/short term || Ox-sub/pure/mix = Oxalic acid - sublimation, pure or mixture || Biot = Another biotechnical methods || Thy = Thymol || Hyp = Hyperthermia || syn Met = synthetic Methods. Categories with less than 30 reported answers were not included.

```{r treatment-network, include=T, fig.cap="(ref:treatment-network)"}
include_custom("output/figs/treatment-network.png")
```

<!--chapter:end:03-results-7-network.Rmd-->

\newpage
## Efficacy & Economical
```{r, include=FALSE}
if (file.exists("output/r-treatment.rds")) {
    r_treatment <- readRDS("output/r-treatment.rds")
} else {
    source("src/r-treatment.R", local = knitr::knit_global())
    saveRDS(r_treatment, "output/r-treatment.rds")
}
```

The total observed loss rate over the three survey years was in 2018/19 `r round(r_treatment$year_loss$year_loss[1], 1)`% (95% CI: `r round(r_treatment$year_loss$loss_lower_ci[1], 1)`-`r round(r_treatment$year_loss$loss_upper_ci[1], 1)`%), in 2019/20 `r round(r_treatment$year_loss$year_loss[2], 1)`% (95% CI: `r round(r_treatment$year_loss$loss_lower_ci[2], 1)`-`r round(r_treatment$year_loss$loss_upper_ci[2], 1)`%) and in 2020/21 `r round(r_treatment$year_loss$year_loss[3], 1)`% (95% CI: `r round(r_treatment$year_loss$loss_lower_ci[3], 1)`-`r round(r_treatment$year_loss$loss_upper_ci[3], 1)`%) (Fig.&nbsp;\@ref(fig:efficient-economic), blue dashed vertical lines). Over the three survey years there were 17 different treatment method combinations (without drone brood removal) which had in sum at least 30 participants employing it (Table&nbsp;\@ref(tab:ee-table)). The most common treatment combination is long term formic acid in summer followed by oxalic acid trickling in winter (**A** 'SU-Fa-LT & WI-Ox-pure'). It showed in each year on mean a lower loss rate compared to the total loss rate but only in one year a low expenses category. In comparison the application of short term formic acid treatment in summer and oxalic acid sublimation in summer and winter (**K** 'SU-Fa-ST & SU/WI-Ox-Sub') was observed with a low loss rate category in each year and low expense category in two out of three years. A total of three combinations showed in all three years in mean a higher loss rate than the yearly total loss rate, which are formic acid short term in summer followed by oxalic acid trickling pure or ready-made mixture in winter (**E** 'SU-Fa-ST & WI-Ox-pure'; **L** 'SU-Fa-ST & WI-Ox-mix') and the sole application of formic acid long term method in summer, without winter treatment, which also showed high expenses in each year (**M** 'SU-Fa-LT'). The three combinations with biotechnical method (without drone brood removal or hyperthermia) in summer showed low expenses in each year (**G** 'SU-Biot. & SU-Ox-sub & WI-Ox-sub'; **P** 'SU-Biot. & SU-Fa-LT & WI-Ox-pure'; **Q** 'SU-Biot. & SU-Fa-ST & WI-Ox-pure'). Additionally the treatment method of oxalic acid sublimation in summer and winter resulted in low mean expenses each year (**F** 'SU-Ox-sub & WI-Ox-sub') (Fig.&nbsp;\@ref(fig:efficient-economic)).

(ref:efficient-economic) Efficiency and economical plot for treatment method combinations with at least *n* = 30 as sum over the three survey years. Letters corresponding to different treatment methods, see Table \@ref(tab:ee-table). Vertical axis the mean expenses for each treatment combination with standard error (SE). Horizontal axis the mean observed winter colony loss rates for each treatment combination with 95% confidence interval (CI). The blue dashed line indicating means total expenses (horizontal) and total observed winter colony loss rates (vertical) per year. Categorizing of treatment combinations based on location in relation to the blue dashed lines (Fig.&nbsp;\@ref(fig:efficient-economic-mm)). The four categories are colour coded as seen in the plot legend.

```{r efficient-economic, include=T, fig.cap="(ref:efficient-economic)"}
include_custom("output/figs/efficient-economic.png")
```

(ref:ee-table) Treatment combinations with at least *n*=30 as sum over the three survey years. Ordered by number of participants applying the combinations. 'Efficient' and 'Economical' are colour coded based on low (teal) and high (orange), categorized based to total loss rates and total mean expenses (Fig.&nbsp;\@ref(fig:efficient-economic)). Abbrevations: SP/SU/WI = Spring/Summer/Winter || AS-LT/ST = Formic acid - long/short term || Ox-sub/pure/mix = Oxalic acid - sublimation, pure or mixture || Biot = Another biotechnical methods.

```{r ee-table, include=T}
cn <- c(" ", "Treatment", "Year", "\\textit{n}", "Median", "Mean", "SD", "Mean", "Lower CI", "Upper CI", "Efficient", "Economical")
cell_colors_price <- c("high-expense" = "cred!80", "low-expense" = "cgreen!80")
cell_colors_loss <- c("high-loss" = "cred!80", "low-loss" = "cgreen!80")

tab <- kbl(
    r_treatment$tab %>%
        mutate(
            c_short_od = str_replace_all(c_short_od, "&", " \\\\& "),
            across(letter:c_short_od, ~ ifelse(lag(as.character(.x), default = "0") == .x, "", .x)),
            across(observed_costs:observed_costs_sd, ~ ifelse(lag(.x, default = 0) == .x, "", round(.x, 2)))
        ) %>%
        mutate(
            letter = cell_spec(letter, bold = TRUE),
            efficient = cell_spec(
                efficient,
                background = cell_colors_loss[efficient]
            ),
            economical = cell_spec(
                economical,
                background = cell_colors_price[economical]
            )
        ),
    linesep = c("", "", "\\addlinespace"),
    booktabs = TRUE,
    col.names = cn,
    digits = 1,
    align = c("c", "l", "c", rep("r", 6), "c", "c"),
    caption = "(ref:ee-table)",
    escape = FALSE
) %>%
    add_header_above(c(" " = 4, "Expenses [EUR]" = 3, "Observed Loss Rate [\\\\%]" = 3, " " = 2)) %>%
    # landscape()
    kable_styling(latex_option = "scale_down")
tab
rm(cn, c, tab)
```

<!--chapter:end:03-results-8-treatment.Rmd-->

# Discussion

## Sample Representative and Limitations

Although the popularity and rise of new beekeepers in Austria, a decrease in absolute and relative numbers of participants over the three survey years was observed (Table&nbsp;\@ref(tab:registered)). I would have expected the opposite because in 2016 new rules made it obligatory for beekeepers who want to participate in the bee health program to fill out the yearly winter loss survey. Beekeeper who are part of the bee health programme are allowed to apply and/or receive a higher rate of government aid for investments in technical equipment to improve production quality and hygiene in their beekeeping operation.^[AgrarMarkt Austria, 06/2021 - Version 04] Before this indirect monetary incentive there were no advantages or personal rewards for participants in the voluntary survey. Incentives showed in an analysis of two meta-studies an increase of survey responses and also a lower drop-out rate [@goritz2006]. However, the effect or higher motivation due to incentives is comparatively small [@goritz2006;@bruggen2011]. One of the strong major motivations for beekeeper to participate in citizens science ecological field studies is to be able to contribute to the scientific knowledge [@gratzer2021]. The high rate of participants with intrinsically motivation was also identified for online surveys [@bruggen2011]. The winter colony loss survey cannot be described purely as a citizens science study or an ordinary online survey and is likely an in-between. Nevertheless, to increase and improve the retention of beekeepers participating the project should strive for better communication on how it improves the beekeeping knowledge and identify ways to acknowledge each personal contribution [@nov2014]. One idea would be a personalized automated report of the survey results, were the independent results are compared to the total observed results. This could be perceived as an acknowledge service in the form of a free professional data analysis for the volunteers. Which is one of the motivations for citizens scientists [@gratzer2021].

As with the majority of monitoring survey studies, the design is subject to limitations especially in their representativeness. From a statistical standpoint, we cannot say our sample represents the beekeeper population in Austria as the sample was not selected at random from the population [@vanderzee2013]. In Austria all beekeepers must officially register in VIS[^r-vis], but access to this data is limited and only available for prevention of diseases. A random stratified selection out of all beekeepers would be an ideal situation, which would help to prevent potential selection biases and survey errors [@heeringa2010;@vanderzee2013]. With this in mind point estimates and results should be viewed with caution. The participating beekeepers in the survey are from close to all country districts, indicating a thorough spatial coverage, and a response rate of 3.5-4.0% summarises to around 1.000 participants each year. Based on the rules of thumbs from @roscoe1975, a multivariate analyses should typically comprise 10% more responses than questions and cover all answering groups with an appropriate number of answers (*n*=30). In addition for permutation tests, which were used in the two level factor analysis (Section \@ref(two-level)), the combination of skewed data and unequal sample sizes could lead to possible false inferences [@christensen2021]. The problem of low sample sizes in certain groups can somewhat be opposed with the data collection over three years. Results can be viewed as more robust, if the same results were observed in each year, or weak, if this is not the case. For the second concern a rank based permutation test was performed with bootstrap confidence interval to not merely rely on a single point estimate and in addition it allows to visualize the test results.

Methodically a lot of care was undertaken to identify outliers and imputation of some answers. For good practice and to minimize arbitrary decisions different statistical methods were applied and my knowledge about the different expenses of various treatment methods was taken into account [@kwak2017]. However, this imputation, which can be seen as an act of data manipulation, should not be taken too lightly. With my methodical transparent approach and the public availability of the source code I try to take care of possible arguments of manipulative intentions. 

The reported inference based statistics are without correction of the alpha error, since different hypotheses are tested in the questionnaire as it is mostly explorative and possible overlaps of the hypotheses are difficult to classify. The advantage of the COLOSS questionnaire is that the questions are the same over several years. This is typically because of interactions with temporal effects for example the weather which has shown to impact the winter colony survival [@switanek2017;@becsi2021]. The repetition additionally helps to reduce the probability of mistaken inferences in statistics, if consistent results can be observed over multiple years. The questionnaire itself can be seen as a tool to identify possible risk factors and compare the results in a standardised way with other countries part of COLOSS [@brodschneider2016;@brodschneider2018]. The identified risk factors can then be followed up by research with specific experimental design, for example the observed crippled wings by beekeepers with a study about the viral load [@morawetz2019].

Another point which may represent a limitation is the survey data quality. Especially for the question about estimated treatment expenses per colony without labor costs. The data distribution showed participants mostly answered in full round numbers, eg. 5 or 10 Euro. This rounding leads to less variance between groups. Beekeepers answering on this question without ordinary estimates would need to keep books on their expenses or invest more effort when answering the survey, to find out how much they actually spend. I believe this is not the case for the majority, as most participants presumably regard beekeeping as a hobby and do not keep track of their expenses. Additionally, the position of the question in the questionnaire could cause a negative effect as it placed at the end of the survey. Although the structure did not seem to portray a major role on data quality as shown in a postal survey [@dunn2003].

## Motivation choosing Varroa Control Method

The motivation why participants practise a specific method was a novel question and was only asked in the last survey year. 'Efficacy' (67.3%) which was picked by the majority, as one of the categories, is the most relevant factor. The obvious follow up question would be, how do the participant know the chosen control method is efficient. In Austria the treatment methods are dominated by organic acids and the loss rates for these methods are similar [@oberreiter2020]. The majority follows the official recommendation of formic acid after honey harvest in summer and oxalic acid treatment in winter [@moosbeckhofer2015]. In comparison to Czech, which showed in four successive winters lower loss rates than Austria, synthetic acaricides are common (eg. Amitraz) and can even be commanded by the authorities if the varroa infestation level is unacceptably high [@brodschneider2019a]. Even though synthetic acaricides show a specific and effectiveness against the varroa mite, residues and resistance remain a problem [@rosenkranz2010;@noel2020;@kast2021]. At least the problem of resistances with synthetic acaricides can be minimized by employing different methods in each year [@noel2020]. On this note, the motivation category 'Residues' (42.9%) is the third most picked category in the survey, showing the survey participants are aware of the problem, especially the bigger beekeeping operations. 'Resistances' was selected by 17.1% and seem for most not a primary motivation for choosing their varroa control method. The category 'Resistances' is more of a concern or motivational reason for bigger operations in Austria (28%). In Austria the high concern or possible knowledge of 'Residues' and 'Resistances' could be the reason, why compared to the treatment strategy in Czech or the common practices of sideline and commerical U.S. beekeepers, the use of synthetic acaricides is rather low [@underwood2019;@brodschneider2019a]. On the other side, that 'Residues' is a higher motivation than 'Resistances' could be because of the low uptake on synthetic acaricides and beekeepers are not as afraid of resistances because they did not experience it.

The second most picked answer in the survey represent 'Side effects on bees' (49%). Probably all treatment methods can have side effects on honeybees, but research on this topic is not available for all products. For example the accumulation of the synthetic acaricide Coumaphos in the bees wax can lead to toxic levels for larvae [@kast2020]. The organic formic acid is highly dependent on outside temperature and the fumes can damage adult bees and larvae [@bolli1993;@underwood2003]. In addition various treatment methods showed under controlled conditions negative influence on the gen expression for memory of adult honeybees [@gashout2020] and overdoses of oxalic acid trickling in winter can lead to higher bee mortality [@charriere2002;@adjlane2016].

In the survey the most common sole answer was 'Recommendations from beekeeping associations, extension workers, veterinarians, magazines, etc.' (*n*=16), which was picked in total as one of their choices by 26.1%. It is a more dominant motivation for 'Recreational Beekeepers' than for bigger operations. This means that around one-fourth in our survey are motivated by external resources in their decision making. This demonstrate the importance of knowledge transfer in the beekeeping community and of projects like B-RAP from COLOSS, which tries to improve the information flow between beekeepers and scientific results and vice versa [@kristiansen2021]. For example the communication of ways for integrated pest management (IPM) in beekeeping could be fruitful as it would be one way to measure 'Efficacy', lower 'Residues' and 'Side effects on bees' and minimize the chance of 'Resistances' [@roth2020;@bubnic2021]. Only a small amount of beekeepers are picking their treatment methods purely from 'Habit', showing that beekeepers are somewhat flexible and willing to change their methods. A U.S. national survey identified two main groups of beekeepers in their varroa management beliefs, the first one 'Treatment Skeptics', which were less likely to treat and 'Treatment Adherents' who actively try to keep bees healthy [@thoms2019]. The authors discuss further that communication needs to be adapted to these two beliefer-types and beekeepers should be made aware that their decision in management practices also influences neighbouring colonies to prevent upcoming conflicts between the groups [@thoms2019]. In our survey we can see clearly two different goals and motivations for selecting varroa treatment methods based on the operation size. Where motivations on some parts as for 'Efficacy' overlap but on other parts are quite different, for example in the categories 'Residues' or 'Recommendations'. Beekeeping management differentiation and philosophy, if categorized based on operation size, was also found in a study from the U.S., whereas the employed practices from commerical and sideline beekeepers did differ significantly from backyard beekeepers [@underwood2019].

Interestingly the category 'Economic Costs' was only picked by 6.4% of the participants, showing that expenses are only a low motivational driver for choosing a specific method. @thoms2019 showed in their U.S. survey that 'Money' was reported as one of the main disadvantages of managing varroa mites at fourth place after 'Kills, weakens, or hurts bees', 'Chemicals in colony' and 'Time'. As most beekeepers in Austria are hobbyists it is apparently of no surprise that economic is no major motivation factor for keeping bees and the expenses for treatment against varroa, even tough some can be substitutional, are low in comparison to other expenses, for example sugar or costs of hives. That the economic categories are more of a motivational factor for bigger operations (>25 colonies) can also be seen in the results, as the categories 'Economic Costs' and 'Time and number of applicants' were more frequently chosen.

The two least chosen two motivations 'Others' (0.5%) and 'Marketing' (0.2%) tell us that the selected 21 categories are complete and beekeepers in Austria are low influenced or seldom come into contact with advertisement of treatment methods. Although advertisement as such could also be hidden as recommendation in beekeeping journals.

## Data Distribution and Estimation 

The non normal distribution of reported treatment expenses per colony with a mean of EUR&nbsp;11.1 and median EUR&nbsp;9 was expected, because costs for the different treatment methods in investment, material and consumables vary (Fig.&nbsp;\@ref(fig:distr-year)). Although not directly comparable, in a US research about the pollination markets they went with an approximating of USD&nbsp;18 per colony for treatment consumables and application costs [@rucker2012]. Adjusted for inflation this would be USD&nbsp;44 which are EUR&nbsp;39.[^exchange] This estimate likely additionally includes the labour costs in application, which was not the case in our survey. On that note a survey from Albania showed that acaricides only account for a minor percentage of the production costs for honey and drug treatment in general did account on average for 5% of the total work load in beekeeping operations [@dedej2000]. Sadly the researcher do not report the data in sufficient manner to extract more information. Nevertheless, compared to the loss when colonies do not survive the winter, which was estimated with EUR&nbsp;318 in Austria [@popovskastojanov2021], the treatment expenses purely on an economic standpoint are rather worthwhile. At least if the treatment method is also efficient enough to prevent colony losses due to varroa mites. Furthermore, this is only the direct economic loss for the beekeeper, as the indirect socio-economic loss is even much higher due to honeybees important activity as pollinator for our food supply [@southwick1992].

As there are no previous market analyses or pilot studies in Austria, I implemented the calculated estimates to successfully validate the survey results for plausibility (Table&nbsp;\@ref(tab:table-estimates)). The overlap of survey responses and estimated expenses, despite the fact that they are on average similar, to show some differences when looking at the survey treatment method combinations. One of these overestimations are combinations including 'Another Biotechnical Method', excluding the separately asked methods 'Hyperthermia' and 'Drone brood removal'. One reason for the wrong estimation is that the method itself includes a broad range of applications and was difficult to calculate, eg. less expensive small plastic clip for queen caging compared to the more expensive full frame stainless steel cages. Another reason could be the more sophisticated calculation of the yearly depreciation of investment and material costs. The underestimation of the two exclusive use of formic acid treatment methods could be due to the application frequencies, as in the estimate I only calculate one application per month. Additionally, we only compared in this case the arithmetic mean expenses, which is influenced by outliers as shown by the underestimated treatment combination of summer and winter oxalic acid trickling of ready-made mixture. Although a lot of care was taken in data cleaning and validating before the analysis, this is one of the reasons why only treatment method combinations with at least 15 answers were further analysed.

[^exchange]: Calculation date: 13th December 2021. Inflation based on Consumer Price Index (CPI) with 58.64% since 2002. Conversion with exchange rate published by central banks 1&nbsp;USD&nbsp;=&nbsp;0.886682&nbsp;EUR.

## Extrapolation

The varroa treatment agent market was to my knowledge never before estimated in Austria. The total market value increased with the growing number of colonies registered in Austria over the three survey years. In total value, as a rough extrapolation, this is sums to expenses of EUR 3 to 5 million for each year. That these numbers could be off is shown by the interpolation of the survey total expenses, where the value of the real survey total sum lies between the two applied methods (Table&nbsp;\@ref(tab:extrapolation-tab)). The arithmetic mean as one method, which is influenced by high outliers and the geometric mean as second method, which probably underestimates as it is higher influenced by lower values in our case. Nevertheless, the extrapolation shows a possible economical burden for Austrian beekeepers as direct impact of an introduced aggressor since its extend into Europa in the 1980s [@griffiths1981;@potts2010a]. This calculated value does not even account for the hours of labour, which each treatment method requires. The expenses for treating against varroa mites and preventing possible colony losses is still lower than the loss of the colony if no prevention against the varroa mite was carried out. As shown in a multi-country study the estimated combined national economic losses due to colony losses and weak colonies over winter in the year 2016/17 sums to an estimate of EUR&nbsp;32 million in Austria [@popovskastojanov2021]. As a matter colony losses are not alone caused by varroa mites, but comprise of a combination of multiple abiotic and biotic factors [@neov2019]. Even so, a acute degree of parasitisation is correlated with multiple of negative effects on the honeybee colony and their individuals and in the worst case it can lead to their death [e.g. @rosenkranz2010;@noel2020;@morawetz2019;@traynor2020].

## Operational Factor Analysis

All selected operational questions did show in the two levels factor analysis a statistically significant difference in their expenses per colony. The highest difference was between small beekeeping operations with 25 or less colonies, which can be seen as recreational beekeepers, and bigger operations with more colonies, who can be seen as side-line and more professional beekeepers. As for winter colony losses it was shown before that bigger sized operations are loosing less colonies [e.g. @lee2015;@brodschneider2018;@gray2020]. The reasons for lower losses and lower expenses could be in their different beliefs and following practices to manage the varroa mite [@thoms2019;@underwood2019]. Bigger operations are financially driven and and their goals are probably more economical [@underwood2019]. Additionally some investments are more expensive for smaller operations, as you can only use them for a smaller number of colonies, e.g. sublimation tool for oxalic acid.

That the other two operational question come to the conclusion that migratory beekeepers and certified organic beekeepers are also spending less than the counterpart, can only partly be explained by the interaction of these groups and the operation size (Fig.&nbsp;\@ref(fig:distr-groups)). As the third and fourth most common group in our survey are recreational beekeepers (<= 25 colonies) which are certified organic or migratory beekeepers. Both operational factors are probably an indicator for experience and professionalism, which is also believed to be one of the main reasons for lower colony losses [@lee2015;@underwood2019;@oberreiter2020]. This professionalism which leads to lower loss rates could also lead to a more economical colony management. @underwood2019 showed in their beekeeping management study in the U.S. that organic beekeepers tend to use biotechnical and alternative treatment methods. The outcome of lower expenses for certified organic beekeeper was not expected, as I expected higher expenses due to higher uptake of biotechnical methods. As shown in the estimate calculation, these methods were overestimated by me and in the later section of efficacy and economics it is shown that 'Another Biotechnical Method' treatments are actually less expensive than the other ones applied. On this note, the difference in expenses in the two levels for the two operational factors is lower than the operational size and in the multiple regression model the effect is too small to be of significant impact.

## Multiple Regression

The accuracy of the selected factors in a multiple regression model for estimating correctly the expenses per colony is rather low. If the only goal is to predict the expenses, adding additional features from the survey and using more sophisticated methods like supervised ML or a 'black box'-variant could be a viable option [@sarker2021]. My inference based regression approach is simple and the sample size is limiting. The advantage of the selected linear regression models is that it can accommodate both prediction and explanation. Additionally, more samples or in my case participants would probably not increase the model fit, as the variation inside the selected factors is too spread, but better analysis of some treatment methods with low rates of application would be possible.

Nevertheless, the results from the regression model with number of colonies indicate the low explanatory value of the two operational question 'Certified Organic Beekeepers' and 'Migratory Beekeeper' as additional covariates. The two factors were previously discussed in the operational factor analysis as single factor analysis. In the model and now with finer resolution of colony numbers, than in the two level factor analysis, the possible correlations between the operational questions and colony numbers are probably one of the reasons for the low gains.

In addition the model shows the variation of expenses per colony inside the treatment methods. The asked methods in the questionnaire to include a wide variety of possible tools and treatment application. For example different investment into equipment for 'Hyperthermia' or frequency and application tools of 'Formic acid' treatment.

Interesting is that 'Another Biotechnical Method' is the only treatment method, besides 'Lactic Acid', which reduces the estimate if applied. The same trend can be seen in the later discussed efficacy and economical section. The treatment methods 'Synthetic Methods', 'Hyperthermia' and 'Thymol' did increase the estimate for expenses per colony if applied the most. This products are compared to organic acids are more expensive as consumables or for 'Hyperthermia' in investment. From the three different available oxalic treatment options the 'Oxalic acid - sublimation' did increase the estimate the least, showing that it is a less expensive option compared two the ready-made mix or pure trickling solution. Although, the investment of sublimation tools is more expensive, it can be used over multiple years and for multiple colonies and the low amount of oxalic acid needed would possible offset this investment quickly. On the efficiency side in field studies the sublimation of oxalic acid showed better results compared to trickling or no treatment [@altoufailia2015].


## Efficacy & Economical

In Austria it is common to apply organic acids (eg. formic acid, oxalic acid) as varroa treatment, whereas oxalic acid is more or less the sole option for the last treatment in autumn as shown in the network analysis (Section \@ref(network)) and was published previously [@oberreiter2020]. Although, alternative for example 'Hyperthermia' or synthetic methods to exist, they are only applied by a minority. In general each treatment method has their own strengths and weaknesses. As some treatment methods are more dependent on outside environmental conditions like temperate or humidity [@rosenkranz2010;@noel2020]. The multiple year approach allows me to find out which treatment methods are efficient on average over multiple years. Nevertheless, the interpretation of the results in my case are on point estimates but confidence intervals for loss rates are mostly overlapping between the analysed treatment combinations. The same is true for the expenses per colonies, as the variation is high and low sample size for some treatments resulting in wide standard errors.

The most common treatment combination over the three years of survey was long term formic acid in summer followed by oxalic acid trickling in winter (**A** 'SU-Fa-LT & WI-Ox-pure'). The combination of formic acid in summer and oxalic acid in winter is the national recommendation for Austrian beekeepers [@moosbeckhofer2015]. The combination showed on average consistently low loss rates but only in one year low expenses. Formic acid evaporation is known to be an effective treatment variant with a low risk of residues or resistances [@rosenkranz2010;@noel2020]. The long term treatment variant compared to to the short term variant with the same combination, which showed in each year higher losses (**E** 'SU-Fa-ST & WI-Ox-pure'), shows a better efficiency. This was also observed previously in Austria [@oberreiter2020]. The difference could stem from the disadvantages of using formic acid as it is dependent on outside temperature and humidity and can have considerable negative impact on the health of honeybees in high concentrations [@bolli1993;@underwood2003;@noel2020]. The long term evaporation of formic acid is typically used with applicators (evaporators) designed for a more controlled release of the fumes due to smaller exposure area, resulting in slower release and maybe better efficiency [@underwood2003]. The disadvantages of the short term formic acid treatment could be minimized when used in combination with sublimation of oxalic acid in summer and winter, as it did show promising results in both efficacy and expenses (**K** 'SU-Fa-ST & SU-Ox-sub & WI-Ox-sub'). One of the least efficient and expensive treatment method was the single application of formic acid long term treatment in summer (**M** 'SU-Fa-LT'). This shows that only a single treatment in summer is not enough and a treatment in late autumn is needed or at least needs to be evaluated for example with IPM methods [@roth2020;@bubnic2021].

Interesting was that the trickling of oxalic acid as pure or ready-made mixture did perform in the same combination better, but with overlapping confidence intervals, than the oxalic acid sublimation (**B** 'SU-Fa-LT & WI-Ox-sub'). In addition oxalic acid sublimation in summer and winter (**F** 'SU-Ox-sub & WI-Ox-sub') was less expensive but on average resulted in higher loss rates. In a direct comparison the sublimation of oxalic acid did archive better results than the trickling or spraying [@altoufailia2015]. In this regard oxalic acid does only work effectively in bloodless periods [@rosenkranz2010;@noel2020]. Therefore, I expect that beekeeper performing in summer the sublimation repeat it multiple times to be effective, but we do not have any more information on the frequency or how they applied it.

One way to minimize the amount of acaricides usage is with biotechnical methods. These combinations, without drone brood removal or hyperthermia, were on average more economically than the formic acid treatments. On efficacy the point estimates showed for the combinations using biotechnical methods higher loss rates. Previous studies in Austria showed that the biotechnical methods in summer, in direct comparison to the beekeepers who don't use these methods, resulted in significant lower observed loss rates [@crailsheim2018;@oberreiter2020]. Disadvantageously is that the methods itself are often labour intensive and the positives effect of being more economically without labour costs could be diminished if one accounts for time spend on manual labour [@rosenkranz2010;@buchler2020]. Nevertheless, the usage of biotechnical methods showed in field studies promising results and can be seen as a good alternative to organic acid or synthetic treatment and as an addition for integrated pest management (IPM) [@lodesani2019;@buchler2020].


<!--chapter:end:04-disc.Rmd-->

# Conclusion

The total economic burden by the introduced aggressor *Varroa destructor* are substantial on the other side the importance of honeybees for the society is indisputable. The results of the survey gives a first ever estimation and extrapolation of the treatment expenses for beekeepers in Austria. It underlays empirically a small fraction of this burden with sums already to a few million Euro on national scale. The motivations why beekeepers choose their treatment method showing the high resonance of intrinsics motives for their honeybees, 'Side effects on bees', and awareness of 'Residues'. It also shows a differences in goals and beliefs of smaller hobbyist beekeepers, who depend more on external information and less motivated by time or costs, which is more important for bigger operation who are probably more financial driven. 

In Austria the domination of organic acids as treatment methods is probably reflected as part of their motivations. Single treatments in summer with formic acid showed once again not enough efficacy on average. The most common treatment combination of formic acid long term and oxalic acid trickling in winter results on  consistently lower loss rates. Without labour costs the uptake of biotechnical methods seems to be more economical and depending on the method it may is less harmful for honeybees with no risk of residues or resistances. The reduction of pesticides should be a main goal, in which biotechnical methods can help. As the treatment against varroa mites is now since over 40 years in Europa the dominant topic in beekeeping, a lot of effort and time is invested by breeder and scientist to select for honeybees who can defend or live with the aggressor without treatment, but so far without any major adaptation or success [@mondet2020a;@mondet2020;@guichard2020]. The economic way of beekeeping with high density of colonies and swarm reduction techniques will probably always rely on active pest control. Nevertheless, natural directed selection may be a solution to keep healthier honeybee colonies [@seeley2007a;@locke2016;@neumann2017;@buchler2020;@oddie2021]. This phenomena is already observed as several honeybee populations in Europe that actively hinder the Varroa mite in its development through their behaviour and they may also exist in Austria [@mondet2020;@oberreiter2021].

The uncertain range of the my results and the big effort on data cleaning did show some disadvantages of our survey. An alternative approach for evaluating the economic values or impact would be to work more closely with beekeeping operations for a thorough result. This was done for example in a detailed economic breakdown of honeybee queen breeding in Canada [@bixby2020]. Although, the advantages of the survey are probably higher, e.g. using an already established survey and getting more answers, lower workload and costs to get the data. My thesis is showing the importance of ongoing surveys to evaluate and observe trends and changes in the beekeeping community. For the beekeepers the results are information of their coequal and for national beekeeping clubs and politics the results can be used to find out more about the needs in the beekeeping community. 

<!--chapter:end:05-conclusio.Rmd-->

# (APPENDIX) Appendix {-}
# Appendix

(ref:size-permutation) Median difference null distribution (gray bars) based on 5,000 permutations. Red line is the point estimate for the median difference with yellow area as the 95% confidence interval. Red shadow indicating the calculated two-sided *p*-value range on the null distribution. 

(ref:size-permutation2) The *p*-values reported as "\*\*\*" would need more permutation to calculate a precise *p*-value.

```{r size-permutation, out.width="100%", include=T, fig.cap="Two level factor analysis statistical test results for 'Operation Size'. (ref:size-permutation) (ref:size-permutation2)"}
include_custom("output/figs/size-permutation.png")
```

```{r organic-permutation, out.width="100%", include=T, fig.cap="Two level factor analysis statistical test results for 'Certified Organic Beekeeper'. (ref:size-permutation)"}
include_custom("output/figs/organic-permutation.png")
```

```{r migratory-permutation, out.width="100%", include=T, fig.cap="Two level factor analysis statistical test results for 'Migratory Beekeeper'. (ref:size-permutation)"}
include_custom("output/figs/migratory-permutation.png")
```

```{r model-qq-residuals, out.width="80%", include=T, fig.cap="Distribution as QQ-plot for the model residuals after fitting the selected best model 'Base + Treatment Methods' on the training dataset."}
include_custom("output/figs/model-qq-residuals.png")
```

```{r motivation-state-rank, include=T}
cn <- c("State", "Category", "Answers [%]", "Rank")
c <- "Top 7 motivation answers for choosing varroa control method for each state separately. Participants could select out of a pool of 21 categories and where able to pick up to 5 categories maximum. 'Answers' is the relative amount the corresponding category was picked. Rank simply the decreasing order of the relative value in each state."
tab <- knitr::kable(
    r_motivation$state_rank,
    align = c("l", "l", "r", "c"),
    booktabs = TRUE,
    row.names = FALSE,
    col.names = cn,
    caption = c
) %>%
    kable_styling(latex_options = "scale_down") %>%
    kableExtra::collapse_rows(
        1,
        valign = "top",
        latex_hline = "major",
    )
tab
rm(tab, c, cn)
```

```{r motivation-comb-full, include=T}
cn <- c("Motivation", "Answers", "Motivations")
c <- "Top 25 motivation answers combinations for choosing varroa control method given by participants. Participants could select out of a pool of 21 categories and where able to pick up to 5 categories maximum. 'Answers' is the total number this combination was picked and 'Motivations' are the number of different categories in this combination, whereas 1 would indicate a single answer."
tab <- knitr::kable(
    r_motivation$comb_list %>% select(motivation, n, motivations) %>% dplyr::slice_max(n, n = 25),
    align = c("l", rep("r", 2)),
    booktabs = TRUE,
    row.names = FALSE,
    col.names = cn,
    # format = "markdown",
    caption = c
) %>%
    kable_styling(latex_options = "scale_down")
tab
rm(tab, c, cn)
```




<!--chapter:end:06-appendix.Rmd-->

