---
title: "Treatment Cost in Austria"
subtitle: 'University of Graz, Institut for Biology'
author: "Hannes Oberreiter & Robert Brodschneider"
date: "2020"
header-includes:
    - \usepackage{hyperref}
    - \usepackage{float}
    - \usepackage{caption}
    - \captionsetup[figure]{font=small}
output:
  html_document
bibliography: bib.bibtex
biblio-style: authoryear
link-citations: true
urlcolor: blue
abstract: Summary of Stats and Statistics for Treatment Costs Varroa. Our Hypothesis is that the costs have an influence or correlated to the winter losses. H0 = Costs have no influence on survival rate, H1 = Costs have an influence on the survival rate.
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  fig.pos   = "H", 
  out.extra = "", 
  fig.align = "center", 
  out.width = "100%",
  echo      = FALSE,
  include   = FALSE
  )

# Our Libraries
VLIBS <- c("tidyverse", "here", "readxl", "bookdown", "pander")

F_LoadLibs <- function(x){
  print(x)
  if(!require(x, character.only = TRUE)) install.packages(x); library(x, character.only = TRUE);
  return(T)
}

sapply(VLIBS, F_LoadLibs)
rm(VLIBS, F_LoadLibs)
```

```{r load-files, warning = FALSE}
source("partials/files.R", local = knitr::knit_global())

# Extract only answers with valid answer for costs
DATA <- RAW[!is.na(RAW$costs),]
```

# Cleanup

## No Treatment Answers

```{r}
# Extract No Treatment Answers
NO_TREATMENT <- DATA[DATA$varroa_treated != "Ja",]
NO_TREATMENT <- NO_TREATMENT[,c("id", "costs", "varroa_treated", "comments", "year", "T_amount", "c_short")]
DATA <- DATA[!(DATA$id %in% NO_TREATMENT$id),]
```

In our survey `r nrow(NO_TREATMENT)` did answer "No Treatment against Varroa" and did give no answers for any treatment method, but included costs for varroa treatment. These `r nrow(NO_TREATMENT)` illogical entries were removed from our data analysis.

```{r}
rm(NO_TREATMENT)
```

## Zero Cost Answers

```{r}
# Entries with zero costs
COST_ZERO <- DATA[DATA$costs == 0,]
COST_ZERO <- COST_ZERO[,c("id", "costs", "varroa_treated", "comments", "year", "T_amount", "c_short")]
# sponsorship (e.g. Imkereiförderung, Gemeinde)
id_sponsor <- c(353, 752, 855, 1624)
# biotechnical & hyperthermie
id_keep <- c(1750)
id_remove <- COST_ZERO$id[!(COST_ZERO$id %in% id_keep)]
DATA <- DATA[!(DATA$id %in% id_remove),]
```

In total `r nrow(COST_ZERO)` participants said their treatment costs were zero. Only `r length(id_keep)` answers were kept, as the treatment was a combination of hyperthermia and biotechnical. We have no assumption. Ofcourse the equipment for the hyperthermia do costs something, but we don't know more details. `r length(id_remove)-length(id_sponsor)` were removed as they said zero costs but actually did treatment with organic acids. In addition `r length(id_sponsor)` participants got a sponsorship (eg from beekeeping club or local community) for their treatment costs, which were also removed from the cost analysis dataset.

```{r}
rm(COST_ZERO, id_sponsor, id_remove, id_keep)
```

## High Cost Answers

```{r}
csum <- summary(DATA$costs)
upper_limit <- csum[5] * 10
COST_UPPER <- DATA[DATA$costs > upper_limit,]
COST_UPPER <- COST_UPPER[,c("id", "hives_winter", "hives_lost", "costs", "t_estimated", "varroa_treated", "comments", "year", "T_amount", "c_short")]

# Remove these entries, as they make sense
id_nochange <- c(3030, 3959)
COST_UPPER <- COST_UPPER %>% filter(!(id %in% id_nochange))

# Calculate new prices
COST_UPPER$new <- COST_UPPER$costs / COST_UPPER$hives_winter
DATA$costs[(DATA$id %in% COST_UPPER$id)] <- COST_UPPER$new
```

We defined upper limit cost outliner with 3rd quantile and multiplied by 10 which resulted in a cost limit of `r upper_limit`. In total `r nrow(COST_UPPER)` participants did answers above this limit. Upon investigating the `r nrow(COST_UPPER)` answers, we found out that most of them did answer a total cost and no cost by colony. Therefore, we divided the costs by the number of colonies going into winter. Only `r length(id_nochange)` participants were not changed, as one only had one colony and did hyperthermia treatment and the second one did explicit explain the high costs as comment.

```{r}
rm(id_nochange, upper_limit, COST_UPPER, csum)
```
# Standard Params

```{r calc-n, include=FALSE}
RAW_REPORTS <- RAW %>% group_by(year) %>% summarize(
  n = n()
)

DATA_REPORTS <- DATA %>% group_by(year) %>% summarize(
  n = n()
)

REPORTS <- tibble(
  Year = RAW_REPORTS$year,
  Total = RAW_REPORTS$n,
  Cost = DATA_REPORTS$n,
  Percent = round(DATA_REPORTS$n/RAW_REPORTS$n*100,1)
)
rm(DATA_REPORTS, RAW_REPORTS)
```

The number of reports which did include costs and after cleanup of invalid or illogical answers was similar for the survey of 2018/19 with `r REPORTS$Percent[1]`% and for 2019/20 with `r REPORTS$Percent[2]` \@ref(tab:reports).

```{r reports, include=T}
cn = c("Year", "Total [n]", "Answered [n]", "Percent [%]")
knitr::kable(
  REPORTS,
  booktabs = TRUE, row.names = FALSE,
  col.names = cn, format = "markdown"
)
rm(REPORTS, cn)
```

```{r summary, include=T}
cap = "Table the total descriptive statistics for costs in our survey and estimates based on our standard cost estimates, see table x."

cs <- summary(DATA$costs)
es <- summary(DATA$t_estimated)

ct <- tibble(
  Type = c("Survey", "Estimated"),
  min  = c(cs[1], es[1]),
  Q1   = c(cs[2], es[2]),
  Median = c(cs[3], es[3]),
  Mean = round(c(cs[4], es[4]), 2),
  Q3 = round(c(cs[5], es[5]), 2),
  Max = c(cs[6], es[6])
)


knitr::kable(
  ct,
  booktabs = TRUE, row.names = FALSE,
  format = "markdown",
  caption = cap
)

rm(cap, cs, es, ct)

```

# Standard Cost Estimation

To calculate a standard value for data cleaning, we researched the official price in 2020 for several products and average usage amount, see table \@ref(tab:standard-cost).

```{r standard-cost, include=T}
c <- "Our calculated estimates for treatment. Investments can be used long term (we calculate 5 years, eg. Hyperthermia) and are divided by the number of colonies. Material can also be used long term (~ 3 years, eg. Queen Cages) but needs to be bought for each colony. Consumables are used for each colony (eg. Formic Acid, 50ml)."
knitr::kable(
  treatmentList[-1, c(3, 5:7)],
  booktabs = TRUE, row.names = FALSE,
  caption = c
)
rm(c)
```


```{r}
# TODO my AS-KZ is too low
d <- DATA %>% group_by(t_short) %>% summarise(
  c_mean = mean(costs),
  e_mean = mean(t_estimated),
  n = n()
)


```




# Colonies and Costs Linear Modell





