\pdfbookmark[0]{Acknowledgement}{Acknowledgement}
\addcontentsline{toc}{chapter}{Acknowledgements}
\chapter*{Acknowledgement}
\label{sec:acknowledgement}
\vspace*{-10mm}

Before all else I would like to express my deepest love for Mag.\,Anna Dünser\,MSc, who is for me not only an emotional support, but also an excellent scientific discussion partner and proof reading expert. Additionally, my family who are always open minded and support me in any circumstances. Johannes Schlagbauer\,BSc for always being a helpful student colleague. I would like to thank Prof.\,Dr.\,Edith Stabentheiner for aiding me in my endeavour to obtain the approval for my individual master programme. The University of Graz and Technical University of Graz for providing the environment for my education. The manifold unsung heroes of the internet, who answer questions on Stack Overflow or create amazing open-source software, for without them this work would simply be impossible. A particular thanks to the 'R4DS Online Learning Community' which I'm a proud member of and August Comte for always answering my statistical questions.

Needless to say I could not have produced my master thesis without the freedom given by my supervisor Dr.\,Robert Brodschneider, the driving person behind the annual survey on winter colony losses in Austria and who was always available if needed. My appreciation for the effort of the COLOSS Network in their active honey bee research. Furthermore and the most formal thanks belongs to the participating beekeepers who were willing to help with the research about honey bees and by this helping me directly with my master thesis.